---
title: "Cleaning WCO Data and adding Tag info"
output: wco_detections+tags_20240613.csv
Author: "Morgan Youngblood"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

# Load Libraries to read csv and merge/ clean data
```{r}
library(dplyr)
library(readr)
library(tidyr)
```

# Read in the WCO data
```{r}

tag_data <- read_csv("outputs/updated_tag_id_20240621.csv")
str(tag_data)


wco_raw <- read_csv("data/hollings_detection-data_cinms_wco_20240617.csv")
str(wco_raw)
problems(wco_raw)
head(wco_raw)
```

# WCO Data Cleaning and Setup

## Remove NA columns in the WCO data
```{r}
# Find columns where all values are NA
columns_to_remove <- sapply(wco_raw, function(x) all(is.na(x)))

columns_to_remove

# Remove Transmitter Type and Sensor Precision columns
wco_clean <- select(wco_raw, -'Transmitter Type', -'Sensor Precision')

colnames(wco_clean)
```

## Set up time columns
```{r}
# Convert Date and Time (UTC) to local timezone
wco_clean$Local_Time <- as.POSIXct(wco_clean$`Date and Time (UTC)`, tz = "America/Los_Angeles")

# Extract date component
wco_clean$Date <- as.Date(wco_clean$Local_Time)

# Extract month and year from Date
wco_clean$Month <- format(wco_clean$Date, "%m")
wco_clean$Year <- format(wco_clean$Date, "%Y")
wco_clean$Hour <- format(wco_clean$Local_Time, "%H")

# Define seasons based on the month
wco_clean$Season <- with(wco_clean, ifelse(Month %in% c("03", "04", "05"), "Spring",
                        ifelse(Month %in% c("06", "07", "08"), "Summer",
                        ifelse(Month %in% c("09", "10", "11"), "Fall",
                        "Winter"))))

head(wco_clean)

colnames(wco_clean)

# Evaluate summary statistics for the new columns
summary(wco_clean$Local_Time)
summary(wco_clean$Date)
summary(wco_clean$Month)
summary(wco_clean$Year)
summary(wco_clean$Hour)

```

### Filter Spurious detections using the date
```{r}

# Filter detections where there are < 2 detections per day
wco_clean <- wco_clean %>%
  group_by(Date) %>%
  filter(n() > 1)

  # Filter for detections > 1
wco_clean <- wco_clean %>%
  group_by(Transmitter) %>%
  filter(n() > 1) %>%
  ungroup()

# Evaluate the number of detections filtered
nrow(wco_clean)
nrow(wco_raw)

```

## Set up spatial data

### Group stations by island
```{r}

# Create a mapping of stations to islands
station_to_island <- data.frame(
  `Station Name` = c("WCO_Cuyler", "WCO_Tyler Bight",
                     "WCO_Carrington Point", "WCO_Johnson's Lee",
                     "WCO_Hazards_Profile", "WCO_Trailers",
                     "WCO_MorsePoint", "WCO_Valley", "WCO_Smugglers",
                     "WCO_Scorpion", "WCO_Pelican",
                     "WCO_Winfield Scott", "WCO_Anacapa South",
                     "WCO_Mohawk LTER", "WCO_Carpinteria"),
  Island = c("San Miguel", "San Miguel",
             "Santa Rosa", "Santa Rosa",
             "Santa Cruz", "Santa Cruz",
             "Santa Cruz", "Santa Cruz", "Santa Cruz",
             "Santa Cruz", "Santa Cruz",
             "Anacapa", "Anacapa",
             "Mainland", "Mainland")
)

# Ensure column names match exactly
colnames(station_to_island) <- c("Station Name", "Island")

# Perform the join to add island information
wco_clean <- wco_clean %>%
  left_join(station_to_island, by = "Station Name")

colnames(wco_clean)

# Make sure Islands and Stations are matched correctly
unique(wco_clean$Island)
unique(wco_clean$`Station Name`)

# Make all stations where Island is NA 'Santa Barbara' Island
wco_clean$Island[is.na(wco_clean$Island)] <- "Santa Barbara"

# Group by Island and count unique stations
island_station_count <- wco_clean %>%
  group_by(Island) %>%
  summarise(Unique_Stations = n_distinct(`Station Name`))

# Print the table
island_station_count

# Set levels for Islands (West to East)
island_order <- c("San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Mainland", "Santa Barbara")

# Group by Station and find the longitude and latitude of each
station_location <- wco_clean %>%
  group_by(`Station Name`) %>%
  summarise(Longitude = mean(Longitude, na.rm = TRUE),
            Latitude = mean(Latitude, na.rm = TRUE))

# Print the full table
View(station_location)

# Set levels for stations based on longitude (West to East)
station_order <- station_location %>%
  arrange(Longitude) %>%
  pull(`Station Name`)

  # Convert Island and Station Name to factors with specified order
  wco_clean <- wco_clean %>%
    mutate(Island = factor(Island, levels = island_order),
           `Station Name` = factor(`Station Name`, levels = station_order))

```

# Finally we can merge the tag data and write the .csv
```{r}

# Merge the tag_data into the wco_clean dataset based on the Transmitter column
wco_merged <- left_join(wco_clean, tag_data, by = "Transmitter")

# Verify the structure of the merged dataset
str(wco_merged)

# Export the merged dataset to a CSV file
write_csv(wco_merged, "data/wco_detections+tags_20240618.csv")

```

# let's look for orphaned tag IDs in the WCO data to see if we can identify unique recievers and months associated with the detection data to make a vrl file package. 
```{r}

# Create a new dataset with unique transmitters and the list of unique month-year combos detected, including Station.Name
orphaned_transmitter_months <- wco_merged %>%
  filter(is.na(SciName)) %>%
  group_by(Transmitter, Receiver, Year, Month) %>%
  summarise(Month_Year = paste(Year, Month, sep = "-"), 
            Station.Name = first(`Station Name`), # Assuming `Station Name` is unique or taking the first one
            .groups = 'drop') %>%
  ungroup() %>%
  select(Transmitter, Month_Year, Receiver, Station.Name) %>%
  distinct() %>%
  arrange(Month_Year)

View(orphaned_transmitter_months)

# Export the orphaned transmitter months to a CSV file
write_csv(orphaned_transmitter_months, "outputs/orphaned_transmitter_months_20240701.csv")

# Determine the number of unique Receiver-Month_Year combinations in the dataset
unique(paste(orphaned_transmitter_months$Month_Year, orphaned_transmitter_months$Receiver)) %>% length()

# Make a list of these unique combinations
unique_receiver <- unique(paste(orphaned_transmitter_months$Month_Year, orphaned_transmitter_months$Receiver))

# make a dataframe of the unique receiver-month-year combinations
unique_receiver <- data.frame(unique_receiver)

View(unique_receiver)

# Export the unique receiver-month-year combinations to a CSV file
write_csv(unique_receiver, "outputs/unique_receiver_month_year_20240701.csv")
```