---
title: "NOAA DDT DATA"
author: "MORGAN YOUNGBLOOD"
date: "05/21/2024"
output:
  html_document: default
  pdf_document: default
---

#_______________
```{r setup, include=FALSE}

# Set the root directory for notebook chunks
knitr::opts_knit$set(root.dir = "Documents/Personal/NOAA/Hollings Project/Analysis")

# Set the working directory for the session
setwd("Documents/Personal/NOAA/Hollings Project/Analysis")

# Verify the working directory
getwd()
```

# Load packages
```{r packages, warning=FALSE, message=FALSE}
#install.packages("fuzzyjoin")

library(readr)
library(gganimate)
library(lubridate)
library(adehabitatHR)
library(adehabitatHS)
library(adehabitatLT)
library(adehabitatMA)
library(deldir)
library(dplyr)
library(sp)
library(ggplot2)
library(sf)
library(ggmap)
library(ggplot2)
library(stats)
library(viridis)
library(geosphere)
library(zoo)
library(gifski)
library(fuzzyjoin)

```

# Import Data
```{r, warning=FALSE}

library(readr)

WCO <- read_csv("hollings_detection-data_cinms-wco_20231024_01.csv")
head(WCO)
str(WCO)

```
#_______________
# Data Cleaning and Organizing

## Setup time data 
```{r, warning=FALSE}

# Convert Date and Time (UTC) to local timezone
WCO$Local_Time <- as.POSIXct(WCO$`Date and Time (UTC)`, tz = "America/Los_Angeles")

# Extract date component
WCO$Date <- as.Date(WCO$Local_Time)

# Extract month and year from Date
WCO$Month <- format(WCO$Date, "%m")
WCO$Year <- format(WCO$Date, "%Y")
WCO$Hour <- format(WCO$Local_Time, "%H")

# Define seasons based on the month
WCO$Season <- with(WCO, ifelse(Month %in% c("03", "04", "05"), "Spring",
                        ifelse(Month %in% c("06", "07", "08"), "Summer",
                        ifelse(Month %in% c("09", "10", "11"), "Fall",
                        "Winter"))))

head(WCO)

```

## Clean Data
```{r, warning=FALSE}

# Remove columns with all NA values
WCO <- WCO %>% select_if(~ !all(is.na(.)))

```

## Inspect Station Data and Group Associated Islands
```{r, warning=FALSE}

# Define the order of islands and stations in descending order
island_order <- c("San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Mainland")
station_order <- c("WCO_Cuyler", "WCO_Tyler Bight", "WCO_Carrington Point", "WCO_Johnson's Lee",
                   "WCO_Hazards_Profile", "WCO_Trailers", "WCO_MorsePoint", "WCO_Valley",
                   "WCO_Smugglers", "WCO_Scorpion", "WCO_Pelican", "WCO_Winfield Scott",
                   "WCO_Anacapa South", "WCO_Mohawk LTER", "WCO_Carpinteria")

# Inspect Station Data and Group Associated Islands
# Get unique stations
unique_stations <- WCO %>%
  select(`Station Name`, Latitude, Longitude) %>%
  distinct()

# Display the unique stations and their corresponding lat/long
print(unique_stations)

# Count the number of unique stations and unique lat/long combinations
num_unique_stations <- n_distinct(WCO$`Station Name`)
num_unique_latlong <- n_distinct(paste(WCO$Latitude, WCO$Longitude))

cat("Number of unique stations: ", num_unique_stations, "\n")
cat("Number of unique latitude/longitude combinations: ", num_unique_latlong, "\n")

# Create a mapping of stations to islands
station_to_island <- data.frame(
  `Station Name` = c("WCO_Cuyler", "WCO_Tyler Bight",
                     "WCO_Carrington Point", "WCO_Johnson's Lee",
                     "WCO_Hazards_Profile", "WCO_Trailers",
                     "WCO_MorsePoint", "WCO_Valley", "WCO_Smugglers",
                     "WCO_Scorpion", "WCO_Pelican",
                     "WCO_Winfield Scott", "WCO_Anacapa South",
                     "WCO_Mohawk LTER", "WCO_Carpinteria"),
  Island = c("San Miguel", "San Miguel",
             "Santa Rosa", "Santa Rosa",
             "Santa Cruz", "Santa Cruz",
             "Santa Cruz", "Santa Cruz", "Santa Cruz",
             "Santa Cruz", "Santa Cruz",
             "Anacapa", "Anacapa",
             "Mainland", "Mainland")
)

# Ensure column names match exactly
colnames(station_to_island) <- c("Station Name", "Island")

# Perform the join to add island information
WCO <- WCO %>%
  left_join(station_to_island, by = "Station Name")

# Check if Island column exists after join
print(colnames(WCO))
print(head(WCO))

# Add the island information to the main dataset
if ("Island" %in% colnames(WCO)) {
  WCO <- WCO %>%
    mutate(Island = ifelse(is.na(Island), "Santa Barbara", Island))
} else {
  stop("The 'Island' column was not created in the join.")
}

# Convert Island and Station Name to factors with specified order
WCO <- WCO %>%
  mutate(Island = factor(Island, levels = island_order),
         `Station Name` = factor(`Station Name`, levels = station_order))

# Display the structure and first few rows of the updated dataset
str(WCO)
head(WCO)

```

## More cleaning and Remove Santa Barbara
```{r, warning=FALSE}

# Filter for detections > 1
WCO <- WCO %>%
  group_by(`Station Name`, Island, Transmitter) %>%
  filter(n() > 1) %>%
  ungroup()

# Remove the specified columns
WCO <- WCO %>%
  select(-c(`Transmitter Name`, `Transmitter Serial`, `Sensor Value`, `Sensor Unit`))

# Remove data associated with Santa Barbara Island
WCO <- WCO %>%
  filter(Island != "Santa Barbara")

# Display the structure and first few rows of the cleaned dataset
str(WCO)
head(WCO)

```

## Make tag ID lists
```{r}

LoweLabTagInfo <- readRDS("LoweLabTagInfo.rds")

# Perform an inner join to find matching transmitters
matching_Lowe_transmitters <- WCO %>%
  inner_join(LoweLabTagInfo, by = "Transmitter")

# Count the number of matching transmitters
n_matching_Lowe_transmitters <- nrow(matching_Lowe_transmitters)
print(n_matching_Lowe_transmitters)

# Count the number of unique matching transmitters
unique_matching_Lowe_transmitters <- matching_Lowe_transmitters %>%
  select(Transmitter) %>%
  distinct() %>%
  nrow()

print(unique_matching_Lowe_transmitters)

Tag_ID_List <- read_csv("Tag_ID_List.csv")
#View(Tag_ID_List)

# Merge the datasets on the Transmitter column
Tag_ID_6_1 <- Tag_ID_List %>%
  left_join(LoweLabTagInfo, by = "Transmitter")

# Combine SciName.x and SciName.y into a single SciName column
Tag_ID_6_1 <- Tag_ID_6_1 %>%
  mutate(SciName = if_else(is.na(SciName.y), SciName.x, SciName.y)) %>%
  select(-SciName.x, -SciName.y)

# View the merged data
str(Tag_ID_6_1)
head(Tag_ID_6_1)

# Export the merged dataset to a CSV file
write_csv(Tag_ID_6_1, "Updated_Tag_ID_6_1.csv")

# Filter the dataset for rows where SciName is NA
NA_Tag_ID_6_1 <- Tag_ID_6_1 %>%
  filter(is.na(SciName)) %>%
  select(Transmitter, First_Detection, Last_Detection)

# Export the filtered dataset to a .csv file
write_csv(NA_Tag_ID_6_1, "NA_Tag_ID_6_1.csv")

```

## Merge IDs to WCO Data
```{r}

# Merge the Tag_ID_6_1 data into the WCO dataset based on the Transmitter column
WCO <- WCO %>%
  left_join(Tag_ID_6_1, by = "Transmitter")

# Verify the structure of the merged dataset
str(WCO)

# Export the merged dataset to a CSV file
write_csv(WCO, "WCO_with_Tag_ID_6_6_24.csv")

```

#_______________
# Summary

## Transmitter overview
```{r}

# Detections per Transmitter
detections_per_transmitter <- WCO %>%
  group_by(Transmitter) %>%
  summarise(Detections = n())

print(detections_per_transmitter)

# Bar plot for Detections per Transmitter
ggplot(detections_per_transmitter, aes(x = reorder(Transmitter, -Detections), y = Detections)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Detections per Transmitter", x = "Transmitter", y = "Detections") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Detections per SciName
detections_per_SciName <- WCO %>%
  group_by(SciName) %>%
  summarise(Detections = n())

# Filter out rows where SciName is NA
detections_per_SciName <- detections_per_SciName %>%
  filter(!is.na(SciName))

print(detections_per_SciName)

# Bar plot for Detections per Transmitter
ggplot(detections_per_SciName, aes(x = reorder(SciName, -Detections), y = Detections)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Detections per Species", x = "Species", y = "Detections") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Unique transmitters per SciName
unique_transmitters_per_SciName <- WCO %>%
  filter(!is.na(SciName)) %>%  # Filter out rows where SciName is NA
  group_by(SciName) %>%
  summarise(Unique_Transmitters = n_distinct(Transmitter))

print(unique_transmitters_per_SciName)

# Bar plot for Unique Transmitters per SciName
ggplot(unique_transmitters_per_SciName, aes(x = reorder(SciName, -Unique_Transmitters), y = Unique_Transmitters)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Unique Transmitters per Species", x = "Species", y = "Unique Transmitters") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Count unique transmitters for each species and sex
count_sex_species <- WCO %>%
  group_by(SciName, sex) %>%
  summarise(Unique_Transmitters = n_distinct(Transmitter)) %>%
  ungroup()

# Filter out rows where sex is NA
count_sex_species <- count_sex_species %>%
  filter(!is.na(sex))

# Print the result
print(count_sex_species)

# Create the bar plot for unique transmitters per species and sex
ggplot(count_sex_species, aes(x = reorder(SciName, -Unique_Transmitters), y = Unique_Transmitters, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  labs(title = "Unique Transmitters per Species and Sex", x = "Species", y = "Number of Unique Transmitters") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Spatial Overview
```{r, warning=FALSE}

# Detections per Station
detections_per_station <- WCO %>%
  group_by(`Station Name`) %>%
  summarise(Detections = n())

# Detections per Island
detections_per_island <- WCO %>%
  group_by(Island) %>%
  summarise(Detections = n())

# Print summary statistics tables
print(detections_per_station)
print(detections_per_island)

# Bar plot for Detections per Station
ggplot(detections_per_station, aes(x = reorder(`Station Name`, -Detections), y = Detections)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Detections per Station", x = "Station Name", y = "Detections") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Bar plot for Detections per Island
ggplot(detections_per_island, aes(x = reorder(Island, -Detections), y = Detections)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Detections per Island", x = "Island", y = "Detections")

# Summarize detections per station
detections_per_station <- WCO %>%
  group_by(`Station Name`, Latitude, Longitude) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Read the shapefiles
islands <- st_read("cinms_py2/cinms_py.shp")
coast <- st_read("stanford-vx275xn8886-shapefile/vx275xn8886.shp")

# Create the scatter plot for detections per station with shapefile
map_total_detections <- ggplot() +
   geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the islands
  geom_point(data = detections_per_station, aes(x = Longitude, y = Latitude, color = Detections, size = Detections)) +
  scale_color_viridis_c() +  # Color gradient
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Detections per Station", x = "Longitude", y = "Latitude", color = "Detections", size = "Detections")

print(map_total_detections)

# Save the plot as an image
ggsave("total_detections_per_station_map.png", plot = map_total_detections, width = 10, height = 8, dpi = 300, units = "in")

```

### Transmitters per Location
```{r}

# Summarize number of unique transmitters detected per station
transmitters_per_station <- WCO %>%
  group_by(`Station Name`, Latitude, Longitude) %>%
  summarise(Transmitters = n_distinct(Transmitter)) %>%
  ungroup()

# Create the bar plot for transmitters detected per station
ggplot(transmitters_per_station, aes(x = reorder(`Station Name`, -Transmitters), y = Transmitters)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Unique Transmitters Detected per Station",
       x = "Station Name",
       y = "Number of Transmitters") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Summarize number of unique transmitters detected per island
transmitters_per_island <- WCO %>%
  group_by(Island) %>%
  summarise(Transmitters = n_distinct(Transmitter)) %>%
  ungroup()

# Create the bar plot for transmitters detected per island
ggplot(transmitters_per_island, aes(x = reorder(Island, -Transmitters), y = Transmitters)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Unique Transmitters Detected per Island",
       x = "Island",
       y = "Number of Transmitters") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Average Detections Per Transmitter at each Location
```{r, warning=FALSE}

# Calculate the number of detections per transmitter at each station
detections_per_individual <- WCO %>%
  group_by(`Station Name`, Transmitter) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Calculate average detections per individual, standard error, and sample size for each station
avg_detections_per_individual_per_station <- detections_per_individual %>%
  group_by(`Station Name`) %>%
  summarise(
    avg_detections = mean(Detections),
    se_detections = sd(Detections) / sqrt(n()),
    num_transmitters = n_distinct(Transmitter)
  ) %>%
  ungroup() %>%
  mutate(myaxis = paste0(`Station Name`, "\n", "n=", num_transmitters))

# Create the bar plot for average detections per individual with standard error
ggplot(avg_detections_per_individual_per_station, aes(x = reorder(myaxis, -avg_detections), y = avg_detections)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_errorbar(aes(ymin = avg_detections - se_detections, ymax = avg_detections + se_detections), 
                width = 0.2, color = "red") +
  theme_minimal() +
  labs(title = "Average Detections per Individual with Standard Error at Each Station",
       x = "Station Name",
       y = "Average Detections per Individual") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Calculate the number of detections per transmitter at each island
detections_per_individual_island <- WCO %>%
  group_by(Island, Transmitter) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Calculate average detections per individual, standard error, and sample size for each island
avg_detections_per_individual_per_island <- detections_per_individual_island %>%
  group_by(Island) %>%
  summarise(
    avg_detections = mean(Detections),
    se_detections = sd(Detections) / sqrt(n()),
    num_transmitters = n_distinct(Transmitter)
  ) %>%
  ungroup() %>%
  mutate(myaxis = paste0(Island, "\n", "n=", num_transmitters))

# Create the bar plot for average detections per individual with standard error for each island
ggplot(avg_detections_per_individual_per_island, aes(x = reorder(myaxis, -avg_detections), y = avg_detections)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_errorbar(aes(ymin = avg_detections - se_detections, ymax = avg_detections + se_detections), 
                width = 0.2, color = "red") +
  theme_minimal() +
  labs(title = "Average Detections per Individual with Standard Error at Each Island",
       x = "Island",
       y = "Average Detections per Individual") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Median detections per individual at each Location
```{r, warning=FALSE}

# Calculate the number of detections per transmitter at each station
detections_per_individual <- WCO %>%
  group_by(`Station Name`, Transmitter) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Calculate median detections per individual, IQR, and sample size for each station
median_detections_per_individual_per_station <- detections_per_individual %>%
  group_by(`Station Name`) %>%
  summarise(
    median_detections = median(Detections),
    IQR_lower = quantile(Detections, 0.25),
    IQR_upper = quantile(Detections, 0.75),
    num_transmitters = n_distinct(Transmitter)
  ) %>%
  ungroup() %>%
  mutate(myaxis = paste0(`Station Name`, "\n", "n=", num_transmitters))

# Create the bar plot for median detections per individual with IQR for each station
ggplot(median_detections_per_individual_per_station, aes(x = reorder(myaxis, -median_detections), y = median_detections)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_errorbar(aes(ymin = IQR_lower, ymax = IQR_upper), width = 0.2, color = "red") +
  theme_minimal() +
  labs(title = "Median Detections per Individual with IQR at Each Station",
       x = "Station Name",
       y = "Median Detections per Individual") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Calculate the number of detections per transmitter at each island
detections_per_individual_island <- WCO %>%
  group_by(Island, Transmitter) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Calculate median detections per individual, IQR, and sample size for each island
median_detections_per_individual_per_island <- detections_per_individual_island %>%
  group_by(Island) %>%
  summarise(
    median_detections = median(Detections),
    IQR_lower = quantile(Detections, 0.25),
    IQR_upper = quantile(Detections, 0.75),
    num_transmitters = n_distinct(Transmitter)
  ) %>%
  ungroup() %>%
  mutate(myaxis = paste0(Island, "\n", "n=", num_transmitters))

# Create the bar plot for median detections per individual with IQR for each island
ggplot(median_detections_per_individual_per_island, aes(x = reorder(myaxis, -median_detections), y = median_detections)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_errorbar(aes(ymin = IQR_lower, ymax = IQR_upper), width = 0.2, color = "red") +
  theme_minimal() +
  labs(title = "Median Detections per Individual with IQR at Each Island",
       x = "Island",
       y = "Median Detections per Individual") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Map detections per transmitter per station
```{r,warning=FALSE}

map_transmitters_per_station <- ggplot() +
   geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the islands
  geom_point(data = transmitters_per_station, aes(x = Longitude, y = Latitude, color = Transmitters, size = Transmitters)) +
  scale_color_viridis_c() +  # Color gradient
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Transmitters per Station", x = "Longitude", y = "Latitude", color = "Transmitters", size = "Transmitters")

map_transmitters_per_station

# Save the plot as an image
ggsave("map_transmitters_per_station.png", plot = map_transmitters_per_station, width = 10, height = 8, dpi = 300, units = "in")

# Add lat/long to average detections data
avg_detections_per_individual_per_station <- detections_per_individual %>%
  group_by(`Station Name`) %>%
  summarise(
    avg_detections = mean(Detections),
    se_detections = sd(Detections) / sqrt(n()),
    num_transmitters = n_distinct(Transmitter)
  ) %>%
  left_join(detections_per_station %>% select(`Station Name`, Latitude, Longitude), by = "Station Name") %>%
  ungroup()

map_avg_detections_per_station <- ggplot() +
   geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the islands
  geom_point(data = avg_detections_per_individual_per_station, aes(x = Longitude, y = Latitude, color = avg_detections, size = avg_detections)) +
  scale_color_viridis_c() +  # Color gradient
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Average Detections per Individual at Each Station", x = "Longitude", y = "Latitude", color = "Average Detections Per Individual", size = "Average Detections Per Individual")

map_avg_detections_per_station

# Save the plot as an image
ggsave("map_avg_detections_per_station.png", plot = map_avg_detections_per_station, width = 10, height = 8, dpi = 300, units = "in")

# Add lat/long to median detections data
median_detections_per_individual_per_station <- detections_per_individual %>%
  group_by(`Station Name`) %>%
  summarise(
    median_detections = median(Detections),
    IQR_lower = quantile(Detections, 0.25),
    IQR_upper = quantile(Detections, 0.75),
    num_transmitters = n_distinct(Transmitter)
  ) %>%
  left_join(detections_per_station %>% select(`Station Name`, Latitude, Longitude), by = "Station Name") %>%
  ungroup()

# Create the map for median detections per transmitter per station
map_median_detections_per_station <- ggplot() +
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the islands
  geom_point(data = median_detections_per_individual_per_station, aes(x = Longitude, y = Latitude, color = median_detections, size = median_detections)) +
  scale_color_viridis_c() +  # Color gradient
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Median Detections per Individual at Each Station", x = "Longitude", y = "Latitude", color = "Median Detections Per Individual", size = "Median Detections Per Individual")

# Display the map
print(map_median_detections_per_station)

# Save the plot as an image
ggsave("map_median_detections_per_station.png", plot = map_median_detections_per_station, width = 10, height = 8, dpi = 300, units = "in")

```

#______________

## Temporal Stats 
```{r, warning=FALSE}


# Calculate detections per year and island
detections_per_year_island <- WCO %>%
  group_by(Year, Island) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Bar plot for Detections per Year colored by Island
ggplot(detections_per_year_island, aes(x = Year, y = Detections, fill = Island)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(title = "Detections per Year", x = "Year", y = "Detections", fill = "Island")

# Calculate detections per month and island (aggregated across all years)
detections_per_month_island <- WCO %>%
  group_by(Month, Island) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Bar plot for Detections per Month colored by Island
ggplot(detections_per_month_island, aes(x = Month, y = Detections, fill = Island)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(title = "Detections per Month", x = "Month", y = "Detections", fill = "Island")

# Calculate detections per season and island
detections_per_season_island <- WCO %>%
  group_by(Season, Island) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Ensure the Season factor is ordered correctly
detections_per_season_island$Season <- factor(detections_per_season_island$Season, levels = c("Spring", "Summer", "Fall", "Winter"))

# Bar plot for Detections per Season colored by Island
ggplot(detections_per_season_island, aes(x = Season, y = Detections, fill = Island)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(title = "Detections per Season", x = "Season", y = "Detections", fill = "Island")

```

## Line Plot for Detections Over Time (2020 and After) Colored by Island
```{r, warning=FALSE}

# Filter for detections from 2020 and after
WCO_2020to2023 <- WCO %>%
  filter(Date >= as.Date("2021-06-01"))

# Calculate 7-day moving average for detections
detections_per_day_island <- WCO_2020to2023 %>%
  group_by(Date, Island) %>%
  summarise(Detections = n()) %>%
  ungroup() %>%
  group_by(Island) %>%
  mutate(MovingAvg = rollmean(Detections, k = 7, fill = NA, align = "right"))

# Line plot with moving average
ggplot(detections_per_day_island, aes(x = Date, y = MovingAvg, color = Island, group = Island)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Detections Over Time with Moving Average (2021 and After)", x = "Date", y = "7-day Moving Average of Detections", color = "Island")

```
## Animated Detections over time
```{r}

# Aggregate detections per station and month
detections_per_station_month <- WCO_2020to2023 %>%
  group_by(Longitude, Latitude, Island, Year, Month) %>%
  summarise(Detections = n()) %>%
  ungroup()

# Extract unique coordinates and station names
unique_coordinates <- WCO_2020to2023 %>%
  select(`Station Name`, Longitude, Latitude) %>%
  distinct()

# Check for stations with multiple coordinates
stations_with_multiple_coords <- unique_coordinates %>%
  group_by(`Station Name`) %>%
  filter(n() > 1) %>%
  arrange(`Station Name`)

# Print the results
print(unique_coordinates)
print(stations_with_multiple_coords)

# Create a Bi-weekly column for easier plotting
detections_per_station_biweekly <- WCO_2020to2023 %>%
  mutate(Biweekly = as.Date(cut(Date, breaks = "15 days"))) %>%
  group_by(Longitude, Latitude, Island, Year = year(Biweekly), Biweekly) %>%
  summarise(Detections = n()) %>%
  ungroup()

base_plot <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_biweekly, aes(x = Longitude, y = Latitude, color = Detections, size = Detections)) +
  scale_color_viridis_c() +  # Color gradient
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Detections", size = "Detections")

# Create the animation
animation_plot <- base_plot +
  geom_point(data = detections_per_station_biweekly, aes(x = Longitude, y = Latitude, color = Detections, size = Detections)) +
  transition_manual(Biweekly) +
  labs(title = "Detections per Station: {current_frame}")

```


### Animate and save
```{r, eval=FALSE}

animate(animation_plot, nframes = length(unique(detections_per_station_biweekly$Biweekly)), fps = 1, width = 800, height = 600, renderer = gifski_renderer("detections_animation_biweekly.gif"))

```
#_______________

# Tables

## Overview All Transmitters
```{r, warning=FALSE}

# Create summary table for with Freedman Parameters
transmitter_summary_freedman <- WCO %>%
  group_by(Transmitter) %>%
  summarise(
    `Date First Detection` = min(Date),
    `Date Last Detection` = max(Date),
    `Detection Span (Days)` = as.numeric(difftime(max(Date), min(Date), units = "days")),
    `Total Days Detected` = n_distinct(Date),
    `Total Detections` = n(),
    `Total Stations Detected` = n_distinct(`Station Name`),
    `Total Islands Detected` = n_distinct(Island)
  ) %>%
  arrange(desc(`Total Stations Detected`))

# Print the summary table
print(transmitter_summary_freedman)

# Export the summary table as a CSV file
write.csv(transmitter_summary_freedman, "transmitter_summary_freedman.csv", row.names = FALSE)

# Conversion factor for degrees of longitude to kilometers at the equator
km_per_degree_longitude <- 111.32

# Create summary table for the  Nosal parameters
transmitter_summary_nosal <- WCO %>%
  group_by(Transmitter) %>%
  summarise(
    `Total Days Detected` = n_distinct(Date),
    `Min Longitude` = min(Longitude),
    `Max Longitude` = max(Longitude)
  ) %>%
  mutate(
    `Max Distance Traveled (km)` = (`Max Longitude` - `Min Longitude`) * km_per_degree_longitude
  ) %>%
  arrange(desc(`Max Distance Traveled (km)`))

# Print the summary table
print(transmitter_summary_nosal)

# Export the summary table as a CSV file
write.csv(transmitter_summary_nosal, "transmitter_summary_nosal.csv", row.names = FALSE)

```

## Species Table
```{r}

# Function to create the summary table for each species
species_summary <- function(species_name) {
  species_filtered <- WCO %>%
    filter(SciName == species_name)
  
  # Calculate the number of days each transmitter was detected
  days_detected_per_transmitter <- species_filtered %>%
    group_by(Transmitter) %>%
    summarise(`Days Detected` = n_distinct(Date))
  
  summary <- species_filtered %>%
    summarise(
      Species = species_name,
      Transmitters = n_distinct(Transmitter),
      Detections = n(),
      `Date First Detected` = min(Date),
      `Date Last Detected` = max(Date),
      `Detection Span (Days)` = as.numeric(difftime(max(Date), min(Date), units = "days")),
      `Average Number of Days Detected per Transmitter` = mean(days_detected_per_transmitter$`Days Detected`),
      `SD Number of Days Detected per Transmitter` = ifelse(n_distinct(Transmitter) > 2, 
                                                            sd(days_detected_per_transmitter$`Days Detected`), NA),
      `Total Stations Detected` = n_distinct(`Station Name`),
      `Total Islands Detected` = n_distinct(Island)
    )
  
  return(summary)
}

# Get the list of unique species names excluding NA values
species_list <- WCO %>%
  filter(!is.na(SciName)) %>%
  distinct(SciName) %>%
  pull(SciName)

# Create the summary table for each species and combine into one dataframe
species_comparison <- lapply(species_list, species_summary) %>%
  bind_rows() %>%
  arrange(desc(Transmitters))

# Print the summary table
print(species_comparison)

# Export the summary table as a CSV file
write.csv(species_comparison, "species_comparison_summary.csv", row.names = FALSE)

```

## Top Species Tables for White Shark, Bat Rays, and Soupfin
```{r}

# Define the species names
Top_Species <- c("Carcharodon carcharias", "Galeorhinus galeus", "Myliobatis californica")

# Conversion factor for degrees of longitude to kilometers at the equator
km_per_degree_longitude <- 111.32

# Function to create summary table for each species
Top_Species_table <- function(species_name) {
  Top_Species_table_filtered <- WCO %>%
    filter(SciName == species_name)
  
  Top_Species_summary <- Top_Species_table_filtered %>%
    group_by(Transmitter) %>%
    summarise(
      `when_encountered` = first(na.omit(when_encountered)),
      `Date First Detection` = min(Date),
      `Date Last Detection` = max(Date),
      `est_tag_expiry_date` = first(na.omit(est_tag_expiry_date)),
      `Detection Span (Days)` = as.numeric(difftime(max(Date), min(Date), units = "days")),
      `Total Days Detected` = n_distinct(Date),
      `Total Detections` = n(),
      `Total Stations Detected` = n_distinct(`Station Name`),
      `Total Islands Detected` = n_distinct(Island),
      `Min Longitude` = min(Longitude),
      `Max Longitude` = max(Longitude),
      `Max Distance Traveled (km)` = (`Max Longitude` - `Min Longitude`) * km_per_degree_longitude,
      `sex` = first(na.omit(sex))
    ) %>%
    arrange(desc(`Total Detections`))
  
  if (species_name %in% c("Carcharodon carcharias", "Galeorhinus galeus")) {
    Top_Species_summary <- Top_Species_summary %>%
      left_join(WCO %>% select(Transmitter, TNL_cm) %>% distinct(), by = "Transmitter")
  } else if (species_name == "Myliobatis californica") {
    Top_Species_summary <- Top_Species_summary %>%
      left_join(WCO %>% select(Transmitter, disk_length) %>% distinct(), by = "Transmitter")
  }

  # Print the summary table
  print(Top_Species_summary)
  
  # Export the summary table as a CSV file
  filename <- paste0("transmitter_summary_", gsub(" ", "_", species_name), ".csv")
  write.csv(Top_Species_summary, filename, row.names = FALSE)
}

# Create summary tables for each species
for (species in Top_Species) {
  Top_Species_table(species)
}

```

#_______________

#Filtered for Species
## Top Species

### Map
```{r}

# Define the species names
Top_Species <- c("Carcharodon carcharias", "Galeorhinus galeus", "Myliobatis californica")

# Function to create map for each species
create_species_map <- function(species_name) {
  # Filter the data for the species
  species_data <- WCO %>%
    filter(SciName == species_name)
  
  # Calculate the summaries for the species
  detections_per_individual <- species_data %>%
    group_by(`Station Name`, Transmitter) %>%
    summarise(Detections = n()) %>%
    ungroup()
  
  avg_detections_per_individual_per_station <- detections_per_individual %>%
    group_by(`Station Name`) %>%
    summarise(
      avg_detections = mean(Detections),
      num_transmitters = n_distinct(Transmitter)
    ) %>%
    left_join(detections_per_station %>% select(`Station Name`, Latitude, Longitude), by = "Station Name") %>%
    ungroup()
  
  # Create the map
  map <- ggplot() +
    geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
    geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the islands
    geom_point(data = avg_detections_per_individual_per_station, 
               aes(x = Longitude, y = Latitude, color = num_transmitters, size = avg_detections)) +
    scale_color_viridis_c() +  # Color gradient
    coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
    theme_minimal() +
    theme(
      panel.background = element_rect(fill = "white", color = NA),  # White background
      plot.background = element_rect(fill = "white", color = NA),   # White plot background
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(title = paste("Detections per Individual for", species_name, "at Each Station"),
         x = "Longitude", y = "Latitude", color = "Number of Individuals", size = "Average Detections Per Individual")
  
  # Save the plot as an image
  filename <- paste0("map_avg_detections_per_station_", gsub(" ", "_", species_name), ".png")
  ggsave(filename, plot = map, width = 10, height = 8, dpi = 300, units = "in")
  
  return(map)
}

# Create and save maps for each top species
for (species in Top_Species) {
  create_species_map(species)
}

```

### Animated Map
```{r, eval=FALSE}

# Function to create animated map for a species
create_species_animation <- function(species_name) {
  # Filter the data for the species
  species_data <- WCO %>%
    filter(SciName == species_name)
  
  # Aggregate detections per station and biweekly period
  detections_per_station_biweekly <- species_data %>%
    mutate(Biweekly = as.Date(cut(Date, breaks = "15 days"))) %>%
    group_by(Longitude, Latitude, Island, Year = year(Biweekly), Biweekly) %>%
    summarise(Detections = n()) %>%
    ungroup()
  
  # Create the base plot
  base_plot <- ggplot() + 
    geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
    geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
    geom_point(data = detections_per_station_biweekly, aes(x = Longitude, y = Latitude, color = Detections, size = Detections)) +
    scale_color_viridis_c() +  # Color gradient
    coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
    theme_minimal() +
    theme(
      panel.background = element_rect(fill = "white", color = NA),  # White background
      plot.background = element_rect(fill = "white", color = NA),   # White plot background
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(x = "Longitude", y = "Latitude", color = "Detections", size = "Detections")
  
  # Create the animation
  animation_plot <- base_plot +
    geom_point(data = detections_per_station_biweekly, aes(x = Longitude, y = Latitude, color = Detections, size = Detections)) +
    transition_manual(Biweekly) +
    labs(title = paste("Detections per Station for", species_name, ": {current_frame}"))

  # Save the animation
  filename <- paste0("detections_animation_biweekly_", gsub(" ", "_", species_name), ".gif")
  animate(animation_plot, nframes = length(unique(detections_per_station_biweekly$Biweekly)), fps = 1, width = 800, height = 600, renderer = gifski_renderer(filename))
}

# Define the species names
Top_Species <- c("Carcharodon carcharias", "Galeorhinus galeus", "Myliobatis californica")

# Create animations for each top species
for (species in Top_Species) {
  create_species_animation(species)
}

```

### Detections Over time
```{r}

# Function to create a line plot with moving average for a species
create_species_plot <- function(species_name) {
  # Filter the data for the species
  species_data <- WCO %>%
    filter(SciName == species_name)
  
  # Calculate 7-day moving average for detections
  detections_per_day_island <- species_data %>%
    group_by(Date, Island) %>%
    summarise(Detections = n(), .groups = "drop") %>%
    group_by(Island) %>%
    mutate(MovingAvg = rollmean(Detections, k = 7, fill = NA, align = "right"))
  
  # Line plot with moving average
  plot <- ggplot(detections_per_day_island, aes(x = Date, y = MovingAvg, color = Island, group = Island)) +
    geom_line() +
    theme_minimal(base_size = 15) +  # Increase base font size for better readability
    theme(
      panel.background = element_rect(fill = "white", color = NA),  # White background
      plot.background = element_rect(fill = "white", color = NA),   # White plot background
      panel.grid.major = element_line(color = "grey80"),            # Light grey grid lines
      panel.grid.minor = element_line(color = "grey90")             # Even lighter grid lines
    ) +
    labs(title = paste("Detections Over Time with Moving Average for", species_name), x = "Date", y = "7-day Moving Average of Detections", color = "Island")
  
  # Save the plot
  filename <- paste0("detections_moving_avg_", gsub(" ", "_", species_name), ".png")
  ggsave(filename, plot = plot, width = 10, height = 6, dpi = 300)
  
  return(plot)
}

# Create and save plots for each top species
for (species in Top_Species) {
  create_species_plot(species)
}

```

#_______________

# Filtered For Transmitters 
## Abacus Plots for top 10 Detections and Distance
```{r, warning=FALSE}

# Combined list top 10 transmitters
Top10 <- c("A69-9001-22774", "A69-9001-24655", "A69-1601-27075", "A69-9001-12159", "A69-1303-46140", "A69-1601-62047", "A69-1601-62096", "A69-1602-3228", "A69-1602-8666", "A69-1602-58589")

# List of specified transmitters
Top10_1 <- c("A69-9001-22774", "A69-9001-24655", "A69-1601-27075", "A69-9001-12159", "A69-1303-46140")
Top10_2 <- c("A69-1601-62047", "A69-1601-62096", "A69-1602-3228", "A69-1602-8666", "A69-1602-58589")

# Define the order of islands and stations in descending order
island_order <- c("San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Mainland")
station_order <- c("WCO_Cuyler", "WCO_Tyler Bight", "WCO_Carrington Point", "WCO_Johnson's Lee",
                   "WCO_Hazards_Profile", "WCO_Trailers", "WCO_MorsePoint", "WCO_Valley",
                   "WCO_Smugglers", "WCO_Scorpion", "WCO_Pelican", "WCO_Winfield Scott",
                   "WCO_Anacapa South", "WCO_Mohawk LTER", "WCO_Carpinteria")

# Create a list of shapes to use for each transmitter
shapes <- c(15, 16, 17, 18, 19, 20, 21, 22, 23, 24)

# Custom colors for the transmitters
custom_colors <- c("#C58ABB", "#B73944", "#8ACE89", "#C15901", "#0072B2", 
                   "#656565", "#009E73", "#56B4E9", "#E69F00", "#000000")

# Function to process the dataset
process_data <- function(data, transmitters) {
  data %>%
    filter(Transmitter %in% transmitters) %>%
    mutate(Island = factor(Island, levels = island_order),
           `Station Name` = factor(`Station Name`, levels = station_order))
}

# Process the datasets
WCO_Top10_1 <- process_data(WCO, Top10_1)
WCO_Top10_2 <- process_data(WCO, Top10_2)

# Create the abacus plot for Station
ggplot(WCO_Top10_1, aes(x = `Date`, y = `Station Name`, color = Transmitter, shape = Transmitter)) +
  geom_jitter(size = 3, width = 0.4, height = 0.35) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  theme_minimal() +
  labs(title = NULL,
       x = "Year",
       y = "Station",
       color = "Transmitter",
       shape = "Transmitter") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create the abacus plot for Islands
ggplot(WCO_Top10_1, aes(x = `Date`, y = Island, color = Transmitter, shape = Transmitter)) +
  geom_jitter(size = 3, width = 0.4, height = 0.35) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  theme_minimal() +
  labs(title = NULL,
       x = "Year",
       y = "Island",
       color = "Transmitter",
       shape = "Transmitter") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create the abacus plot for Station
ggplot(WCO_Top10_2, aes(x = `Date`, y = `Station Name`, color = Transmitter, shape = Transmitter)) +
  geom_jitter(size = 3, width = 0.4, height = 0.35) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  theme_minimal() +
  labs(title = NULL,
       x = "Year",
       y = "Station",
       color = "Transmitter",
       shape = "Transmitter") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create the abacus plot for Islands
ggplot(WCO_Top10_2, aes(x = `Date`, y = Island, color = Transmitter, shape = Transmitter)) +
  geom_jitter(size = 3, width = 0.4, height = 0.35) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  theme_minimal() +
  labs(title = NULL,
       x = "Year",
       y = "Island",
       color = "Transmitter",
       shape = "Transmitter") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create the faceted abacus plot for individual transmitters
ggplot(WCO_Top10_1, aes(x = `Date`, y = `Island`)) +
  geom_point(color = "black") +
  theme_minimal() +
  labs(title = NULL,
       x = "Year",
       y = "Island") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Transmitter, scales = "free_y", ncol = 1)

# Create the faceted abacus plot for individual transmitters
ggplot(WCO_Top10_2, aes(x = `Date`, y = `Island`)) +
  geom_point(color = "black") +
  theme_minimal() +
  labs(title = NULL,
       x = "Year",
       y = "Island") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Transmitter, scales = "free_y", ncol = 1)

```

## Map for top 10 Detections and Distance
```{r, warning=FALSE}

# Filter the dataset for the top 10 transmitters
WCO_Top10 <- WCO %>%
  filter(Transmitter %in% Top10)

# Summarize detections per station for the specified transmitters
detections_per_station_Top10 <- WCO_Top10 %>%
  group_by(`Station Name`, Latitude, Longitude) %>%
  summarise(Detections = n()) %>%  # Using n() to count the number of detections
  ungroup()

# Create the scatter plot for detections per station with shapefile
map_Top10_detections <- ggplot() +
  geom_sf(data = coast, color= "black") +
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_Top10, aes(x = Longitude, y = Latitude, color = Detections, size = Detections)) +
  scale_color_viridis_c() +  # Color gradient
  scale_size_continuous(range = c(3, 10)) +  # Scale point sizes up
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Detections per Station", x = "Longitude", y = "Latitude", color = "Detections", size = "Detections")

map_Top10_detections

# Save the plot as an image
ggsave("detections_per_station_Top10.png", plot = map_Top10_detections, width = 10, height = 8, dpi = 300, units = "in")

```

## Johnson's Lee April-July 2023
```{r, warning=FALSE}

# Filter the dataset for the specific station and date range
transmitters_johnsons_lee <- WCO %>%
  filter(`Station Name` == "WCO_Johnson's Lee",
         Date >= as.Date("2023-04-01") & Date <= as.Date("2023-07-31")) %>%
  select(Transmitter) %>%
  distinct()

# Print the unique transmitters
print(transmitters_johnsons_lee)

# Abacus for Johnson's Lee Transmitters

# List of specified transmitters
Johnsons_Transmitters <- c("A69-1601-62047", "A69-1601-62096", "A69-9001-55586", "A69-9001-2552")

# Filter the dataset for the specified transmitters
Filtered_Johnsons_Transmitters <- WCO %>%
  filter(Transmitter %in% Johnsons_Transmitters)

# Create the faceted abacus plot for individual transmitters
ggplot(Filtered_Johnsons_Transmitters, aes(x = Date, y = Island)) +
  geom_point(color = "black") +
  theme_minimal() +
  labs(title = NULL,
       x = "Year",
       y = "Island") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Transmitter, scales = "free_y", ncol = 1)

```

# Make Abacus for Myliobatis californica
```{r, warning=FALSE}

str(WCO)

# Filter the dataset for the specific species
Filtered_Myliobatis <- WCO %>%
  filter(SciName == "Myliobatis californica")

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.
mylobatis_abacus <- ggplot(Filtered_Myliobatis, aes(x = Island, y = Date, color = Transmitter, shape = Transmitter)) +  # Specify the x and y aesthetics      
  geom_jitter(size = 3, width = 0.4, height = 0.35) +  # Add jittered points with specified size and width/height
  scale_color_manual(values = custom_colors) +  # Set custom
  scale_shape_manual(values = shapes) +  # Set custom shapes
  theme_minimal() +  # Apply a minimal theme
  labs(title = NULL, x = "Island", y = "Date", color = "Transmitter", shape = "Transmitter") +  # Set the axis labels and title
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

  print(mylobatis_abacus)

  # export the plot as an image with white background
ggsave("Abacus_Plot_Myliobatis.png", plot = mylobatis_abacus, width = 10, height = 8, dpi = 300, units = "in")
```