---
title: "Kickback Tag Data"
author: "Morgan Youngblood"
date: "06/9/2024"
---
 

# Load data
```{r load_data, echo=FALSE}

# Load libraries
library(dplyr)
library(readr) 
library(readxl)

# Load reciever data
reciever_data <- read_csv("data/og_detection-data_cinms-wco_20231024_01.csv")

# Load tag data
LoweLab <- readRDS("data/LoweLabTagInfo (1).rds")

# View structure of each dataset to determine how to merge
str(reciever_data)
str(LoweLab)

# which columns have NA values in the LoweLab dataset
colSums(is.na(LoweLab))

```

# Merge Data
```{r merge_Lowe_data, echo=FALSE}

# Merge data from both datasets by the "Transmitter" column
Lowe_Detection_data <- reciever_data %>%
  left_join(LoweLab, by = c ("Transmitter" = "Transmitter"))

# View structure of merged data
str(Lowe_Detection_data)

# Filter data for matching transmitters
Lowe_Detection_data <- Lowe_Detection_data %>%
    filter(!is.na(Owner))

# Write filtered data to a .csv file
write_csv(Lowe_Detection_data, "data/Lowe_Detection_20231024.csv")

# Write filtered data to a .rds file
saveRDS(Lowe_Detection_data, "data/Lowe_Detection_20231024.rds")

# View the first few rows of the merged data
head(Lowe_Detection_data)

```

# Load and Merge Data
```{r merge_Nosal_data, echo=FALSE}

# Load Nosal data
Nosal <- readxl::read_xlsx("data/Nosal_Soupfin_Tag_IDs_2024.xlsx")

# View structure of each dataset to determine how to merge
str(Nosal)
str(reciever_data)

# Which columns have NA values in the Nosal dataset
colSums(is.na(Nosal))

 # Merge "Tag_ID_Combined" from Nosal data with "Transmitter" from reciever data
Nosal_Detection_data <- reciever_data %>%
  left_join(Nosal, by = c("Transmitter" = "Tag_ID_Combined"))

# View structure of merged data
str(Nosal_Detection_data)

# Filter data for matching transmitters
Nosal_Detection_data <- Nosal_Detection_data %>%
    filter(!is.na(Species))

# Write filtered data to a .csv file
write_csv(Nosal_Detection_data, "data/Nosal_Detection_20231024.csv")

# Look at the number of unique Transmitters in the Nosal_Detection_data
length(unique(Nosal_Detection_data$Transmitter))

```

# Load and Merge Data
```{r update_Block_data, echo=FALSE}

# Load Block data
Block <- read.csv("data/block_tags_2024june7.csv")

# View structure of each dataset to determine how to merge
str(Block)
str(reciever_data)

# Which columns have NA values in the Block dataset
colSums(is.na(Block))

# Merge "codespace" from Block data with "Transmitter" from reciever data
Block_Detection_data <- reciever_data %>%
  left_join(Block, by = c("Transmitter" = "codespace"))

# View structure of merged data
str(Block_Detection_data)

# Filter data for matching transmitters
Block_Detection_data <- Block_Detection_data %>%
    filter(!is.na(commonname))

# Write filtered data to a .csv file
write_csv(Block_Detection_data, "data/Block_Detection_20231024.csv")

# Look at how many unique Transmitters are in the Block_Detection_data
length(unique(Block_Detection_data$Transmitter))

```

# Merge Block Data into ongoing Tag Data
```{r update_unknown_data, echo=FALSE}

Unknown_data <- read_csv("data/Unknown_Tag_Data.csv")

# View structure of sorted data and Block data to determine how to merge
str(Unknown_data)
str(Block)

# How many unique "codespace" values are in the Block data
length(unique(Block$codespace))

# See how many "codespace" values in the Block data match the "Transmitter" values in the Unknown data
sum(Unknown_data$Transmitter %in% Block$codespace)

# Merge Block data with sorted_Tag_Data.csv by the "Transmitter" column
merged_Block_data <- Unknown_data %>%
    left_join(Block, by = c("Transmitter" = "codespace"))

# View structure of merged data
str(merged_Block_data)

# Change the column names of "merged_Block_data" to match the column names of "sorted_Tag_Data.csv"
colnames(merged_Block_data) <- c("Transmitter", "number_of_detections", "first_detection", "last_detection", "SciName", "when_encountered", "seriesname", "TNL_cm", "eventid")

# Check colnames of "merged_Block_data"
colnames(merged_Block_data)

# Change all SciName = "White Shark" to "Carcharodon carcharias" in the merged_Block_data
merged_Block_data$SciName[merged_Block_data$SciName == "White Shark"] <- "Carcharodon carcharias"

# Filter data for matching transmitters
merged_Block_data <- merged_Block_data %>%
    filter(!is.na(SciName))

    # Remove eventid columns from merged_Block_data
merged_Block_data <- merged_Block_data %>%
    select(-eventid)

# Load sorted_Tag_Data.csv
sorted_data <- read_csv("data/sorted_Tag_Data.csv")

#view structure of sorted data
str(sorted_data)

# Merge the datasets on the Transmitter column
Tag_ID_6_9 <- sorted_data %>%
  left_join(merged_Block_data, by = "Transmitter")

# Merge seriesname column from Block data with tag_location column from sorted data as a single tag_location column
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(tag_location = if_else(is.na(tag_location), tag_location, seriesname)) %>%
  select(-seriesname)

# Combine SciName.x and SciName.y into a single SciName column
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(SciName = if_else(is.na(SciName.y), SciName.x, SciName.y)) %>%
  select(-SciName.x, -SciName.y)

  # Do the same for "TNL_cm", "when_encountered", "last_detection", 'first_detection', and "number_of_detections" columns
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(TNL_cm = if_else(is.na(TNL_cm.y), TNL_cm.x, TNL_cm.y)) %>%
  select(-TNL_cm.x, -TNL_cm.y)

# Change when_encountered column to Date format
Tag_ID_6_9$when_encountered.y <- as.Date(Tag_ID_6_9$when_encountered.y, format = "%m-%d-%y")
# Convert the Date format to year-month-day
Tag_ID_6_9$when_encountered.y <- format(Tag_ID_6_9$when_encountered.y, "%Y-%m-%d")

Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(when_encountered = if_else(is.na(when_encountered.x), as.character(when_encountered.x), as.character(when_encountered.y))) %>%
  select(-when_encountered.x, -when_encountered.y)

Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(last_detection = if_else(is.na(last_detection.y), last_detection.x, last_detection.y)) %>%
  select(-last_detection.x, -last_detection.y)

Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(first_detection = if_else(is.na(first_detection.y), first_detection.x, first_detection.y)) %>%
  select(-first_detection.x, -first_detection.y)

Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(number_of_detections = if_else(is.na(number_of_detections.y), number_of_detections.x, number_of_detections.y)) %>%
  select(-number_of_detections.x, -number_of_detections.y)

# View the merged data
str(Tag_ID_6_9)
head(Tag_ID_6_9)

# Export the merged dataset to a CSV file
write_csv(Tag_ID_6_9, "outputs/Updated_Tag_ID_20240609.csv")

# Filter the dataset for rows where SciName is NA
NA_Tag_ID_6_9 <- Tag_ID_6_9 %>%
  filter(is.na(SciName)) %>%
  select(Transmitter, first_detection, last_detection)

# Export the filtered dataset to a .csv file
write_csv(NA_Tag_ID_6_9, "outputs/NA_Tag_ID_20240609.csv")

```

# Merge Nosal Data into ongoing Tag Data
```{r update_Nosal_data, echo=FALSE}

# View structure of Nosal data and Tag_ID_6_9 to determine how to merge
str(Nosal)
str(Tag_ID_6_9)

# Filter unnecessary columns from the Nosal data
Nosal <- Nosal %>%
  select(-V16_ID, -Code_Space)

# Merge the datasets on the Transmitter column
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  left_join(Nosal, by = c("Transmitter" = "Tag_ID_Combined"))

# View structure of merged data
str(Tag_ID_6_9)

# Use mutate to combine the columns from the two datasets
## mutate SciName
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(Species = if_else(is.na(SciName), Species, SciName)) %>%
  select(-Species)

## mutate TNL_cm
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(TNL_cm = if_else(is.na(TNL_cm), TL, TNL_cm)) %>%
  select(-TL)

## mutate sex
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(sex = if_else(is.na(sex), Sex, sex)) %>%
  select(-Sex)

## mutate when_encountered
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(when_encountered = if_else(is.na(when_encountered), Tagging_Date, when_encountered)) %>%
  select(-Tagging_Date)

## mutate tag_location
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(tag_location = if_else(is.na(tag_location), Tagging_Region, tag_location)) %>%
  select(-Tagging_Region)

## mutate Transmitter_SN
Tag_ID_6_9 <- Tag_ID_6_9 %>%
  mutate(Transmitter_SN = if_else(is.na(Transmitter_SN), V16_SN, Transmitter_SN)) %>%
  select(-V16_SN)

# View the merged data
str(Tag_ID_6_9)

# Export the merged dataset to a CSV file
write_csv(Tag_ID_6_9, "outputs/Updated_Tag_ID_20240611.csv")

# Filter the dataset for rows where SciName is NA
NA_Tag_ID_6_11 <- Tag_ID_6_9 %>%
  filter(is.na(when_encountered)) %>%
  select(Transmitter, SciName, first_detection, last_detection)

# Export the filtered dataset to a .csv file
write_csv(NA_Tag_ID_6_11, "outputs/NA_Tag_ID_20240611.csv")

# Look at how many FL are NA in the Tag_ID_6_9 dataset to see how many Nosal Tag data were added
sum(!is.na(Tag_ID_6_9$FL))

# Look at Transmitters where SciName are Galeorhinus galeus in the Tag_ID_6_9 dataset and FL is NA
Unknown_Galeorhinus <- Tag_ID_6_9 %>%
  filter(SciName == "Galeorhinus galeus" & is.na(FL))

# Count the number of unique transmitters in the Unknown_Galeorhinus dataset
length(unique(Unknown_Galeorhinus$Transmitter))

# Write .csv file for Unknown_Galeorhinus with columns Transmitter, number_of_detections, first_detection, and last_detection
write_csv(Unknown_Galeorhinus, "outputs/Unknown_Galeorhinus_20240611.csv")
```

# Create Filtered Detections Data file for Myliobatis californica for Alyssa
```{r create_Myliobatis_data, echo=FALSE}

Total_detections <- read.csv("data/og_detection-data_cinms-wco_20231024_01.csv")

str(Total_detections)

# Filter data for transmitter A69-1602-12253
Detections_12253 <- Total_detections %>%
  filter(Transmitter == "A69-1602-12253")

view structure of Myliobatis data
str(Detections_12253)

# Write filtered data to a .csv file
write_csv(Detections_12253, "data/Detections_12253_20240611.csv")
```

# Filter detection data for OSU Sevengill Transmitter # A69-9001-2552
```{r filter_2552_data, echo=FALSE}

# Filter data for Transmitter A69-9001-2552
A69_9001_2552_data <- Total_detections %>%
  filter(Transmitter == "A69-9001-2552")

# Write filtered data to a .csv file
write_csv(A69_9001_2552_data, "outputs/A69_9001_2552_detections_20240611.csv")

```

# Incorporate Lowe Update into ongoing Tag Data
```{r update_Lowe_data, echo=FALSE}

# Read in updated data
tag_id_up <- read.csv("data/Tag_ID_20240611.csv")


# Read in Lowe Lab update xlxs file
Lowe_update <- readxl::read_xlsx("data/DDT+taggingJune2024-Lowe.xlsx")

# View structure of each dataset to determine how to merge
str(tag_id_up)
str(Lowe_update)

# Extract the last 5 digits from the "Transmitter" column
tag_id_up$last_5_digits <- substr(tag_id_up$Transmitter, nchar(tag_id_up$Transmitter) - 4, nchar(tag_id_up$Transmitter))

# Change "internal tag ID" column to character in the Lowe data
Lowe_update$`internal tag ID` <- as.character(Lowe_update$`internal tag ID`)

# make sure the last_5_digits column is character
tag_id_up$last_5_digits <- as.character(tag_id_up$last_5_digits)

# Merge based on the "last_5_digits" column in tag_id_up and the "internal tag ID" column in Lowe_update
tag_id_up <- tag_id_up %>%
  left_join(Lowe_update, by = c("last_5_digits" = "internal tag ID"))

# Remove the "last_5_digits" column
tag_id_up <- tag_id_up %>%
  select(-last_5_digits)

# View structure of merged data
str(tag_id_up)

# see how many transmitters merged by looking at how many "Internal tag type" values are not NA
sum(!is.na(tag_id_up$`Internal tag type`))

# Make list of which transmitters were updated by looking at the "Internal tag type" values that are not NA and seeing which transmitters are associated with them
new_transmitters <- unique(tag_id_up$Transmitter[!is.na(tag_id_up$`Internal tag type`)])

new_transmitters

```

# Send Noah Ben-Aderet's yellowtail data
```{r send_yellowtail_data, echo=FALSE}

# Load OG detection data
OG_detection_data <- read.csv("data/og_detection-data_cinms-wco_20231024_01.csv")

# Filter for Transmitter A69-1601-27071
A69_1601_27071 <- OG_detection_data %>%
  filter(Transmitter == "A69-1601-27071")

# view structure of A69_1601_27071
str(A69_1601_27071)

# Write filtered data to a .csv file
write_csv(A69_1601_27071, "outputs/A69_1601_27071_detections_20240611.csv")
```

# Send Alyssa data for transmitter A69-1602-12244
```{r send_12244_data, echo=FALSE}

# Load new detection data
new_detection_data <- read.csv("data/hollings_detection-data_cinms_wco_20240617.csv")

# Filter for Transmitter A69-1602-12244
A69_1602_12244 <- new_detection_data %>%
  filter(Transmitter == "A69-1602-12244")

# view structure of A69_1602_12244
str(A69_1602_12244)

# Write filtered data to a .csv file
write_csv(A69_1602_12244, "outputs/A69_1602_12244_detections_20240617.csv")
```

# Send Sal Jorgensen's data 
```{r send_Sal_data, echo=FALSE}

# Load updated detection data
updated_detection_data <- read.csv("data/wco_detections+tags_20240618.csv")

# Filter for Owner Salvador Jorgensen
Salvador_Jorgensen <- updated_detection_data %>%
  filter(Owner == "Salvador Jorgensen")

# view structure of Salvador_Jorgensen
str(Salvador_Jorgensen)

# Clean Data
Salvador_Jorgensen <- Salvador_Jorgensen %>%
  dplyr::select(Date.and.Time..UTC., Receiver, Transmitter, Transmitter.Name, Transmitter.Serial, Sensor.Value, Sensor.Unit, Station.Name, Latitude, Longitude, Local_Time)

# Identify unique transmitters
unique_transmitters <- unique(Salvador_Jorgensen$Transmitter)

unique_transmitters

# Write filtered data to a .csv file
write_csv(Salvador_Jorgensen, "outputs/Salvador_Jorgensen_CINMS_detections_20240618.csv")

```

# Send Noah Ben-Aderet's data
```{r send_Noah_data, echo=FALSE}

# Load updated detection data
updated_detection_data <- read.csv("data/wco_detections+tags_20240618.csv")

# Filter for Owner Noah Ben-Aderet
Noah_Ben_Aderet <- updated_detection_data %>%
  filter(Owner == "Noah Ben-Aderet")

# view structure of Noah_Ben_Aderet
str(Noah_Ben_Aderet)

# Clean Data
Noah_Ben_Aderet <- Noah_Ben_Aderet %>%
  dplyr::select(Date.and.Time..UTC., Receiver, Transmitter, Transmitter.Name, Transmitter.Serial, Sensor.Value, Sensor.Unit, Station.Name, Latitude, Longitude, Local_Time)

# Identify unique transmitters
unique_transmitters <- unique(Noah_Ben_Aderet$Transmitter)

unique_transmitters

# Write filtered data to a .csv file
write_csv(Noah_Ben_Aderet, "outputs/Noah_Ben_Aderet_CINMS_detections_20240618.csv")

```

# Send Andy Nosal's data
```{r send_Andy_data, echo=FALSE}

# Load updated detection data
updated_detection_data <- read.csv("data/wco_detections+tags_20240618.csv")

# Filter for Owner Andy Nosal
Andy_Nosal <- updated_detection_data %>%
  filter(Owner == "Andrew Nosal")

# view structure of Andy_Nosal
str(Andy_Nosal)

# Clean Data
Andy_Nosal <- Andy_Nosal %>%
  dplyr::select(Date.and.Time..UTC., Receiver, Transmitter, Transmitter.Name, Transmitter.Serial, Sensor.Value, Sensor.Unit, Station.Name, Latitude, Longitude, Local_Time)

# Identify unique transmitters
unique_transmitters <- unique(Andy_Nosal$Transmitter)

unique_transmitters

# Write filtered data to a .csv file
write_csv(Andy_Nosal, "outputs/Andy_Nosal_CINMS_detections_20240618.csv")

```

# Send Block data for tag ID A69-9001-64475 and A69-1303-33545
```{r send_Block_data, echo=FALSE}

# Load updated detection data
updated_detection_data <- read.csv("data/wco_detections+tags_20240618.csv")

# Filter for Transmitter A69-9001-64475 and A69-1303-33545
Mike_detection_data_64475_33545 <- updated_detection_data %>%
  filter(Transmitter %in% c("A69-9001-64475", "A69-1303-33545"))

  # Clean Data
Mike_detection_data_64475_33545 <- Mike_detection_data_64475_33545 %>%
  dplyr::select(Date.and.Time..UTC., Receiver, Transmitter, Transmitter.Name, Transmitter.Serial, Sensor.Value, Sensor.Unit, Station.Name, Latitude, Longitude, Local_Time)


# View Mike_detection_data_64475_33545
View(Mike_detection_data_64475_33545)

# Write filtered data to a .csv file
write_csv(Mike_detection_data_64475_33545, "outputs/Mike_detection_data_64475_33545_20240618.csv")
```