---
title: "Manuscript Analysis"
author: "Morgan Youngblood"
date: "`r Sys.Date()`"
output: html_document
---

# Notes

- Cleaned data is located at: `data/wco_detections+tags_20241226.csv`
- Updated tag ID data is located at: `data/updated_tag_id_20240621.csv`

# Load Libraries
```{r}
# install.packages("sf")
# install.packages("geosphere")
library(dplyr)
library(readr)
library(tidyr)
library(sf)
library(ggplot2)
library(ggthemes)
library(geosphere)
```

# Read in the WCO data
```{r}
tag_data <- read_csv("data/updated_tag_id_20241226.csv")
str(tag_data)

detection_data <- read_csv("data/wco_detections+tags_20241226.csv")
str(detection_data)
problems(detection_data)
head(detection_data)
```

# Let's look at some summary statistics for the detection data. I want to know the time span of the data and total number of detections. 
```{r}
# Calculate the time span of the data
time_span <- range(detection_data$`Date and Time (UTC)`, na.rm = TRUE)
time_span

# Calculate the total number of detections
total_detections <- nrow(detection_data)
total_detections

# How many unique transmitters are there?
unique_transmitters <- n_distinct(detection_data$Transmitter)
unique_transmitters

# How many unique transmittrs have an Owner listed?
unique_transmitters_with_owner <- detection_data %>%
  filter(!is.na(Owner)) %>%
  summarise(n_distinct(Transmitter)) %>%
  pull()
unique_transmitters_with_owner
```

# Let's make sure there aren't any duplicate rows in the detection data
```{r}
# Check for duplicate rows
detection_data %>%
    distinct() %>%
    nrow()

# How many total rows are there?
nrow(detection_data)
```


# Figure 1.1: Map of station locations
```{r}
# make a table of unique station locations with Station Name, Latitude, and Longitude columns
station_locations <- detection_data %>%
  select(`Station Name`, Latitude, Longitude, Island) %>%
  distinct()

# Remove stations with Island = Santa Barbara
station_locations <- station_locations %>%
  filter(Island != "Santa Barbara")

View(station_locations)

# Make a locator map of site_locations

# name shape files for the map
sanctuary <- st_read("data/cinms_py2/cinms_py.shp")
coast <- st_read("data/stanford-vx275xn8886-shapefile/vx275xn8886.shp")
islands <- st_read("data/California_Counties/California_Counties.shp")
baja <- st_read("data/nyu_2451_36780/bcs_entidad.shp")
bcali <- st_read("data/nyu_2451_36772/bc_municipio.shp")

# Make a map of the station locations for reciever sites
wco_locator <- ggplot() +
    geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
    geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
    geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
    geom_sf(data = baja, fill = "#B2B2B2", color = "grey40") +
    geom_sf(data = bcali, fill = "#B2B2B2", color = "grey40") +
    coord_sf(xlim = c(-123.025, -114.45), ylim = c(38.3, 27.85)) +
    theme_map() +
    theme(
        legend.position = "bottomleft",
        legend.title = element_text(size = 25),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
        panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 20),
        legend.text = element_text(size = 20)
    ) +
    labs(title = NULL, x = "Longitude", y = "Latitude") +
    scale_x_continuous(breaks = seq(-120, -114, by = 3)) +
    scale_y_continuous(breaks = seq(27, 39, by = 3)) +
    geom_rect(aes(xmin = -120.6, xmax = -119.3, ymin = 33.75, ymax = 34.5), fill = NA, color = "black", size = 2)


print(wco_locator)

# export the map
ggsave("outputs/wco_locator.png", width = 15, height = 20, units = "in", dpi = 300)

# make a map of the site locations

# What is the range of lat and lon?
lat_range <- range(station_locations$Latitude, na.rm = TRUE)
lon_range <- range(station_locations$Longitude, na.rm = TRUE)

lat_range
lon_range

# Adjust the ggplot object 
station_locations_map <- ggplot() +
    geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
    geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
    geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
    geom_point(data = station_locations, aes(x = Longitude, y = Latitude), color = "black", size = 10) + # Increased size
    coord_sf(xlim = c(-120.6, -119.3), ylim = c(33.75, 34.5)) +
    theme_map() +
    theme(
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold"), # Increased size and bold
        panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 30, face = "bold"), # Increased size and bold
        legend.text = element_text(size = 24)
    ) +
    scale_y_continuous(breaks = c(33.8, 34.2, 34.6)) + # Custom y-axis breaks
    scale_x_continuous(breaks = c(-120.4, -120.0, -119.6)) +
    labs(x = "Longitude", y = "Latitude")

print(station_locations_map)

# Export the map
ggsave("outputs/station_locations_map.png", width = 20, height = 15, units = "in", dpi = 300)


```

# Figure 2.1: Map with total detections per station
## I took this from previous work and modified it to fit the Manuscript
```{r}
# Calculate total detections per station
total_detections_per_station <- detection_data %>%
    group_by(`Station Name`, Latitude, Longitude) %>%
    summarise(Total_Detections = n(), .groups = "drop")

# Adjust the ggplot object
detections_map <- ggplot() +
    geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
    geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
    geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
    geom_point(data = total_detections_per_station, aes(x = Longitude, y = Latitude, size = Total_Detections, color = Total_Detections)) +
    scale_color_gradient(low = "blue", high = "red") +
    scale_size_continuous(range = c(3, 10)) +  # Adjust the range to make points bigger
    coord_sf(xlim = c(-120.6, -119.3), ylim = c(33.75, 34.5)) +
    theme_map() +
    theme(
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold"),
        panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 30, face = "bold"),
        legend.text = element_text(size = 24),
        legend.title = element_text(size = 30, face = "bold")  # Make legend title bold and larger
    ) +
    scale_y_continuous(breaks = c(33.8, 34.2, 34.6)) +
    scale_x_continuous(breaks = c(-120.4, -120.0, -119.6)) +
    labs(x = "Longitude", y = "Latitude", color = "Detections") +
    guides(size = "none") +
    geom_vline(xintercept = -119.8, linetype = "dashed", color = "black", size = 1)

print(detections_map)

# Export the map
ggsave("outputs/fig_2.1_detections_map.png", width = 20, height = 15, units = "in", dpi = 300)
```

## Let's look at some stats associated with figure 2.2. I want to know the total number of detections at each island, which station has the most detections, and if there were more detections at eastern or western stations.
```{r}
# Filter out detections associated with Island = Santa Barbara
detection_data <- detection_data %>%
    filter(Island != "Santa Barbara")

# Count the total number of detections at each island
detections_per_island <- detection_data %>%
    group_by(Island) %>%
    summarise(Total_Detections = n())

    # count the percentage of detections at each island
detections_per_island <- detection_data %>%
    group_by(Island) %>%
    summarise(Total_Detections = n()) %>%
    mutate(Percentage = Total_Detections / sum(Total_Detections) * 100)


# List the stations in order of total detections
stations_by_detections <- detection_data %>%
    group_by(`Station Name`) %>%
    summarise(Total_Detections = n()) %>%
    arrange(desc(Total_Detections))

# Print the list of stations
stations_by_detections

# Count the number of detections at eastern and western stations
east_west_detections <- detection_data %>%
    mutate(East_West = ifelse(Longitude < -119.8, "West", "East")) %>%
    group_by(East_West) %>%
    summarise(Total_Detections = n())

# Print the results
detections_per_island
most_detections_station
east_west_detections
```

# Figure 2.2: Detections per species
```{r}

# We need to make a bar plot of the number of detections per species I want to use common names for the species (White Shark = carcharodon carcharias, Soupfin Shark = galeorhinus galeus, Bat Ray = myliobatis californica, Leopard Shark = triakis semifasciata, Yellowtail = seriola lalandi, Giant Sea Bass = stereolepis gigas, Sevengill Shark = notorynchus cepedianus)
# Create a bar plot of the number of detections per species using common names
species_common_names <- c(
    "Carcharodon carcharias" = "White Shark",
    "Galeorhinus galeus" = "Soupfin Shark",
    "Myliobatis californica" = "Bat Ray",
    "Triakis semifasciata" = "Leopard Shark",
    "Seriola lalandi" = "Yellowtail",
    "Stereolepis gigas" = "Giant Sea Bass",
    "Notorynchus cepedianus" = "Sevengill Shark"
)

# Add common names to detection_data
detection_data <- detection_data %>%
    mutate(CommonName = species_common_names[SciName]) %>%
    filter(!is.na(CommonName))  # Exclude NA species

# Count the number of detections per species
detections_per_species <- detection_data %>%
    group_by(CommonName) %>%
    summarise(Detections = n())

# Create the bar plot
ggplot(detections_per_species, aes(x = reorder(CommonName, -Detections), y = Detections)) +
    geom_bar(stat = "identity", fill = "#095558") +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0)) +
    labs(title = NULL, x = "Species", y = "Detections") +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 40),
        axis.text.y = element_text(size = 40),
        axis.title.x = element_text(size = 45),
        axis.title.y = element_text(size = 45)
    )

# Export the plot
ggsave("outputs/fig_2.2_detections.png", width = 20, height = 15, units = "in", dpi = 300)

```

# lets look at some statistics associated with figure 2.2. I want to know the total number of detections for each species, the average number of detections per day for each species, and the percent of detections each species represents.
```{r}
# Calculate the number of unique transmitters for each species
transmitter_counts <- detection_data %>%
    group_by(CommonName) %>%
    summarise(Transmitter_Count = n_distinct(Transmitter))

# Print the table
transmitter_counts
# Calculate the total number of detections for each species
total_detections_per_species <- detection_data %>%
    group_by(CommonName) %>%
    summarise(Total_Detections = n())

# Calculate the average number of detections per day for each species
average_detections_per_day <- detection_data %>%
    group_by(CommonName) %>%
    summarise(Avg_Detections_Per_Day = n() / as.numeric(difftime(max(`Date and Time (UTC)`), min(`Date and Time (UTC)`), units = "days")))

# Calculate the percentage of detections each species represents
percent_detections_per_species <- detection_data %>%
  group_by(CommonName) %>%
  summarise(Percentage = n() / nrow(detection_data) * 100)

# Print the results
total_detections_per_species
average_detections_per_day
percent_detections_per_species
```

# Table 1 from tag ID data: Species names, nuumber if transmitters, tag owners, average size (plus or minus standard error)
```{r}

head(tag_data)
str(tag_data)

# list the unique species
unique(tag_data$SciName)

# Count the number of unique transmitters for each species
transmitter_counts <- tag_data %>%
  group_by(SciName) %>%
  summarise(Transmitter_Count = n_distinct(Transmitter))

# Print the table
transmitter_counts

# Ok, now let's find the tag owners starting with Carcharodon carcharias along with the average size plus or minus the standard error for each tag owner
carcharodon_table <- tag_data %>%
    filter(SciName == "Carcharodon carcharias") %>%
    group_by(Owner) %>%
    summarise(Avg_Size = mean(TNL_cm, na.rm = TRUE),
                        SE_Size = sd(TNL_cm, na.rm = TRUE) / sqrt(n()),
                        Transmitter_Count = n_distinct(Transmitter))

# Print the table
carcharodon_table

# Ok, now let's find the tag owners for Galeorhinus galeus along with the average size plus or minus the standard error for each tag owner
galeorhinus_table <- tag_data %>%
    filter(SciName == "Galeorhinus galeus") %>%
    group_by(Owner) %>%
    summarise(Avg_Size = mean(TNL_cm, na.rm = TRUE),
                        SE_Size = sd(TNL_cm, na.rm = TRUE) / sqrt(n()),
                        Transmitter_Count = n_distinct(Transmitter))

# Print the table
galeorhinus_table

# Ok, now let's find the tag owners for Myliobatis californica along with the average size plus or minus the standard error for each tag owner
myliobatis_table <- tag_data %>%
    filter(SciName == "Myliobatis californica") %>%
    group_by(Owner) %>%
    summarise(Avg_Size = mean(disk_length, na.rm = TRUE),
                        SE_Size = sd(disk_length, na.rm = TRUE) / sqrt(n()),
                        Transmitter_Count = n_distinct(Transmitter))

# Print the table
myliobatis_table

# Ok, now let's find the tag owners for Triakis semifasciata along with the average size plus or minus the standard error for each tag owner
triakis_table <- tag_data %>%
    filter(SciName == "Triakis semifasciata") %>%
    group_by(Owner) %>%
    summarise(Avg_Size = mean(TNL_cm, na.rm = TRUE),
                        SE_Size = sd(TNL_cm, na.rm = TRUE) / sqrt(n()),
                        Transmitter_Count = n_distinct(Transmitter))

# Print the table
triakis_table

# Ok, now let's find the tag owners for Seriola lalandi along with the average size plus or minus the standard error for each tag owner
seriola_table <- tag_data %>%
    filter(SciName == "Seriola lalandi") %>%
    group_by(Owner) %>%
    summarise(Avg_Size = mean(TNL_cm, na.rm = TRUE),
                        SE_Size = sd(TNL_cm, na.rm = TRUE) / sqrt(n()),
                        Transmitter_Count = n_distinct(Transmitter))

# Print the table
seriola_table

# Ok, now let's find the tag owners for Stereolepis gigas along with the average size plus or minus the standard error for each tag owner
stereolepis_table <- tag_data %>%
    filter(SciName == "Stereolepis gigas") %>%
    group_by(Owner) %>%
    summarise(Avg_Size = mean(TNL_cm, na.rm = TRUE),
                        SE_Size = sd(TNL_cm, na.rm = TRUE) / sqrt(n()),
                        Transmitter_Count = n_distinct(Transmitter))

# Print the table
stereolepis_table

# Ok, now let's find the tag owners for Notorynchus cepedianus
notorynchus_table <- tag_data %>%
    filter(SciName == "Notorynchus cepedianus") %>%
    group_by(Owner) %>%
    summarise(Avg_Size = mean(TNL_cm, na.rm = TRUE),
                        SE_Size = sd(TNL_cm, na.rm = TRUE) / sqrt(n()),
                        Transmitter_Count = n_distinct(Transmitter))

# Print the table
notorynchus_table
```

## Let's look at some statistics associated with table 1. I want to know the percentages of male and female for each species, and average tag latitude for white sharks associated with each owner.
```{r}
# Calculate percentage of male and female for each species
# Calculate percentage of male and female for each species
sex_percentage <- tag_data %>%
    group_by(SciName, sex) %>%
    summarise(Count = n(), .groups = "drop") %>%
    group_by(SciName) %>%
    mutate(Percentage = Count / sum(Count) * 100)

# Print the table
sex_percentage

# Calculate average tag latitude for white sharks associated with each owner
white_shark_latitude <- tag_data %>%
    filter(SciName == "Carcharodon carcharias") %>%
    group_by(Owner) %>%
    summarise(Avg_Latitude = mean(tag_lat, na.rm = TRUE))

# Print the table
white_shark_latitude
```
# Table 2: Species name, maximum days at liberty, maximum detection span, maximum daily displacement, mean Residency Index (plus or minus standard error)
## Maximum Days at liberty is the maximum number of days between the when_encountered and last detection of a transmitter for the species
## Maximum Detection Span is the maximum number of days between the first and last detection of a transmitter for the species
## Maximum Daily Displacement is the maximum straight line distance between detection locations on any given day for the species
## Mean Residency Index is a metric that's calculated for each individual transmitter as the number of days it was detected at a particular station divided by the total number of days it was detected. 
### Mean Residency Index is then averaged across all transmitters for the species
```{r}

# Exclude detections associated with Island = Santa Barbara
detection_data_filtered <- detection_data %>%
    filter(Island != "Santa Barbara")

# Calculate the maximum days at liberty for each species
max_days_at_liberty <- detection_data_filtered %>%
    group_by(SciName, Transmitter) %>%
    summarise(Days_at_liberty = max(as.numeric(difftime(Last_Detection, when_encountered, units = "days"))))
    max_days_at_liberty <- detection_data_filtered %>%
        group_by(SciName) %>%
        summarise(Max_Days_at_Liberty = max(as.numeric(difftime(Last_Detection, when_encountered, units = "days")), na.rm = TRUE))

# Calculate the maximum detection span for each species (maximum number of days between the first and last detection of a transmitter for each species)
max_detection_span <- detection_data_filtered %>%
    group_by(SciName, Transmitter) %>%
    summarise(Detection_Span = max(as.numeric(difftime(Last_Detection, First_Detection), units = "days")))
    max_detection_span <- detection_data_filtered %>%
        group_by(SciName) %>%
        summarise(Max_Detection_Span = max(as.numeric(difftime(Last_Detection, First_Detection), units = "days")), na.rm = TRUE)

# Calculate the maximum daily displacement for each species. We have to calculate the distance between each detection location on any given day and take the maximum value for each transmitter then find the transmitter with the maximum value for each species. 
##To do this we will use the Date and Time columns to distinguish days and Latitude and Longitude columns to calculate the distance between each detection location on any given day.
## We will use the geosphere package to calculate the distance between each detection location on any given day.
# Calculate the maximum daily displacement for each species and transmitter

max_daily_displacement <- detection_data_filtered %>%
  # Group data by Transmitter and Date
  group_by(SciName, Transmitter, Date) %>%
  # Arrange detections by Date and Time
  arrange(`Date and Time (UTC)`) %>%
  # Compute the pairwise distances for sequential detections on the same day
  mutate(Distance = ifelse(is.na(lag(Longitude)), 0,
                           distHaversine(cbind(Longitude, Latitude), 
                                         lag(cbind(Longitude, Latitude))))) %>%
  # Aggregate maximum displacement per day
  summarise(Max_Displacement_Daily = max(Distance, na.rm = TRUE) / 1000,  # Convert meters to kilometers
            .groups = "drop") %>%
  # Replace any -Inf values with 0 for Max_Displacement_Daily
  mutate(Max_Displacement_Daily = ifelse(Max_Displacement_Daily == -Inf, 0, Max_Displacement_Daily)) %>%
  # Find the maximum daily displacement for each transmitter
  group_by(SciName, Transmitter) %>%
  summarise(Max_Daily_Displacement = max(Max_Displacement_Daily, na.rm = TRUE), 
            .groups = "drop") %>%
  # Replace any -Inf values with 0 for Max_Daily_Displacement
  mutate(Max_Daily_Displacement = ifelse(Max_Daily_Displacement == -Inf, 0, Max_Daily_Displacement)) %>%
  # Find the maximum daily displacement across all transmitters for each species
  group_by(SciName) %>%
  summarise(Max_Daily_Displacement = max(Max_Daily_Displacement, na.rm = TRUE))

# View the result
print(max_daily_displacement)

# Print the table
max_days_at_liberty
max_detection_span
```

# I want to look at the distribution of the Residency Index for each species
```{r}

# Calculate the mean Residency Index for each species
## To do this we need to calculate the number of days each transmitter was detected at a particular station and divide it by the total number of days it was detected.
## We will then average this value across all transmitters for each species.

# Step 1: Calculate the Residency Index (RI) for each detection
ri_data <- detection_data_filtered %>%
  # Group by species, transmitter, and station name
  group_by(SciName, Transmitter, `Station Name`) %>%
  # Count the number of unique days detected at each station
  summarise(Days_Detected_At_Station = n_distinct(Date), .groups = "drop") %>%
  
  # Step 2: Calculate the total days detected across all stations for each transmitter
  group_by(SciName, Transmitter) %>%
  mutate(Total_Days_Detected = sum(Days_Detected_At_Station)) %>%
  
  # Step 3: Calculate the Residency Index (RI) for each detection
  mutate(RI = Days_Detected_At_Station / Total_Days_Detected) %>%
  
  # Remove groupings for clean plotting
  ungroup()

# View the result
print(ri_data)

# Step 4: Calculate the mean RI for each species for reference in the plot
species_means <- ri_data %>%
    group_by(SciName) %>%
    summarise(
        mean_RI = mean(RI, na.rm = TRUE),
        se_RI = sd(RI, na.rm = TRUE) / sqrt(n())
    )

# Print a table with species means and SE for the Residency Index
print(species_means)

# name shape files for the map
sanctuary <- st_read("data/cinms_py2/cinms_py.shp")
coast <- st_read("data/stanford-vx275xn8886-shapefile/vx275xn8886.shp")
islands <- st_read("data/California_Counties/California_Counties.shp")
baja <- st_read("data/nyu_2451_36780/bcs_entidad.shp")
bcali <- st_read("data/nyu_2451_36772/bc_municipio.shp")

# Define the mapping of scientific names to common names
species_common_names <- c(
  "Triakis semifasciata" = "Leopard Shark",
  "Carcharodon carcharias" = "White Shark",
  "Seriola lalandi" = "Yellowtail",
  "Myliobatis californica" = "Bat Ray",
  "Galeorhinus galeus" = "Soupfin Shark",
  "Notorynchus cepedianus" = "Sevengill Shark",
  "Stereolepis gigas" = "Giant Sea Bass"
)

# Add common names to ri_data and remove NA species
ri_data <- ri_data %>%
    mutate(CommonName = factor(species_common_names[SciName], 
                                                         levels = c("Leopard Shark", "White Shark", "Yellowtail", 
                                                                                "Bat Ray", "Soupfin Shark", "Sevengill Shark", 
                                                                                "Giant Sea Bass"))) %>%
    filter(!is.na(CommonName))

# Update species_means with common names and remove NA species
species_means <- species_means %>%
    mutate(CommonName = factor(species_common_names[SciName], 
                                                         levels = c("Leopard Shark", "White Shark", "Yellowtail", 
                                                                                "Bat Ray", "Soupfin Shark", "Sevengill Shark", 
                                                                                "Giant Sea Bass"))) %>%
    filter(!is.na(CommonName))

# Step 5: Create the plot
ggplot(ri_data, aes(x = RI)) +
        geom_histogram(
                fill = "#084d2a", bins = 30
        ) +  # Adjust bins for better resolution
        facet_wrap(~CommonName, scales = "free_y") +
        scale_y_continuous(expand = c(0, 0)) +  # Ensure y-axis starts at the corner
        scale_x_continuous(expand = c(0, 0)) +  # Ensure x-axis starts at the corner
        geom_vline(
                data = species_means, aes(xintercept = mean_RI), color = "red",
                linetype = "solid", size = 1
        ) +  # Add mean line for each species
        geom_vline(
                data = ri_data %>% group_by(CommonName) %>% summarise(median_RI = median(RI, na.rm = TRUE)),
                aes(xintercept = median_RI), color = "blue", linetype = "solid", size = 1
        ) +  # Add median line for each species
        labs(
                title = NULL,
                x = NULL,
                y = NULL
        ) +
        theme_classic() +
        theme(
                axis.text.x = element_text(angle = 45, hjust = 1, size = 30),
                axis.text.y = element_text(color = "black", size = 30),
                legend.text = element_text(size = 24),
                legend.title = element_text(size = 30, face = "bold", margin = margin(b = 10)),
                legend.key.size = unit(1.0, "cm"),
                strip.text = element_text(size = 35, face = "bold")
        )

# export the plot
ggsave("outputs/ri_distribution_plot.png", width = 20, height = 15, units = "in", dpi = 300)

# Step 0: Ensure necessary columns are included in detection_data_filtered
# Resolve many-to-many relationship by ensuring unique rows in detection_data
detection_data_filtered <- detection_data_filtered %>%
  left_join(
    detection_data %>%
      select(`Station Name`, Date, Latitude, Longitude) %>%
      distinct(`Station Name`, Date, Latitude, Longitude),  # Ensure unique rows
    by = c("Station Name", "Date")  # Join on both Station Name and Date
  )

# Step 1: Calculate the Residency Index (RI) for each detection
ri_data <- detection_data_filtered %>%
  # Group by species, transmitter, and station name
  group_by(SciName, Transmitter, `Station Name`) %>%
  # Count the number of unique days detected at each station
  summarise(Days_Detected_At_Station = n_distinct(Date), .groups = "drop") %>%
  
  # Step 2: Calculate the total days detected across all stations for each transmitter
  group_by(SciName, Transmitter) %>%
  mutate(Total_Days_Detected = sum(Days_Detected_At_Station)) %>%
  
  # Step 3: Calculate the Residency Index (RI) for each detection
  mutate(RI = Days_Detected_At_Station / Total_Days_Detected) %>%
  
  # Remove groupings for clean plotting
  ungroup()

# Step 2: Calculate the mean RI at each station for each species
ri_station_data <- ri_data %>%
  group_by(SciName, `Station Name`) %>%
  summarise(
    Mean_RI = mean(RI, na.rm = TRUE),
    Latitude = first(Latitude),  # Ensure Latitude is included
    Longitude = first(Longitude),  # Ensure Longitude is included
    .groups = "drop"
  ) %>%
  filter(!is.na(SciName)) %>%
  mutate(
    Common_Name = factor(
      species_common_names[SciName],
      levels = names(species_common_names)
    )
  )

# Step 3: Create the faceted map plot
ri_station_plot <- ggplot() +
    geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
    geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
    geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
    geom_point(data = ri_station_data, aes(x = Longitude, y = Latitude, size = Mean_RI, color = Mean_RI)) +
    scale_color_gradient(low = "blue", high = "red") +
    scale_size_continuous(range = c(3, 10)) +
    coord_sf(xlim = c(-120.6, -119.3), ylim = c(33.75, 34.5)) +
    theme_map() +
    theme(
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 30, face = "bold"),
        panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 30, face = "bold"),
        legend.text = element_text(size = 24),
        legend.title = element_text(size = 30, face = "bold", margin = margin(b = 10)),
        legend.key.size = unit(1.0, "cm"),
        strip.text = element_text(size = 35, face = "bold"),
        legend.position = c(0.8, 0.001)
    ) +
    scale_y_continuous(breaks = c(33.8, 34.2, 34.6)) +
    scale_x_continuous(breaks = c(-120.4, -120.0, -119.6)) +
    labs(x = "Longitude", y = "Latitude", color = "Mean RI") +
    guides(size = "none") +
    facet_wrap(~ Common_Name)

# Step 4: Print the plot
print(ri_station_plot)

# Export the plot
ggsave("outputs/ri_station_plot.png", width = 20, height = 15, units = "in", dpi = 300)

```

## I want to see mean RI (plus or minus standard error) for each Island. 
```{r}
# Add Island data to ri_data
ri_data <- ri_data %>%
    left_join(detection_data_filtered %>% select(`Station Name`, Island) %>% distinct(), by = "Station Name")

# Calculate the mean Residency Index for each Island
ri_island_data <- ri_data %>%
    group_by(Island) %>%
    summarise(mean_RI = mean(RI, na.rm = TRUE),
              se_RI = sd(RI, na.rm = TRUE) / sqrt(n()))

# Print the table
print(ri_island_data)

```

## I want to look at some statistics associated with each species. Specifically, I want to look for groups of spcies with similar residency characteristics. 
### To do this I'll perform non parametric tests among species to determine if there are significant differences in the Residency Index among species.
```{r}
# Perform non-parametric tests among species to determine if there are significant differences in the Residency Index among species
# Perform Kruskal-Wallis test to compare Residency Index among species
kruskal_test <- kruskal.test(RI ~ SciName, data = ri_data)

# Print the results of the Kruskal-Wallis test
print(kruskal_test)

# If the Kruskal-Wallis test is significant, perform pairwise comparisons using Dunn's test
if (kruskal_test$p.value < 0.05) {
    if (!requireNamespace("FSA", quietly = TRUE)) {
        install.packages("FSA")
    }
    library(FSA)
    dunn_test <- dunnTest(RI ~ SciName, data = ri_data, method = "bonferroni")
    print(dunn_test)
    
    # Extract the results and assign letters to groups
    dunn_results <- dunn_test$res
    dunn_results$SciName <- rownames(dunn_results)
    dunn_results <- dunn_results %>%
        mutate(Group = cut(P.adj, breaks = c(-Inf, 0.05, Inf), labels = c("A", "B")))
    
    print(dunn_results)
}
```



