---
title: "Update_Tag_ID"
author: "Morgan Youngblood"
date: "6/21/2024"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load necessary libraries
```{r}

# Load libraries
library(dplyr)
library(readr) 
library(readxl)
```

# Let's update the tag ID list for the updated data set

## Load updated detection data
```{r load}

detection_data <- read_csv("data/hollings_detection-data_cinms_wco_20240617.csv")
str(detection_data)

```

## Set up time columns
```{r time}
# Convert Date and Time (UTC) to local timezone
detection_data$Local_Time <- as.POSIXct(detection_data$`Date and Time (UTC)`, tz = "America/Los_Angeles")

# Extract date component
detection_data$Date <- as.Date(detection_data$Local_Time)

# Extract month and year from Date
detection_data$Month <- format(detection_data$Date, "%m")
detection_data$Year <- format(detection_data$Date, "%Y")
detection_data$Hour <- format(detection_data$Local_Time, "%H")

# Define seasons based on the month
detection_data$Season <- with(detection_data, ifelse(Month %in% c("03", "04", "05"), "Spring",
                        ifelse(Month %in% c("06", "07", "08"), "Summer",
                        ifelse(Month %in% c("09", "10", "11"), "Fall",
                        "Winter"))))

head(detection_data)

colnames(detection_data)

# Evaluate summary statistics for the new columns
summary(detection_data$Local_Time)
summary(detection_data$Date)
summary(detection_data$Month)
summary(detection_data$Year)
summary(detection_data$Hour)

```

### Filter Spurious detections using the date
```{r filter}

# Filter detections where there are < 2 detections per day
detection_data_clean <- detection_data %>%
  group_by(Date) %>%
  filter(n() > 1)

# Filter for detections > 1
detection_data_clean <- detection_data_clean %>%
  group_by(Transmitter) %>%
  filter(n() > 1) %>%
  ungroup()

# Evaluate the number of unique transmitter IDs filtered out
unique(detection_data$`Transmitter`) %>% length()
unique(detection_data_clean$`Transmitter`) %>% length()

```

## Make Tag ID list from detection data
```{r setup-Tag-ID}

# Create a list of unique transmitter IDs with # of days detected, first detection and last detection
tag_id_list <- detection_data_clean %>%
    group_by(Transmitter) %>%
    summarise(
        Days_Detected = n_distinct(Date),
        First_Detection = min(Date),
        Last_Detection = max(Date)
    ) %>%
    arrange(desc(Days_Detected))

# View the structure of the tag ID list
str(tag_id_list)

```

# Merge Lowe Lab Tag ID list with the updated detection data
```{r Lowe, echo=TRUE}

# Load tag data
LoweLab <- readRDS("data/LoweLabTagInfo (1).rds")

# View the structure of the Lowe Lab tag data
str(LoweLab)

# Merge data from both datasets by the "Transmitter" column
tag_id_list <- tag_id_list %>%
    left_join(LoweLab, by = c("Transmitter" = "Transmitter"))

# View the structure of the merged data
str(tag_id_list)

```

# Merge Nosal Tag ID list with the updated detection data
```{r Nosal, echo=FALSE}

# Load Nosal data
Nosal <- readxl::read_xlsx("data/Nosal_Soupfin_Tag_IDs_2024.xlsx")

# View structure of each dataset to determine how to merge
str(Nosal)

# Which columns have NA values in the Nosal dataset
colSums(is.na(Nosal))

# Filter unnecessary columns from the Nosal data
Nosal <- Nosal %>%
    dplyr::select(-V16_ID, -Code_Space)

    # Add column to Nosal data for "Owner" and make all values "Andrew Nosal"
    Nosal$Owner <- "Andrew Nosal"

    # Merge the datasets on the Transmitter column
    tag_id_list <- tag_id_list %>%
        left_join(Nosal, by = c("Transmitter" = "Tag_ID_Combined"), suffix = c("", ".y")) %>%
        mutate(Owner = coalesce(Owner, Owner.y)) %>%
        dplyr::select(-Owner.y)

# View structure of merged data
str(tag_id_list)

# Use mutate to combine the columns from the two datasets

## mutate SciName
tag_id_list <- tag_id_list %>%
    mutate(SciName = if_else(is.na(SciName), Species, SciName)) %>%
    dplyr::select(-Species)

## mutate TNL_cm
tag_id_list <- tag_id_list %>%
    mutate(TNL_cm = if_else(is.na(TNL_cm), TL, TNL_cm)) %>%
    dplyr::select(-TL)

## mutate sex
tag_id_list <- tag_id_list %>%
    mutate(sex = if_else(is.na(sex), Sex, sex)) %>%
    dplyr::select(-Sex)

## mutate when_encountered
tag_id_list <- tag_id_list %>%
    mutate(when_encountered = if_else(is.na(when_encountered), Tagging_Date, when_encountered)) %>%
    dplyr::select(-Tagging_Date)

## mutate tag_location
tag_id_list <- tag_id_list %>%
    mutate(tag_location = if_else(is.na(tag_location), Tagging_Region, tag_location)) %>%
    dplyr::select(-Tagging_Region)

## mutate Transmitter_SN
tag_id_list <- tag_id_list %>%
    mutate(Transmitter_SN = if_else(is.na(Transmitter_SN), V16_SN, Transmitter_SN)) %>%
    dplyr::select(-V16_SN)

# View the merged data
str(tag_id_list)

```

# Merge Block Data into tag ID list
```{r Block, echo=TRUE}

# Load Block data
Block <- read.csv("data/block_tags_2024june7.csv")

# View structure of each dataset to determine how to merge
str(Block)

# Which columns have NA values in the Block dataset
colSums(is.na(Block))

# Add column to Block data for "Owner" and make all values "Barbara Block"
Block$Owner <- "Barbara Block"

# Add Column to Block data for "SciName" and make all values "Carcharodon carcharias"
Block$Species <- "Carcharodon carcharias"

# Merge the datasets on the Transmitter column
tag_id_list <- tag_id_list %>%
    left_join(Block, by = c("Transmitter" = "codespace"), suffix = c("", ".y")) %>%
        mutate(Owner = coalesce(Owner, Owner.y)) %>%
        dplyr::select(-Owner.y)

# View structure of merged data
str(tag_id_list)

# Change taggingdate to date format
tag_id_list$taggingdate <- as.Date(tag_id_list$taggingdate, format = "%Y-%m-%d")

## mutate when_encountered
tag_id_list <- tag_id_list %>%
    mutate(when_encountered = if_else(is.na(when_encountered), taggingdate, when_encountered)) %>%
    dplyr::select(-taggingdate)

## mutate SciName
tag_id_list <- tag_id_list %>%
    mutate(SciName = if_else(is.na(SciName), Species, SciName)) %>%
    dplyr::select(-Species)

## mutate TNL_cm
tag_id_list <- tag_id_list %>%
    mutate(TNL_cm = if_else(is.na(TNL_cm), tagging_length_cm, TNL_cm)) %>%
    dplyr::select(-tagging_length_cm)

## mutate tag_location
tag_id_list <- tag_id_list %>%
    mutate(tag_location = if_else(is.na(tag_location), seriesname, tag_location)) %>%
    dplyr::select(-seriesname)

# View the merged data
str(tag_id_list)

# Remove unnecessary columns
tag_id_list <- tag_id_list %>%
    dplyr::select(-c(commonname, eventid))

```

# For Transmitter A69-9001-2552, update the Owner to "Taylor Chapple" and SciName to "Notorynchus cepedianus"
```{r sevengill, echo=TRUE}

# Update Owner
tag_id_list <- tag_id_list %>%
    mutate(Owner = if_else(Transmitter == "A69-9001-2552", "Taylor Chapple", Owner))

# Update SciName
tag_id_list <- tag_id_list %>%
    mutate(SciName = if_else(Transmitter == "A69-9001-2552", "Notorynchus cepedianus", SciName))

```

# For Transmitter A69-1601-27071, update the Owner to "Noah Ben-Aderet" and SciName to "Seriola lalandi"
```{r yellowtail, echo=TRUE}

# Update Owner
tag_id_list <- tag_id_list %>%
    mutate(Owner = if_else(Transmitter == "A69-1601-27071", "Noah Ben-Aderet", Owner))

# Update SciName
tag_id_list <- tag_id_list %>%
    mutate(SciName = if_else(Transmitter == "A69-1601-27071", "Seriola lalandi", SciName))

```

# Add SciNames from past updated list
```{r past, echo=TRUE}

# Load past updated tag ID list
past_tag_id_list <- read.csv("outputs/Updated_Tag_ID_20240611.csv")

# Filter out all columns except for Transmitter and SciName
past_tag_id_list <- past_tag_id_list %>%
    dplyr::select(Transmitter, SciName)

# View the structure of the past updated tag ID list
str(past_tag_id_list)

# Merge the past updated tag ID list with the current tag ID list
tag_id_list <- tag_id_list %>%
    left_join(past_tag_id_list, by = "Transmitter", suffix = c("", ".y")) %>%
    mutate(SciName = coalesce(SciName, SciName.y)) %>%
    dplyr::select(-SciName.y)
```

# Save the updated tag ID list
```{r save}

# Save the updated tag ID list as a .csv file
write.csv(tag_id_list, "outputs/updated_tag_id_20240621.csv", row.names = FALSE)

# Create table of orphaned tag IDs
orphaned_tags <- tag_id_list %>%
    filter(is.na(SciName)) %>%
    dplyr::select(Transmitter, Days_Detected, First_Detection, Last_Detection)

# View the table
orphaned_tags

# Save the orphaned tag IDs as a .csv file
write.csv(orphaned_tags, "outputs/orphaned_tags_20240701.csv", row.names = FALSE)

```

