---
title: "White Bat Connectivity Animation"
output: White_Bat_Product Analysis.html
Author: "Morgan Youngblood"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Now for the final feature of our presentation we will look at the detections per individual per station for each species during each month. 

## To display this, we are going to animate the detections per individual per station (represented as the size of the points) on the map for each month. Each month will be a separate frame in the animation, and each species will have its own animation.

### First we need to organize the data so that we can plot it on the map.
```{r organize-data, echo=TRUE, eval=TRUE}

# Load WCO Reciever Data
WCO_reciever <- read.csv("data/wco_detections+tags_20240618.csv")

# Filter for detections associated with Carcharodon carcharias
Carcharias_detections <- WCO_reciever %>%
  filter(SciName == "Carcharodon carcharias")

```

# Filter for transmitters with less than 2 detections
```{r filter-transmitters, echo=TRUE, eval=TRUE}

# Filter for detections > 1
Carcharias_detections <- Carcharias_detections %>%
  group_by(Transmitter) %>%
  filter(n() > 1) %>%
  ungroup()

# Determine the number of unique transmitters in the filtered dataset
unique_transmitter_count <- Carcharias_detections %>%
  summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_count)
```

# Cleanup Islands
```{r cleanup-islands, echo=TRUE, eval=TRUE}

library(dplyr)

# Clean up the Island column by removing the columns with .x and .y
Carcharias_detections <- Carcharias_detections %>%
  rename(Island = Island.x) %>%
  select(-contains(".x"), -contains(".y"))
  

```

# Set Station Order
```{r station-order, echo = False}

# Set levels for Islands (West to East)
island_order <- c("San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Mainland", "Santa Barbara")

# Group by Station and find the longitude and latitude of each
station_location <- Carcharias_detections %>%
  group_by(Station.Name) %>%
  summarise(Longitude = mean(Longitude, na.rm = TRUE),
            Latitude = mean(Latitude, na.rm = TRUE))

print(station_location)

# Write .csv file with station locations
write.csv(station_location, "outputs/station_location.csv")

# Set levels for stations based on longitude (West to East)
station_order <- station_location %>%
  arrange(Longitude) %>%
  pull(Station.Name)

  # Convert Island and Station Name to factors with specified order
  Carcharias_detections <- Carcharias_detections %>%
    mutate(Island = factor(Island, levels = island_order),
           Station.Name = factor(Station.Name, levels = station_order))

```



# Carcharodon carcharias animation

## Make a summary table to determine the number of unique Transmitters, tag_location, tag_lon, and tag_lat for each. We need to decide how we are going to display the tag location for these transmitters since we are going to display the tag location for each individual on the map and color the groups based on the tag location.
```{r carcharodon-summary-table, echo=TRUE, eval=TRUE}
Carcharodon_tag_location <- Carcharias_detections %>%
    group_by(Transmitter) %>%
    summarise(tag_location = first(tag_location),
                        tag_lon = first(tag_lon),
                        tag_lat = first(tag_lat),
                        days_detected = n_distinct(Date),
                        Owner = first(Owner)) %>%
    filter(!is.na(Owner)) %>%
    arrange(tag_location)

print(Carcharodon_tag_location)

# Let's write this table to a .csv file
#write.csv(Carcharodon_tag_location, "outputs/Carcharodon_tag_location.csv")
```
# Map tagged locations
```{r detections-per-individual-animations, echo=TRUE, eval=TRUE}

# Import Carcharodon_tag_location.csv
Carcharodon_tag_location <- read.csv("outputs/Carcharodon_tag_location.csv")

# name shape files for the map
sanctuary <- st_read("data/cinms_py2/cinms_py.shp")
coast <- st_read("data/stanford-vx275xn8886-shapefile/vx275xn8886.shp")
islands <- st_read("data/California_Counties/California_Counties.shp")
baja <- st_read("data/nyu_2451_36780/bcs_entidad.shp")
bcali <- st_read("data/nyu_2451_36772/bc_municipio.shp")

library(ggplot2)

# Make a map of the tag locations for Carcharodon carcharias colored on a gradient based on tag_lat
Carcharodon_tag_Locations_map <- ggplot() +
    geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
    geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
    geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
    geom_sf(data = baja, fill = "#B2B2B2", color = "grey40") +
    geom_sf(data = bcali, fill = "#B2B2B2", color = "grey40") +
    geom_point(data = Carcharodon_tag_location, aes(x = tag_lon, y = tag_lat, color = tag_lat), size = 15) +
    scale_color_gradient(low = "red", high = "blue", guide = guide_legend(title = "Latitude")) +
    coord_sf(xlim = c(-123.025, -114.45), ylim = c(38.3, 27.85)) +
    theme_map() +
    theme(
        legend.position = "bottomleft",
        legend.title = element_text(size = 25),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
        panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 20),
        legend.text = element_text(size = 20)
    ) +
    labs(title = NULL, x = "Longitude", y = "Latitude") +
    scale_x_continuous(breaks = seq(-120, -114, by = 3)) +
    scale_y_continuous(breaks = seq(27, 39, by = 3)) +
    geom_rect(aes(xmin = -120.62, xmax = -119.27, ymin = 33.81, ymax = 34.42), fill = NA, color = "black", size = 2)


print(Carcharodon_tag_Locations_map)

# export the map
ggsave("outputs/Carcharodon_tag_Locations_map.png", width = 15, height = 20, units = "in", dpi = 300)

```

# Now we will animate the detections per month by size of the points with individuals colored by tag_lat. 
```{r animate-detections-per-individual, echo=TRUE, eval=TRUE}

# Load the necessary libraries
library(gganimate)
library(ggplot2)

# Calculate the number of detections per transmitter at each station
detections_per_individual <- Carcharias_detections %>%
    group_by(Station.Name, Transmitter) %>%
    summarise(Detections = n(), .groups = 'drop')

# Calculate average detections per individual, standard error, and sample size for each station
avg_detections_per_individual_per_station <- detections_per_individual %>%
    group_by(Station.Name) %>%
    summarise(
        avg_detections = mean(Detections),
        se_detections = sd(Detections) / sqrt(n()),
        num_transmitters = n_distinct(Transmitter),
        .groups = 'drop'
    ) %>%
    mutate(myaxis = paste0(Station.Name, "\n", "n=", num_transmitters))

# Join the average detections data back to the original Carcharias_detections for plotting
Carcharias_detections <- Carcharias_detections %>%
    left_join(avg_detections_per_individual_per_station, by = c("Station.Name" = "Station.Name"))

# Replace the month numbers with the month names
Carcharias_detections$Month <- factor(Carcharias_detections$Month, levels = 1:12, labels = month.name)

# Filter Detections where tag_lat is NA
Carcharias_detections <- Carcharias_detections %>%
    filter(!is.na(tag_lat))

# Adjust the ggplot object for the animation
Carcharias_animation <- ggplot() +
    geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
    geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
    geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
    geom_point(data = Carcharias_detections, aes(x = Longitude, y = Latitude, size = avg_detections, color = tag_lat, group = Month)) +
    scale_color_gradient(low = "red", high = "blue", guide = guide_colorbar(direction = "horizontal", barwidth = 15, barheight = 3, title.position = "top", title = "Tagging Latitude (°N)", label = list(size = 40)), name = "Tagging Latitude (°N)") +
    scale_size_continuous(name = "Average Detections Per Individual", guide = guide_legend(title.position = "top", title = "Average Detections Per Individual"), range = c(3, 10)) +
    coord_sf(xlim = c(-120.62, -119.27), ylim = c(33.81, 34.42)) +
    theme_map() +
    theme(
        legend.position = "bottom",
        legend.title = element_text(size = 25),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
        panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 20),
        legend.text = element_text(size = 20),
        plot.title = element_text(face = "bold", size = 30) # Ensure title size is consistent
    ) +
    labs(x = "Longitude", y = "Latitude") +
    scale_x_continuous(breaks = seq(-120.4, -119.6, by = 0.4)) +
    scale_y_continuous(breaks = seq(33.9, 34.4, by = 0.4)) +
    transition_states(Month, transition_length = 1, state_length = 1, wrap = FALSE) +
    labs(title = '{closest_state}') + # Dynamic title based on the state (Month)
    shadow_mark(past = FALSE, future = FALSE)

# Animate the plot with adjusted settings for 1-second frames per month
animate(Carcharias_animation, nframes = 12, duration = 12, width = 800, height = 800, fps = 1)

# Save the animation as a .gif file
anim_save("outputs/Carcharias_animation.gif", width = 22, height = 15, dpi = 300, units = "in")
```

# I want to make a static map as an image that shows the total detections per station for Carcarodon carcharias.
```{r carcharias-detections-per-station, echo=TRUE, eval=TRUE}

# Calculate the total detections per station location in the species data
detections_per_station <- Carcharias_detections %>%
    group_by(Station.Name) %>%
    summarise(Detections = n(), Latitude = mean(Latitude), Longitude = mean(Longitude), .groups = 'drop')

# Make a map of the total detections per station for Carcharodon carcharias
Carcharias_detections_per_station_map <- ggplot() +
    geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
    geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
    geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
    geom_point(data = detections_per_station, aes(x = Longitude, y = Latitude, color = Detections, size = Detections), shape = 16) +
  scale_color_gradient(low = "blue", high = "red", guide = guide_colorbar(title = "Detections", label = list(size = 40), direction = "horizontal", barwidth = 15, barheight = 3, title.position = "top", title.hjust = 0.5, title.vjust = 0.5), name = "Detections") +  # Color gradient with yellow as low and red as high
  scale_size(range = c(4, 10), guide = "none") +
    coord_sf(xlim = c(-120.62, -119.27), ylim = c(33.81, 34.42)) +
    theme_map() +
    theme(
        legend.position = "bottom",
        legend.title = element_text(size = 25),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
        panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 20),
        legend.text = element_text(size = 20),
        plot.title = element_text(face = "bold", size = 30)
    ) +
    labs(x = "Longitude", y = "Latitude") +
    scale_x_continuous(breaks = seq(-120.4, -119.6, by = 0.4)) +
    scale_y_continuous(breaks = seq(33.9, 34.4, by = 0.4))

print(Carcharias_detections_per_station_map)

# Export the map as a .png file
ggsave("outputs/Carcharias_detections_per_station_map.png", width = 15, height = 20, units = "in", dpi = 300)

```

# Now I want to make an animation for the bat ray data
```{r filter-bat-ray-data, echo=TRUE, eval=TRUE}

# Load WCO Reciever Data
WCO_reciever <- read.csv("data/wco_detections+tags_20240618.csv")

# Filter for detections associated with Carcharodon carcharias
Myliobatis_detections <- WCO_reciever %>%
  filter(SciName == "Myliobatis californica")

# Filter for detections > 1
Myliobatis_detections <- Myliobatis_detections %>%
    group_by(Transmitter) %>%
    filter(n() > 1) %>%
    ungroup()


  ```

 ```{r bat-detections-station-animation, echo=TRUE, eval=TRUE}

library(dplyr)
library(ggplot2)
library(gganimate)
library(sf) # Assuming spatial data for plotting

# Create a dataset to plot the total detections per month at each station with the myliobatis data
detections_per_station <- Myliobatis_detections %>%
    group_by(Station.Name, Month) %>%
    summarise(Detections = n(), Latitude = mean(Latitude), Longitude = mean(Longitude), .groups = 'drop')

# Replace the month numbers with the month names
detections_per_station$Month <- factor(detections_per_station$Month, levels = 1:12, labels = month.name)

# Calculate the total detections per station for each month
total_detections_per_station <- detections_per_station %>%
    group_by(Month, Station.Name) %>%
    summarise(Total_Detections = sum(Detections))

# Merge the total detections data back to the detections_per_station dataset
detections_per_station <- left_join(detections_per_station, total_detections_per_station, by = c("Month", "Station.Name"))

# Make a ggplot object for the animation
bat_detections_animation <- ggplot() +
  geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +  # Check coast data
  geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +  # Check sanctuary data
  geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +  # Check islands data
  geom_point(data = detections_per_station, aes(x = Longitude, y = Latitude, size = Total_Detections, color = Total_Detections), shape = 16) +  # Check detections_per_station data
  scale_size_continuous(name = "Detections", guide = "none", range = c(3, 10)) +
  scale_color_gradient(low = "blue", high = "red", guide = guide_colorbar(title = "Detections", label = list(size = 40), direction = "horizontal", barwidth = 15, barheight = 3, title.position = "top", title.hjust = 0.5, title.vjust = 0.5), name = "Detections") +
  coord_sf(xlim = c(-120.62, -119.27), ylim = c(33.81, 34.42)) +  # Adjust if necessary
  theme_map() +  # Ensure this is defined or replace with a standard theme
  theme(
    plot.title = element_text(face = "bold", size = 30),
    legend.title = element_text(size = 25),
    legend.text = element_text(size = 20)
  ) +
  labs(x = "Longitude", y = "Latitude") +
  transition_states(Month, transition_length = 1, state_length = 1, wrap = FALSE) +
  labs(title = '{closest_state}') + 
  shadow_mark(past = FALSE, future = FALSE)

# Animate the plot with adjusted settings for 1-second frames per month
animate(bat_detections_animation, nframes = 12, duration = 12, width = 800, height = 800, fps = 1)
# Save the animation as a .gif file
anim_save("outputs/bat_detections_animation.gif", width = 22, height = 15, dpi = 300, units = "in")

 ``` 

 # We're going to look at bat ray detection across a greater area since I have more data for this species. 
 ```{r bat-data-upload, echo=TRUE, eval=TRUE}

library(dplyr)

# Load RDS file with additional bat ray detections
bat_lowe <- readRDS("data/morgan_bat_rays.RDS")

# View the structure of the bat ray data
str(bat_lowe)

# How many unique transmitters are in the bat ray data? I want a list of every unique Transmitter. 

unique_transmitters <- unique(bat_lowe$transmitter)
unique_transmitters

# Load Sorted Tag ID data
Tag_Data <- read.csv("outputs/updated_tag_id_20240621.csv")

str(Tag_Data)

# Filter Tag Data for Bat Rays
Bat_Ray_Tags <- Tag_Data %>%
  filter(SciName == "Myliobatis californica")

# Merge the bat ray data with the tag data
bat_lowe <- bat_lowe %>%
  left_join(Bat_Ray_Tags, by = c("transmitter" = "Transmitter"))


# View the structure of the merged bat ray data
str(bat_lowe)

# How many unique transmitters are in the bat ray data? I want a list of every unique Transmitter.
unique(bat_lowe$transmitter)


 ```

# Load WCO Reciever Data
 ```{r organize-data, echo=TRUE, eval=TRUE}

# Load WCO Reciever Data
WCO_reciever <- read.csv("data/wco_detections+tags_20240618.csv")

# Filter for detections associated with Myliobatis californica
Myliobatis_detections <- WCO_reciever %>%
  filter(SciName == "Myliobatis californica")

```

# Add the bat_low detections to the Myliobatis detections data
```{r add-bat-low-detections, echo=TRUE, eval=TRUE}

str(bat_lowe)
str(Myliobatis_detections)

# Export both datasets to .csv files so that I can edit them manually
#write.csv(bat_lowe, "outputs/bat_lowe.csv")
#write.csv(Myliobatis_detections, "outputs/Myliobatis_detections.csv")

# Load the edited .csv files
bat_lowe <- read.csv("outputs/bat_lowe.csv")
Myliobatis_detections <- read.csv("outputs/Myliobatis_detections.csv")

# Look at the format of the Local_Time columns in both datasets
str(bat_lowe$Local_Time)
str(Myliobatis_detections$Local_Time)

# Convert the Local_Time column in bat_lowe to date time format
bat_lowe$Local_Time <- as.POSIXct(bat_lowe$Local_Time, format = "%Y-%m-%d %H:%M:%S")

# Convert the Local_Time column in Myliobatis_detections to date time format
Myliobatis_detections$Local_Time <- as.POSIXct(Myliobatis_detections$Local_Time, format = "%Y-%m-%dT%H:%M:%SZ")

# Add the bat_low detections to the Myliobatis detections data
colnames(bat_lowe) <- colnames(Myliobatis_detections)
Myliobatis_detections <- rbind(Myliobatis_detections, bat_lowe)

# view the structure of the combined data
str(Myliobatis_detections)

# Look for NA values in the Local_Time column
sum(is.na(Myliobatis_detections$Local_Time))

# How many unique transmitters are in the combined data?
unique(Myliobatis_detections$Transmitter)


# How many unique Stations are in the combined data?
unique(Myliobatis_detections$Station.Name)

# Let's create a table to determine the location of each station
station_location <- Myliobatis_detections %>%
  group_by(Station.Name) %>%
  summarise(Longitude = mean(Longitude, na.rm = TRUE),
            Latitude = mean(Latitude, na.rm = TRUE))

print(station_location)

# Write the station location data to a .csv file
write.csv(station_location, "outputs/station_location_south.csv")

```

# Filter for transmitters with less than 2 detections
```{r filter-transmitters, echo=TRUE, eval=TRUE}

# Filter for detections > 1
Myliobatis_detections <- Myliobatis_detections %>%
  group_by(Transmitter) %>%
  filter(n() > 1) %>%
  ungroup()

```

# Make a summary table to determine the location of each station
```{r bat-ray-summary-location, echo=TRUE, eval=TRUE}

# Make a summary table to determine the Lat and Long of each station
station_location <- Myliobatis_detections %>%
  group_by(Station.Name) %>%
  summarise(Longitude = mean(Longitude, na.rm = TRUE),
            Latitude = mean(Latitude, na.rm = TRUE))


print(station_location)
```

# Let's make a map showing The california Coast with the station locations for Myliobatis californica
```{r map-station-myliobatis, echo=TRUE, eval=TRUE}

library(ggplot2)
library(sf)
library(ggthemes)

# Load the shape files for the map
sanctuary <- st_read("data/cinms_py2/cinms_py.shp")
coast <- st_read("data/stanford-vx275xn8886-shapefile/vx275xn8886.shp")
islands <- st_read("data/California_Counties/California_Counties.shp")
baja <- st_read("data/nyu_2451_36780/bcs_entidad.shp")
bcali <- st_read("data/nyu_2451_36772/bc_municipio.shp")

# Make a map of the station locations for Myliobatis californica
Myliobatis_station_map <- ggplot() +
    geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
    geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
    geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
    geom_sf(data = baja, fill = "#B2B2B2", color = "grey40") +
    geom_point(data = station_location, aes(x = Longitude, y = Latitude), size = 7, color = "#095558") +
    coord_sf(xlim = c(-123.025, -117.5), ylim = c(38, 32.5)) +
    theme_map() +
    theme(
        legend.position = "bottomleft",
        legend.title = element_text(size = 25),
        panel.background = element_rect(fill = "white", color = NA),
        plot.background = element_rect(fill = "white", color = NA),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
        panel.grid.major = element_line(color = "black", linetype = "solid", linewidth = 0.5),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(color = "black", size = 20),
        legend.text = element_text(size = 20)
    ) +
    labs(title = NULL, x = "Longitude", y = "Latitude") +
    scale_x_continuous(breaks = seq(-121, -118, by = 3)) +
    scale_y_continuous(breaks = seq(27, 39, by = 3)) +
    geom_rect(aes(xmin = -120.62, xmax = -117.75, ymin = 33.1, ymax = 34.5), fill = NA, color = "black", size = 2)

print(Myliobatis_station_map)

# Export the map as a .png file
ggsave("outputs/Myliobatis_station_map.png", width = 15, height = 20, units = "in", dpi = 300)

```

# Adjust data to look at total detections per month at each station for Myliobatis californica
```{r total-detections-per-station-myliobatis, echo=TRUE, eval=TRUE}

# Extract the month from the Date and time column
Myliobatis_detections$Month <- as.numeric(format(Myliobatis_detections$Local_Time, "%m"))

# Evaluate the unique months in the Myliobatis data
unique(Myliobatis_detections$Month)

# Create a dataset to plot the total detections per month at each station with the myliobatis data
detections_per_station <- Myliobatis_detections %>%
  group_by(Station.Name, Month, Transmitter) %>%
  summarise(Detections = n(), Latitude = mean(Latitude), Longitude = mean(Longitude), .groups = 'drop')

# Replace the month numbers with the month names
detections_per_station$Month <- factor(detections_per_station$Month, levels = 1:12, labels = month.name)

# Calculate the total detections per station for each month
total_detections_per_station <- detections_per_station %>%
  group_by(Month, Station.Name, Transmitter) %>%
  summarise(Total_Detections = sum(Detections), .groups = 'drop')

# Merge the total detections data back to the detections_per_station dataset
detections_per_station <- left_join(detections_per_station, total_detections_per_station, by = c("Month", "Station.Name", "Transmitter"), suffix = c("", "_total"))

```

# Ok, Now we're going to animate the detections per individual per station for Myliobatis californica
```{r animate-detections-per-individual-myliobatis, echo=TRUE, eval=TRUE}

# Count the number of unique transmitters in the Myliobatis data
unique(Myliobatis_detections$Transmitter)

# Load the necessary libraries
library(gganimate)
library(ggplot2)

# Define the color spectrum
color_spectrum <- colorRampPalette(c("yellow", "orange", "red", "maroon", "purple", "lightblue", "blue", "black"))

# Generate 11 unique colors from the defined spectrum
colors <- color_spectrum(10)


# Adjust the ggplot object for the animation
Myliobatis_animation <- ggplot() +
  geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
  geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
  geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
  geom_point(data = detections_per_station, aes(x = Longitude, y = Latitude, size = Total_Detections, color = Transmitter, group = Month), position = position_jitter(width = 0.02, height = 0.02)) +
  scale_color_manual(values = colors, name = "Rays") +
  scale_size_continuous(name = "Detections", guide = "none", range = c(3, 15)) +
  coord_sf(xlim = c(-120.62, -117.75), ylim = c(33.1, 34.5)) +
  theme_map() +
  theme(
    legend.position = "left",
    legend.title = element_text(size = 25),
    legend.direction = "vertical",
    legend.key.width = unit(1, "cm"),
    legend.key.height = unit(0.5, "cm"),
    legend.text = element_blank(), # Hide legend text
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
    panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(color = "black", size = 20),
    plot.title = element_text(face = "bold", size = 30)
  ) +
  labs(x = "Longitude", y = "Latitude") +
  scale_x_continuous(breaks = seq(-120.4, -118.8, by = 0.8)) +
  scale_y_continuous(breaks = seq(33.1, 34.7, by = 0.8)) +
  transition_states(Month, transition_length = 1, state_length = 1, wrap = FALSE) +
  labs(title = '{closest_state}') +
  shadow_mark(past = FALSE, future = FALSE)

# Animate the plot with adjusted settings for 1-second frames per month
animate(Myliobatis_animation, nframes = 12, duration = 12, width = 800, height = 800, fps = 1)

# Save the animation as a .gif file
anim_save("outputs/Myliobatis_animation.gif", width = 22, height = 15, dpi = 300, units = "in")


```