---
title: "WCO Tag ID Analysis"
author: "Morgan Youngblood"
date: "2024-06-06"
format: html
---

# Introduction

This document presents the analysis of the WCO dataset with Tag ID as of 06/06/24. General animal movement and summary analysis has been performed for the dataset. Some initial analysis has been performed to identify general trends and patterns in the data. These preliminary analysis hve observed the following:

- Detections per transmitter
- Detections per known species
- Transmitters per species
- Detections per station and island
- Unique transmitters detected per station and island
- Number of detections per transmitter at each station
- Average detections per individual, standard error, and sample size for each station
- Median detections per individual, standard error, and sample size for each station
- Detections over time (and other general temporal patterns)

Our analysis will pick up from here and further explore the data to identify any additional patterns or trends that may be present by looking closer at the variation within and among species data in this dataset. First, we will verify the data quality and then proceed with the analysis.

We can start by looking at the number of unique transmitters in the dataset to make sure that the data is complete and that we have all the necessary information to proceed with the analysis.

# load necessary libraries
```{r load-libraries, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)


```

# Data Quality Check
```{r data-quality-check, echo=TRUE, eval=TRUE}

# Load Sorted Tag ID data
Tag_Data <- read.csv("data/sorted_Tag_Data.csv")

str(Tag_Data)

# Load WCO Reciever with Tag ID data
WCO <- read.csv("data/WCO_with_Tag_ID_6_6_24.csv")

# count the number of unique transmitters in Tag_Data 
unique_transmitter_Tag_Data <- Tag_Data %>%
  summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_Tag_Data)

# count the number of unique transmitters in WCO 
unique_transmitter_WCO <- WCO %>% 
    summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_WCO)


```

# We see 253 unique transmitters in the Tag Data and 305 unique transmitters in the WCO data. This is fine because the Tag Data is a subset of the WCO data. We can proceed with the analysis.

# Abacus Plots for Species

## Myliobatis californica
```{r abacus-Myliobatis, echo=TRUE, eval=TRUE}

# check the number of transmitters associated with Mylobatis californica
Number_of_Myliobatis <- WCO %>%
  filter(SciName == "Myliobatis californica") %>%
  summarise(Transmitter_count = n_distinct(Transmitter))

  print(Number_of_Myliobatis)

# Filter the dataset for the specific species
Filtered_Myliobatis <- WCO %>%
  filter(SciName == "Myliobatis californica")

# Create a vector of custom colors and shapes for each individual
custom_colors <- c("#C58ABB", "#B73944", "#8ACE89", "#C15901", "#0072B2", 
                   "#656565", "#009E73", "#56B4E9", "#E69F00", "#000000")
shapes <- c(15, 16, 17, 18, 19, 20, 21, 22, 23, 24)

# Convert 'Date' from character to Date format in the Filtered_Myliobatis dataframe
Filtered_Myliobatis <- Filtered_Myliobatis %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.

# Filter data for pre 2020
Filtered_Myliobatis_pre_2020 <- Filtered_Myliobatis %>% 
  filter(Date < as.Date("2020-01-01"))

# Plot for pre 2020
mylobatis_abacus_pre_2020 <- ggplot(Filtered_Myliobatis_pre_2020, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 month"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
        axis.text.y = element_text(size = 12), # Increase size of y-axis text
        axis.title = element_text(size = 14), # Increase size of axis titles
        plot.title = element_text(hjust = 0.5, size = 16), # Increase size of plot title
        legend.text = element_text(size = 12), # Increase size of legend text
        legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Myliobatis californica (Pre 2020)",
       x = "Island",
       y = "Date",
       color = "Ray",
       shape = "Ray")

print(mylobatis_abacus_pre_2020)

ggsave("outputs/Abacus_Plot_Myliobatis_Pre_2020.png", plot = mylobatis_abacus_pre_2020, width = 10, height = 15, dpi = 300, units = "in")

# Filter data for post 2020
Filtered_Myliobatis_post_2020 <- Filtered_Myliobatis %>% 
  filter(Date >= as.Date("2020-01-01"))

# Plot for post 2020
mylobatis_abacus_post_2020 <- ggplot(Filtered_Myliobatis_post_2020, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 month"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Myliobatis californica (Post 2020)",
       x = "Island",
       y = "Date",
       color = "Ray",
       shape = "Ray")

print(mylobatis_abacus_post_2020)
ggsave("outputs/Abacus_Plot_Myliobatis_Post_2020.png", plot = mylobatis_abacus_post_2020, width = 10, height = 15, dpi = 300, units = "in")

```

## Triakis semifasciata, Seriola lalandi, Stereolepis gigas, and Pseudobatos productus.
```{r abacus-Triakis-Seriola-Stereolepis-Pseudobatos, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Multi_Species <- WCO %>%
  filter(SciName == "Triakis semifasciata" | SciName == "Seriola lalandi" | SciName == "Stereolepis gigas" | SciName == "Pseudobatos productus")

# Create a vector of custom colors and shapes for each species
custom_colors <- c("#C58ABB", "#B73944", "#8ACE89", "#C15901")
shapes <- c(15, 16, 17, 18)

# Convert 'Date' from character to Date format in the Filtered_Multi_Species dataframe
Filtered_Multi_Species <- Filtered_Multi_Species %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where Island is NA
Filtered_Multi_Species <- Filtered_Multi_Species %>%
  filter(!is.na(Island))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the species labeled with unique colors and shapes.
  
  # Plot
Abacus_Multi_Species <- ggplot(Filtered_Multi_Species, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = SciName, shape = SciName)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Triakis semifasciata, Seriola lalandi, Stereolepis gigas, and Pseudobatos productus",
       x = "Island",
       y = "Date",
       color = "Species",
       shape = "Species")

print(Abacus_Multi_Species)

ggsave("outputs/Abacus_Plot_Multi_Species.png", plot = Abacus_Multi_Species, width = 10, height = 15, dpi = 300, units = "in")

```

## Galeorhinus galeus
```{r abacus-Galeorhinus, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Galeorhinus <- WCO %>%
  filter(SciName == "Galeorhinus galeus")

# Create a vector of custom colors and shapes for each individual (We have 31 individuals)
custom_colors <- c("#000000", "grey", "#D7792C", "#D79C2C", "#D7BB2C", 
                   "#D7D42C", "#BFD72C", "#A3D72C", "#7CD72C", "#3AD72C",
                   "#2CD779", "#2CD79F", "#2CD7BF", "#2CCDD7", "#2CA6D7", 
                   "#2C83D7", "#2C60D7", "#2C41D7", "#672CD7", "#8A2CD7",
                   "#AD2CD7", "#CD2CD7", "#D72CA3", "#D72C80", "#D72C60", 
                   "#D72C3A", "#D7592C", "#D7792C", "#D79C2C", "#D7BB2C", "#D7D42C")
shapes <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10,
            1, 2, 3, 4, 5, 6, 7, 8, 9, 10,
            1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)

# Convert 'Date' from character to Date format in the Filtered_Galeorhinus dataframe
Filtered_Galeorhinus <- Filtered_Galeorhinus %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where Island is NA
Filtered_Galeorhinus <- Filtered_Galeorhinus %>%
  filter(!is.na(Island))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.
  
  # Plot
  Abacus_Galeorhinus <- ggplot(Filtered_Galeorhinus, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.3, height = 0.4, alpha = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  guides(color = guide_legend(ncol = 1), shape = guide_legend(ncol = 1)) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Galeorhinus galeus",
       x = "Island",
       y = "Date",
       color = "Shark",
       shape = "Shark")

print(Abacus_Galeorhinus)

ggsave("outputs/Abacus_Plot_Galeorhinus.png", plot = Abacus_Galeorhinus, width = 10, height = 15, dpi = 300, units = "in")

```

# Carcharodon carcharias
```{r abacus-Carcharodon, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Carcharodon <- WCO %>%
  filter(SciName == "Carcharodon carcharias")

# Create a vector of custom shapes to represent sex for the species. 
shapes <- c(1, 3, 2)

# Convert 'Date' from character to Date format in the Filtered_Carcharodon dataframe
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where Island is NA
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  filter(!is.na(Island))

# Filter for detections where sex is NA
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  filter(!is.na(sex))

# Filter for detections where TNL_cm is NA
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  filter(!is.na(TNL_cm))

  # Find the minimum and maximum values of 'TNL_cm' for Carcharodon carcharias. 
min_max_TNL_cm <- Filtered_Carcharodon %>%
  summarise(min_TNL_cm = min(TNL_cm), max_TNL_cm = max(TNL_cm))

print(min_max_TNL_cm)

# Format when_encountered as Date
Filtered_Carcharodon$when_encountered <- as.Date(Filtered_Carcharodon$when_encountered, format = "%Y-%m-%d")

# Create column for adjusted 'TNL_cm' values to account for growth. The species grows 45cm per year until they reach 180cm, then 12cm per year until they reach 330cm. We can keep the values above 330cm as is.

Filtered_Carcharodon$TNL_cm_adjusted <- Filtered_Carcharodon$TNL_cm

# Calculate the number of days since the when_encountered date for each detection
Filtered_Carcharodon$days_since_encountered <- as.numeric(difftime(Filtered_Carcharodon$Date, Filtered_Carcharodon$when_encountered, units = "days"))

# Adjust TNL_cm based on the conditions
Filtered_Carcharodon$TNL_cm_adjusted[Filtered_Carcharodon$TNL_cm < 180] <- Filtered_Carcharodon$TNL_cm[Filtered_Carcharodon$TNL_cm < 180] + 0.12328767 * Filtered_Carcharodon$days_since_encountered[Filtered_Carcharodon$TNL_cm < 180]
Filtered_Carcharodon$TNL_cm_adjusted[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331] <- Filtered_Carcharodon$TNL_cm[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331] + 0.03287671 * Filtered_Carcharodon$days_since_encountered[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331]


  # Create a new variable 'life_stage' based on 'TNL_cm'
Filtered_Carcharodon$life_stage <- cut(Filtered_Carcharodon$TNL_cm_adjusted, 
                                       breaks = c(-Inf, 180, 330, Inf), 
                                       labels = c("YOY", "Juvenile", "Adult"))

# Create a vector of custom colors
custom_colors <- c("YOY" = "#03C41F", "Juvenile" = "#031FC4", "Adult" = "#C4030F")

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.
  
  # Plot
  Abacus_Carcharodon <- ggplot(Filtered_Carcharodon, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = life_stage, shape = sex)) +
  geom_jitter(width = 0.3, height = 0.4, alpha = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  guides(color = guide_legend(ncol = 1), shape = guide_legend(ncol = 1)) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Carcharodon carcharias",
       x = "Island",
       y = "Date",
       color = "Life Stage",
       shape = "Sex")

print(Abacus_Carcharodon)

# Save the plot
ggsave("outputs/Abacus_Plot_Carcharodon.png", plot = Abacus_Carcharodon, width = 10, height = 15, dpi = 300, units = "in")

# Make a table with the following parameters for each individual: when_encountered, date first detected, date last detected, and detections per life stage with columns for (YOY, Juvenile, Adult).
 Filtered_Carcharodon_table <- Filtered_Carcharodon %>%
  group_by(Transmitter) %>%
  summarise(
    when_encountered = min(when_encountered),
    date_first_detected = min(Date),
    date_last_detected = max(Date),
    TNL_cm = mean(TNL_cm, na.rm = TRUE), # Add TNL_cm column
    YOY_detections = sum(life_stage == "YOY"),
    Juvenile_detections = sum(life_stage == "Juvenile"),
    Adult_detections = sum(life_stage == "Adult")
  ) %>%
  arrange(TNL_cm) # Sort by TNL_cm in ascending order

  print(Filtered_Carcharodon_table)

  # Write the table to a csv file
write.csv(Filtered_Carcharodon_table, "data/Carcharodon_Life_Stage_table.csv", row.names = FALSE)

```

