---
title: "WCO Tag ID Analysis"
author: "Morgan Youngblood"
date: "2024-06-06"
format: html
---

# Introduction

This document presents the analysis of the WCO dataset with Tag ID as of 06/06/24. General animal movement and summary analysis has been performed for the dataset. Some initial analysis has been performed to identify general trends and patterns in the data. These preliminary analysis hve observed the following:

- Detections per transmitter
- Detections per known species
- Transmitters per species
- Detections per station and island
- Unique transmitters detected per station and island
- Number of detections per transmitter at each station
- Average detections per individual, standard error, and sample size for each station
- Median detections per individual, standard error, and sample size for each station
- Detections over time (and other general temporal patterns)

Our analysis will pick up from here and further explore the data to identify any additional patterns or trends that may be present by looking closer at the variation within and among species data in this dataset. First, we will verify the data quality and then proceed with the analysis.

We can start by looking at the number of unique transmitters in the dataset to make sure that the data is complete and that we have all the necessary information to proceed with the analysis.

# load necessary libraries
```{r load-libraries, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

```{r packages, warning=FALSE, message=FALSE}
#install.packages("fuzzyjoin")

library(readr)
library(gganimate)
library(lubridate)
library(adehabitatHR)
library(adehabitatHS)
library(adehabitatLT)
library(adehabitatMA)
library(deldir)
library(dplyr)
library(sp)
library(ggplot2)
library(sf)
library(ggmap)
library(stats)
library(viridis)
library(geosphere)
library(zoo)
library(gifski)
library(fuzzyjoin)

```


# Data Quality Check
```{r data-quality-check, echo=TRUE, eval=TRUE}

# Load Sorted Tag ID data
Tag_Data <- read.csv("data/Updated_Tag_ID_20240611.csv")

str(Tag_Data)

# Load WCO Reciever Data
WCO_reciever <- read.csv("data/wco_detections+tags_20240613.csv")

# count the number of unique transmitters in Tag_Data 
unique_transmitter_Tag_Data <- Tag_Data %>%
  summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_Tag_Data)

# count the number of unique transmitters in WCO 
unique_transmitter_WCO <- WCO_reciever %>% 
    summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_WCO)


```

# We see 253 unique transmitters in the Tag Data and 302 unique transmitters in the WCO data. This is fine because the Tag Data is a subset of the WCO data. We can proceed with the analysis.

# Filter for transmitters with less than 2 detections
```{r filter-transmitters, echo=TRUE, eval=TRUE}

# Filter for detections > 1
WCO <- WCO_reciever %>%
  group_by(Station.Name, Transmitter) %>%
  filter(n() > 1) %>%
  ungroup()

# Determine the number of unique transmitters in the filtered dataset
unique_transmitter_count <- WCO %>%
  summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_count)
```

# Abacus Plots for Species

## Myliobatis californica
```{r abacus-Myliobatis, echo=TRUE, eval=TRUE}

# check the number of transmitters associated with Mylobatis californica
Number_of_Myliobatis <- WCO %>%
  filter(SciName == "Myliobatis californica") %>%
  summarise(Transmitter_count = n_distinct(Transmitter))

  print(Number_of_Myliobatis)

# Filter the dataset for the specific species
Filtered_Myliobatis <- WCO %>%
  filter(SciName == "Myliobatis californica")

# Create a vector of custom colors and shapes for each individual
custom_colors <- c("#C58ABB", "#B73944", "#8ACE89", "#C15901", "#0072B2", 
                   "#656565", "#009E73", "#56B4E9", "#E69F00", "#000000")
shapes <- c(15, 16, 17, 18, 19, 20, 21, 22, 23, 24)

# Convert 'Date' from character to Date format in the Filtered_Myliobatis dataframe
Filtered_Myliobatis <- Filtered_Myliobatis %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.

# Filter data for pre 2020
Filtered_Myliobatis_pre_2020 <- Filtered_Myliobatis %>% 
  filter(Date < as.Date("2020-01-01"))

# Plot for pre 2020
mylobatis_abacus_pre_2020 <- ggplot(Filtered_Myliobatis_pre_2020, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 month"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
        axis.text.y = element_text(size = 12), # Increase size of y-axis text
        axis.title = element_text(size = 14), # Increase size of axis titles
        plot.title = element_text(hjust = 0.5, size = 16), # Increase size of plot title
        legend.text = element_text(size = 12), # Increase size of legend text
        legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Myliobatis californica (Pre 2020)",
       x = "Island",
       y = "Date",
       color = "Ray",
       shape = "Ray")

print(mylobatis_abacus_pre_2020)

ggsave("outputs/Abacus_Plot_Myliobatis_Pre_2020.png", plot = mylobatis_abacus_pre_2020, width = 10, height = 15, dpi = 300, units = "in")

# Filter data for post 2020
Filtered_Myliobatis_post_2020 <- Filtered_Myliobatis %>% 
  filter(Date >= as.Date("2020-01-01"))

# Plot for post 2020
mylobatis_abacus_post_2020 <- ggplot(Filtered_Myliobatis_post_2020, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 month"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Myliobatis californica (Post 2020)",
       x = "Island",
       y = "Date",
       color = "Ray",
       shape = "Ray")

print(mylobatis_abacus_post_2020)
ggsave("outputs/Abacus_Plot_Myliobatis_Post_2020.png", plot = mylobatis_abacus_post_2020, width = 10, height = 15, dpi = 300, units = "in")

```

## Triakis semifasciata, Seriola lalandi, Stereolepis gigas, and Pseudobatos productus.
```{r abacus-Triakis-Seriola-Stereolepis-Pseudobatos, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Multi_Species <- WCO %>%
  filter(SciName == "Triakis semifasciata" | SciName == "Seriola lalandi" | SciName == "Stereolepis gigas" | SciName == "Pseudobatos productus")

# Create a vector of custom colors and shapes for each species
custom_colors <- c("#C58ABB", "#B73944", "#8ACE89", "#C15901")
shapes <- c(15, 16, 17, 18)

# Convert 'Date' from character to Date format in the Filtered_Multi_Species dataframe
Filtered_Multi_Species <- Filtered_Multi_Species %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where Island is NA
Filtered_Multi_Species <- Filtered_Multi_Species %>%
  filter(!is.na(Island))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the species labeled with unique colors and shapes.
  
  # Plot
Abacus_Multi_Species <- ggplot(Filtered_Multi_Species, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = SciName, shape = SciName)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Triakis semifasciata, Seriola lalandi, Stereolepis gigas, and Pseudobatos productus",
       x = "Island",
       y = "Date",
       color = "Species",
       shape = "Species")

print(Abacus_Multi_Species)

ggsave("outputs/Abacus_Plot_Multi_Species.png", plot = Abacus_Multi_Species, width = 10, height = 15, dpi = 300, units = "in")

```

## Galeorhinus galeus
```{r abacus-Galeorhinus, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Galeorhinus <- WCO %>%
  filter(SciName == "Galeorhinus galeus")

# Create a vector of custom colors and shapes for each individual (We have 31 individuals)
custom_colors <- c("#000000", "grey", "#D7792C", "#D79C2C", "#D7BB2C", 
                   "#D7D42C", "#BFD72C", "#A3D72C", "#7CD72C", "#3AD72C",
                   "#2CD779", "#2CD79F", "#2CD7BF", "#2CCDD7", "#2CA6D7", 
                   "#2C83D7", "#2C60D7", "#2C41D7", "#672CD7", "#8A2CD7",
                   "#AD2CD7", "#CD2CD7", "#D72CA3", "#D72C80", "#D72C60", 
                   "#D72C3A", "#D7592C", "#D7792C", "#D79C2C", "#D7BB2C", "#D7D42C")
shapes <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10,
            1, 2, 3, 4, 5, 6, 7, 8, 9, 10,
            1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)

# Convert 'Date' from character to Date format in the Filtered_Galeorhinus dataframe
Filtered_Galeorhinus <- Filtered_Galeorhinus %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where Island is NA
Filtered_Galeorhinus <- Filtered_Galeorhinus %>%
  filter(!is.na(Island))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.
  
  # Plot
  Abacus_Galeorhinus <- ggplot(Filtered_Galeorhinus, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.3, height = 0.4, alpha = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  guides(color = guide_legend(ncol = 1), shape = guide_legend(ncol = 1)) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Galeorhinus galeus",
       x = "Island",
       y = "Date",
       color = "Shark",
       shape = "Shark")

print(Abacus_Galeorhinus)

ggsave("outputs/Abacus_Plot_Galeorhinus.png", plot = Abacus_Galeorhinus, width = 10, height = 15, dpi = 300, units = "in")

```

## Carcharodon carcharias
```{r abacus-Carcharodon, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Carcharodon <- WCO %>%
  filter(SciName == "Carcharodon carcharias")

# Look at the number of unique transmitters for Carcharodon carcharias
Number_of_Carcharodon <- Filtered_Carcharodon %>%
  summarise(Transmitter_count = n_distinct(Transmitter))

print(Number_of_Carcharodon)

# Create a vector of custom shapes to represent sex for the species. 
shapes <- c(1, 3, 2)

# Convert 'Date' from character to Date format in the Filtered_Carcharodon dataframe
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where sex is NA
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  filter(!is.na(sex))

# Filter for detections where TNL_cm is NA
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  filter(!is.na(TNL_cm))

  # Find the minimum and maximum values of 'TNL_cm' for Carcharodon carcharias. 
min_max_TNL_cm <- Filtered_Carcharodon %>%
  summarise(min_TNL_cm = min(TNL_cm), max_TNL_cm = max(TNL_cm))

print(min_max_TNL_cm)

# Format when_encountered as Date
Filtered_Carcharodon$when_encountered <- as.Date(Filtered_Carcharodon$when_encountered, format = "%Y-%m-%d")

# Create column for adjusted 'TNL_cm' values to account for growth. The species grows 45cm per year until they reach 180cm, then 12cm per year until they reach 330cm. We can keep the values above 330cm as is.

Filtered_Carcharodon$TNL_cm_adjusted <- Filtered_Carcharodon$TNL_cm

# Calculate the number of days since the when_encountered date for each detection
Filtered_Carcharodon$days_since_encountered <- as.numeric(difftime(Filtered_Carcharodon$Date, Filtered_Carcharodon$when_encountered, units = "days"))

# Adjust TNL_cm based on the conditions
Filtered_Carcharodon$TNL_cm_adjusted[Filtered_Carcharodon$TNL_cm < 180] <- Filtered_Carcharodon$TNL_cm[Filtered_Carcharodon$TNL_cm < 180] + 0.12328767 * Filtered_Carcharodon$days_since_encountered[Filtered_Carcharodon$TNL_cm < 180]
Filtered_Carcharodon$TNL_cm_adjusted[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331] <- Filtered_Carcharodon$TNL_cm[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331] + 0.03287671 * Filtered_Carcharodon$days_since_encountered[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331]


  # Create a new variable 'life_stage' based on 'TNL_cm'
Filtered_Carcharodon$life_stage <- cut(Filtered_Carcharodon$TNL_cm_adjusted, 
                                       breaks = c(-Inf, 180, 330, Inf), 
                                       labels = c("YOY", "Juvenile", "Adult"))

# Create a vector of custom colors
custom_colors <- c("YOY" = "#03C41F", "Juvenile" = "#031FC4", "Adult" = "#C4030F")

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.
  
  # Plot
  Abacus_Carcharodon <- ggplot(Filtered_Carcharodon, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa", "Santa Barbara")), y = Date, color = life_stage, shape = sex)) +
  geom_jitter(width = 0.3, height = 0.4, alpha = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  guides(color = guide_legend(ncol = 1), shape = guide_legend(ncol = 1)) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Carcharodon carcharias",
       x = "Island",
       y = "Date",
       color = "Life Stage",
       shape = "Sex")

print(Abacus_Carcharodon)

# Save the plot
ggsave("outputs/Abacus_Plot_Carcharodon.png", plot = Abacus_Carcharodon, width = 10, height = 15, dpi = 300, units = "in")

# Make a table with the following parameters for each individual: when_encountered, date first detected, date last detected, and detections per life stage with columns for (YOY, Juvenile, Adult).
 Filtered_Carcharodon_table <- Filtered_Carcharodon %>%
  group_by(Transmitter) %>%
  summarise(
    when_encountered = min(when_encountered),
    date_first_detected = min(Date),
    date_last_detected = max(Date),
    TNL_cm = mean(TNL_cm, na.rm = TRUE), # Add TNL_cm column
    YOY_detections = sum(life_stage == "YOY"),
    Juvenile_detections = sum(life_stage == "Juvenile"),
    Adult_detections = sum(life_stage == "Adult")
  ) %>%
  arrange(TNL_cm) # Sort by TNL_cm in ascending order

  print(Filtered_Carcharodon_table)

  # Write the table to a csv file
write.csv(Filtered_Carcharodon_table, "data/Carcharodon_Life_Stage_table.csv", row.names = FALSE)

```

### Make a movie for shark A69-9001-13501 
```{r shark-movie-13501, echo=TRUE, eval=TRUE}

# Filter the data for the specific shark transmitter
A69_9001_13501 <- Filtered_Carcharodon %>% 
  filter(Transmitter == "A69-9001-13501")

# Read the shapefiles
islands <- st_read("data/cinms_py2/cinms_py.shp")
coast <- st_read("data/stanford-vx275xn8886-shapefile/vx275xn8886.shp")

# Create a Daily column for easier plotting
detections_per_station_daily_13501 <- A69_9001_13501 %>%
  mutate(Daily = as.Date(Date)) %>%
  group_by(Longitude, Latitude, Island, Year = year(Daily), Daily) %>%
  summarise(Detections = n(), TNL_cm_adjusted = mean(TNL_cm_adjusted)) %>%
  ungroup()

base_plot_13501 <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_daily_13501, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted), size = 5) +
  scale_color_gradient(low = "yellow", high = "red") +  # Set the color gradient from red to yellow
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Total Length (cm)")

# Create the animation
animation_plot_13501 <- base_plot_13501 +
  geom_point(data = detections_per_station_daily_13501, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted)) +
  transition_states(states = Daily, transition_length = 20, state_length = 20) +
  labs(title = paste("Detections M 13501:", "{closest_state}"))
```

#### Animate 13510 and save
```{r, eval=FALSE}

animate(animation_plot_13501, nframes = length(unique(as.vector(detections_per_station_daily_13501$Daily))), fps = 3, width = 800, height = 600, renderer = gifski_renderer("outputs/13501_animation.gif"))

# Set levels for stations based on longitude (West to East)
station_order <- A69_9001_13501 %>%
  arrange(Longitude) %>%
  pull(Station.Name)

# Remove duplicated levels
station_order <- unique(station_order)

# Convert Island and Station Name to factors with specified order
A69_9001_13501 <- A69_9001_13501 %>%
  mutate(Station.Name = factor(Station.Name, levels = station_order))

# Create abacus for 13510
Abacus_13501 <- ggplot(A69_9001_13501, aes(x = factor(`Station.Name`, levels = station_order), y = Date, color = TNL_cm_adjusted)) +
  geom_point(shape = 19, size = 5) +
  scale_color_viridis_c(option = "plasma", direction = -1) +
  labs(title = "Abacus Plot for 13510", x = "Station", y = "Date", color = "TNL") +
  theme_minimal() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_rect(fill = "darkgrey"),
        plot.background = element_rect(fill = "darkgrey"))

print(Abacus_13501)
ggsave("outputs/Abacus_Plot_13501.png", plot = Abacus_13501, width = 10, height = 15, dpi = 300, units = "in")

```

## Ok, now that we know how to do this, let's look at some other Charcarodon carcharias transmitters that may be of interest.

### Identify the unique transmitters for Carcharodon carcharias with the greatest detection span and the greatest number of detections.
```{r shark-summary, echo=TRUE, eval=TRUE}

# Find the unique transmitters for Carcharodon carcharias with the greatest detection span
Carcharodon_span <- Filtered_Carcharodon %>%
  group_by(Transmitter) %>%
  summarize(First_Detection = min(Date),
            Last_Detection = max(Date),
            Detection_Span = as.numeric(max(Date) - min(Date)),
            Number_of_Detections = n()) %>%
  arrange(desc(Detection_Span))

View(Carcharodon_span)

```

### Make a movie for shark A69-1601-30029 
```{r shark-movie-30029, echo=TRUE, eval=TRUE}

# Filter the data for the specific shark transmitter
A69_1601_30029 <- Filtered_Carcharodon %>% 
  filter(Transmitter == "A69-1601-30029")

# Create a Daily column for easier plotting
detections_per_station_daily_30029 <- A69_1601_30029 %>%
  mutate(Daily = as.Date(Date)) %>%
  group_by(Longitude, Latitude, Island, Year = year(Daily), Daily) %>%
  summarise(Detections = n(), TNL_cm_adjusted = mean(TNL_cm_adjusted)) %>%
  ungroup()

base_plot_30029 <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_daily_30029, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted), size = 5) +
  scale_color_gradient(low = "yellow", high = "red") +  # Set the color gradient from red to yellow
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Total Length (cm)")

# Create the animation
animation_plot_30029 <- base_plot_30029 +
  geom_point(data = detections_per_station_daily_30029, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted)) +
  transition_states(states = Daily, transition_length = 20, state_length = 20) +
  labs(title = paste("Detections M 30029:", "{closest_state}"))
  ```

#### Animate 30029 and save
```{r, eval=FALSE}
animate(animation_plot_30029, nframes = length(unique(as.vector(detections_per_station_daily_30029$Daily))), fps = 3, width = 800, height = 600, renderer = gifski_renderer("outputs/30029_animation.gif"))

# Set levels for stations based on longitude (West to East)
station_order <- A69_1601_30029 %>%
  arrange(Longitude) %>%
  pull(Station.Name)

# Remove duplicated levels
station_order <- unique(station_order)

# Convert Island and Station Name to factors with specified order
A69_1601_30029 <- A69_1601_30029 %>%
  mutate(Station.Name = factor(Station.Name, levels = station_order))

# Create abacus for 30029
Abacus_30029 <- ggplot(A69_1601_30029, aes(x = factor(`Station.Name`, levels = station_order), y = Date, color = TNL_cm_adjusted)) +
  geom_point(shape = 19, size = 5) +
  scale_color_viridis_c(option = "plasma", direction = -1) +
  labs(title = "Abacus Plot for 30029", x = "Station", y = "Date", color = "TNL") +
  theme_minimal() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_rect(fill = "darkgrey"),
        plot.background = element_rect(fill = "darkgrey"))

print(Abacus_30029)
ggsave("outputs/Abacus_Plot_30029.png", plot = Abacus_30029, width = 10, height = 15, dpi = 300, units = "in")
```

### Make a movie for shark A69-9001-24384 
```{r shark-movie-24384, echo=TRUE, eval=TRUE}

# Filter the data for the specific shark transmitter
A69_9001_24384 <- Filtered_Carcharodon %>% 
  filter(Transmitter == "A69-9001-24384")

# Create a Daily column for easier plotting
detections_per_station_daily_24384 <- A69_9001_24384 %>%
  mutate(Daily = as.Date(Date)) %>%
  group_by(Longitude, Latitude, Island, Year = year(Daily), Daily) %>%
  summarise(Detections = n(), TNL_cm_adjusted = mean(TNL_cm_adjusted)) %>%
  ungroup()

base_plot_24384 <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_daily_24384, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted), size = 5) +
  scale_color_gradient(low = "yellow", high = "red") +  # Set the color gradient from red to yellow
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Total Length (cm)")

# Create the animation
animation_plot_24384 <- base_plot_24384 +
  geom_point(data = detections_per_station_daily_24384, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted)) +
  transition_states(states = Daily, transition_length = 20, state_length = 20) +
  labs(title = paste("Detections M 24384:", "{closest_state}"))
  ```

#### Animate 24384 and save
```{r, eval=FALSE}
animate(animation_plot_24384, nframes = length(unique(as.vector(detections_per_station_daily_24384$Daily))), fps = 3, width = 800, height = 600, renderer = gifski_renderer("outputs/24384_animation.gif"))

# Set levels for stations based on longitude (West to East)
station_order <- A69_9001_24384 %>%
  arrange(Longitude) %>%
  pull(Station.Name)

# Remove duplicated levels
station_order <- unique(station_order)

# Convert Island and Station Name to factors with specified order
A69_9001_24384 <- A69_9001_24384 %>%
  mutate(Station.Name = factor(Station.Name, levels = station_order))

# Create abacus for 24384
Abacus_24384 <- ggplot(A69_9001_24384, aes(x = factor(`Station.Name`, levels = station_order), y = Date, color = TNL_cm_adjusted)) +
  geom_point(shape = 19, size = 5) +
  scale_color_viridis_c(option = "plasma", direction = -1) +
  labs(title = "Abacus Plot for 24384", x = "Station", y = "Date", color = "TNL") +
  theme_minimal() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_rect(fill = "darkgrey"),
        plot.background = element_rect(fill = "darkgrey"))

print(Abacus_24384)
ggsave("outputs/Abacus_Plot_24384.png", plot = Abacus_24384, width = 10, height = 15, dpi = 300, units = "in")

```

### Make a movie for all three sharks 
```{r shark-movie-all, echo=TRUE, eval=TRUE}

# Filter the data for the specific shark transmitters
male_cohort <- Filtered_Carcharodon %>% 
  filter(Transmitter %in% c("A69-1601-30029", "A69-9001-24384", "A69-9001-13501"))

# Create a Daily column for easier plotting
detections_per_station_daily_mcoh <- male_cohort %>%
  mutate(Daily = as.Date(Date)) %>%
  group_by(Longitude, Latitude, Island, Year = year(Daily), Daily, Transmitter) %>%
  summarise(Detections = n(), TNL_cm_adjusted = mean(TNL_cm_adjusted)) %>%
  ungroup()

base_plot_mcoh <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_daily_mcoh, aes(x = Longitude, y = Latitude, color = Transmitter), size = 5) +
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Shark")

# Create the animation
animation_plot_mcoh <- base_plot_mcoh +
  geom_point(data = detections_per_station_daily_mcoh, aes(x = Longitude, y = Latitude, color = Transmitter)) +
  transition_states(states = Daily, transition_length = 20, state_length = 20) +
  labs(title = paste("Detections Male Cohort:", "{closest_state}"))
```

#### Animate all three sharks and save
```{r, eval=FALSE}
animate(animation_plot_mcoh, nframes = length(unique(as.vector(detections_per_station_daily_mcoh$Daily))), fps = 3, width = 800, height = 600, renderer = gifski_renderer("outputs/male_cohort_animation.gif"))

# Set levels for stations based on longitude (West to East)
station_order <- male_cohort %>%
  arrange(Longitude) %>%
  pull(Station.Name)

# Remove duplicated levels
station_order <- unique(station_order)

# Convert Island and Station Name to factors with specified order
male_cohort <- male_cohort %>%
  mutate(Station.Name = factor(Station.Name, levels = station_order))

# Create abacus for all
Abacus_mcoh <- ggplot(male_cohort, aes(x = factor(`Station.Name`, levels = station_order), y = Date, color = Transmitter)) +
  geom_point(shape = 19, size = 5) +
  labs(title = "Abacus Plot Male Cohort", x = "Station", y = "Date", color = "Shark") +
  theme_minimal() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_rect(fill = "darkgrey"),
        plot.background = element_rect(fill = "darkgrey"))

print(Abacus_mcoh)
ggsave("outputs/Abacus_Male_Cohort.png", plot = Abacus_mcoh, width = 10, height = 15, dpi = 300, units = "in")
```
