---
title: "WCO Tag ID Analysis"
author: "Morgan Youngblood"
date: "2024-06-06"
format: html
---

# Introduction

This document presents the analysis of the WCO dataset with Tag ID as of 06/06/24. General animal movement and summary analysis has been performed for the dataset. Some initial analysis has been performed to identify general trends and patterns in the data. These preliminary analysis hve observed the following:

- Detections per transmitter
- Detections per known species
- Transmitters per species
- Detections per station and island
- Unique transmitters detected per station and island
- Number of detections per transmitter at each station
- Average detections per individual, standard error, and sample size for each station
- Median detections per individual, standard error, and sample size for each station
- Detections over time (and other general temporal patterns)

Our analysis will pick up from here and further explore the data to identify any additional patterns or trends that may be present by looking closer at the variation within and among species data in this dataset. First, we will verify the data quality and then proceed with the analysis.

We can start by looking at the number of unique transmitters in the dataset to make sure that the data is complete and that we have all the necessary information to proceed with the analysis.


# load necessary libraries
```{r load-libraries, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

```{r packages, warning=FALSE, message=FALSE}
#install.packages("fuzzyjoin")
#options(repos = c(CRAN = "https://cloud.r-project.org"))
#install.packages("reticulate")
#install.packages("ggthemes")
library(ggthemes)
library(reticulate)
library(reticulate)
library(extrafont)
library(readr)
library(gganimate)
library(lubridate)
library(adehabitatHR)
library(adehabitatHS)
library(adehabitatLT)
library(adehabitatMA)
library(deldir)
library(dplyr)
library(sp)
library(ggplot2)
library(sf)
library(ggmap)
library(stats)
library(viridis)
library(geosphere)
library(zoo)
library(gifski)
library(fuzzyjoin)

```


# Data Quality Check
```{r data-quality-check, echo=TRUE, eval=TRUE}

# Load Sorted Tag ID data
Tag_Data <- read.csv("outputs/updated_tag_id_20240621.csv")

str(Tag_Data)

# Load WCO Reciever Data
WCO_reciever <- read.csv("data/wco_detections+tags_20240618.csv")

# count the number of unique transmitters in Tag_Data 
unique_transmitter_Tag_Data <- Tag_Data %>%
  summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_Tag_Data)

# count the number of unique transmitters in WCO 
unique_transmitter_WCO <- WCO_reciever %>% 
    summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_WCO)


# Print the names of unique Receivers in the WCO data
unique_recievers <- WCO_reciever %>%
  distinct(Receiver) %>%
  pull(Receiver)
print(unique_recievers)

print(unique_recievers)
```

# We see 277 unique transmitters in the Tag Data and 326 unique transmitters in the WCO data. This is fine because the Tag Data is a subset of the WCO data. We can proceed with the analysis.

# Filter for transmitters with less than 2 detections
```{r filter-transmitters, echo=TRUE, eval=TRUE}

# Filter for detections > 1
WCO <- WCO_reciever %>%
  group_by(Transmitter) %>%
  filter(n() > 1) %>%
  ungroup()

# Determine the number of unique transmitters in the filtered dataset
unique_transmitter_count <- WCO %>%
  summarise(unique_count = n_distinct(Transmitter))

print(unique_transmitter_count)
```

# Cleanup Islands
```{r cleanup-islands, echo=TRUE, eval=TRUE}

library(dplyr)

# Clean up the Island column by removing the columns with .x and .y
WCO <- WCO %>%
  select(-("Island.x"), -("Island.y"))
  

```

# Set Station Order
```{r station-order, echo = False}

# Set levels for Islands (West to East)
island_order <- c("San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Mainland", "Santa Barbara")

# Group by Station and find the longitude and latitude of each
station_location <- WCO %>%
  group_by(Station.Name) %>%
  summarise(Longitude = mean(Longitude, na.rm = TRUE),
            Latitude = mean(Latitude, na.rm = TRUE))

print(station_location)

# Write .csv file with station locations
write.csv(station_location, "outputs/station_location.csv")

# Set levels for stations based on longitude (West to East)
station_order <- station_location %>%
  arrange(Longitude) %>%
  pull(Station.Name)

  # Convert Island and Station Name to factors with specified order
  WCO <- WCO %>%
    mutate(Island = factor(Island, levels = island_order),
           Station.Name = factor(Station.Name, levels = station_order))

```


# Ok, let's look at the summary statistics for the data
```{r summary-statistics, echo=TRUE, eval=TRUE}

# When was the first and last detection in the dataset?
first_detection <- min(WCO$Date)
last_detection <- max(WCO$Date)

print(first_detection)
print(last_detection)

# Detections per station
detections_per_station <- WCO %>%
  group_by(Station.Name) %>%
  summarise(Detections = n())

# Add latitude and longitude data to detections_per_station
detections_per_station <- left_join(detections_per_station, station_location, by = "Station.Name")

  # Add Island data to detections_per_station
  detections_per_station <- left_join(detections_per_station, WCO %>% dplyr::select(Station.Name, Island) %>% distinct(), by = "Station.Name")

print(detections_per_station)

# Export to CSV
write.csv(detections_per_station, "outputs/detections_per_station.csv")

# Filter out sttions associated with Santa Barbara
detections_per_station_no_sb <- detections_per_station %>%
  filter(Island != "Santa Barbara")

print(detections_per_station_no_sb)

# Export to CSV
write.csv(detections_per_station_no_sb, "outputs/detections_per_station_no_sb.csv")

# Bar plot for Detections per Station
ggplot(detections_per_station, aes(x = Station.Name, y = Detections)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Detections per Station", x = "Station Name", y = "Detections") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Detections per Island
detections_per_island <- WCO %>%
  group_by(Island) %>%
  summarise(Detections = n())

print(detections_per_island)

# Bar plot for Detections per Island
ggplot(detections_per_island, aes(x = Island, y = Detections)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Detections per Island", x = "Island", y = "Detections") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

  # Let's look at detections per species in the WCO data 
  detections_per_species <- WCO %>%
    filter(!is.na(SciName)) %>%
    filter(Island != "Santa Barbara") %>%
    group_by(SciName) %>%
    summarise(Detections = n()) %>%
    arrange(desc(Detections))

  print(detections_per_species)

  # Plot Detections per Species
detections_per_species_plot <- ggplot(detections_per_species, aes(x = reorder(SciName, -Detections), y = Detections)) +
  geom_bar(stat = "identity", fill = "#095558") +
  theme_classic() +
  labs(title = NULL, x = "Species", y = "Detections") +
  theme(text = element_text(family = "Arial"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
        axis.title = element_text(size = 30),
        axis.text.y = element_text(size = 25)) +
  scale_x_discrete(labels = c("Giant Sea Bass", "White Shark", "Soupfin Shark", "Sevengill Shark","Bat Ray",  "Yellowtail Amberjack", "Leopard Shark")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))

print(detections_per_species_plot)

# Save plot as image file
ggsave("outputs/Detections_per_Species.png", plot = detections_per_species_plot, width = 10, height = 8, dpi = 300, units = "in")

# Let's look at transmitters per species in the WCO data
transmitters_per_species <- WCO %>%
  filter(!is.na(SciName)) %>%
  group_by(SciName) %>%
  summarise(Transmitters = n_distinct(Transmitter)) %>%
  arrange(desc(Transmitters))

print(transmitters_per_species)

# Transmitters per species
transmitters_per_species_plot <- ggplot(transmitters_per_species, aes(x = reorder(SciName, -Transmitters), y = Transmitters)) +
  geom_bar(stat = "identity", fill = "#244056") +
  theme_minimal() +
  labs(x = "Species", y = "Transmitters") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 15, color = "white"),
        axis.title = element_text(size = 20, color = "white"),
        axis.text.y = element_text(size = 15, color = "white"),
        plot.background = element_rect(fill = "#5d9fba"),
        panel.background = element_rect(fill = "#5d9fba"),
        axis.line = element_line(color = "white"),
        axis.ticks = element_line(color = "white"),
        axis.ticks.length = unit(0.3, "cm"),
        axis.ticks.margin = unit(0.3, "cm"),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y.left = element_blank(),
        axis.ticks.x.bottom = element_blank(),
        axis.ticks.y.right = element_blank(),
        axis.ticks.x.top = element_blank(),
        axis.ticks.length.y = unit(0.3, "cm"),
        axis.ticks.length.x = unit(0.3, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.spacing = unit(0, "lines"),
        plot.margin = unit(c(1, 1, 1, 1), "cm")) +
  scale_x_discrete(labels = c("White Shark", "Soupfin Shark", "Bat Ray", "Leopard Shark", "Yellowtail Amberjack", "Sevengill Shark", "Giant Sea Bass", "Shovelnose Guitarfish")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))

print(transmitters_per_species_plot)

# Save plot as image file
ggsave("outputs/Transmitters_per_Species.png", plot = transmitters_per_species_plot, width = 10, height = 6, dpi = 300, units = "in")

# Let's look at the size range and number of transmitters of Carcharodon carcharias associated with each tag Owner
Carcharodon_summary <- WCO %>%
  filter(SciName == "Carcharodon carcharias") %>%
  group_by(Owner) %>%
  summarise(num_transmitters = n_distinct(Transmitter),
            avg_TNL_cm = mean(TNL_cm, na.rm = TRUE),
            se_TNL_cm = sd(TNL_cm, na.rm = TRUE) / sqrt(sum(!is.na(TNL_cm))),
            num_detections = n())

print(Carcharodon_summary)

# Let's make the same summary table for Galeorhinus galeus
Galeorhinus_summary <- WCO %>%
  filter(SciName == "Galeorhinus galeus") %>%
  group_by(Owner) %>%
  summarise(num_transmitters = n_distinct(Transmitter),
            avg_TNL_cm = mean(TNL_cm, na.rm = TRUE),
            se_TNL_cm = sd(TNL_cm, na.rm = TRUE) / sqrt(sum(!is.na(TNL_cm))),
            num_detections = n())

print(Galeorhinus_summary)

# Do the same for Mylobatis californica
Myliobatis_summary <- WCO %>%
  filter(SciName == "Myliobatis californica") %>%
  group_by(Owner) %>%
  summarise(num_transmitters = n_distinct(Transmitter),
            avg_disk_length = mean(disk_length, na.rm = TRUE),
            se_disk_length = sd(disk_length, na.rm = TRUE) / sqrt(sum(!is.na(disk_length))),
            num_detections = n())

print(Myliobatis_summary)

# Do the same for Triakis semifasciata
Triakis_summary <- WCO %>%
  filter(SciName == "Triakis semifasciata") %>%
  group_by(Owner) %>%
  summarise(num_transmitters = n_distinct(Transmitter),
            avg_TNL_cm = mean(TNL_cm, na.rm = TRUE),
            se_TNL_cm = sd(TNL_cm, na.rm = TRUE) / sqrt(sum(!is.na(TNL_cm))),
            num_detections = n())

            # Let's look at the TNL_cm size of each individual for Triakis semifasciata
Triakis_individuals <- WCO %>%
  filter(SciName == "Triakis semifasciata") %>%
  group_by(Transmitter) %>%
  reframe(TNL_cm = TNL_cm,
          num_detections = n())

print(Triakis_individuals)

print(Triakis_summary)

# Do the same for Seriola lalandi
Seriola_summary <- WCO %>%
  filter(SciName == "Seriola lalandi") %>%
  group_by(Owner) %>%
  summarise(num_transmitters = n_distinct(Transmitter),
            avg_TNL_cm = mean(TNL_cm, na.rm = TRUE),
            se_TNL_cm = sd(TNL_cm, na.rm = TRUE) / sqrt(sum(!is.na(TNL_cm))),
            num_detections = n())

            # look at all the transmitters associated with Seriola lalandi
Seriola_individuals <- WCO %>%
  filter(SciName == "Seriola lalandi") %>% 
  group_by(Transmitter) %>%
  summarise(num_detections = n()) %>%
  dplyr::select(Transmitter)

print(Seriola_individuals)

print(Seriola_summary)


```

## Map Spatial Data
```{r map-spatial-data, echo=TRUE, eval=TRUE}
  # Read the shapefiles
sanctuary <- st_read("data/cinms_py2/cinms_py.shp")
coast <- st_read("data/stanford-vx275xn8886-shapefile/vx275xn8886.shp")
islands <- st_read("data/California_Counties/California_Counties.shp")

# Create map showing station loacations
map_stations <- ggplot() +
  geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +  # Add the coastline shapefile with thicker line
  geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +  
  geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +  # Plot the islands
  geom_point(data = detections_per_station %>% filter(Island != "Santa Barbara"), aes(x = Longitude, y = Latitude), color = "#08C6F4", size = 7, shape = 19, stroke = 1) +  # Plot points in one color and size with black outline
  theme(legend.position = "none") +  # Remove legends
  coord_sf(xlim = c(-120.62, -119.27), ylim = c(33.81, 34.42)) +  # Set boundaries
  theme_map() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
    panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),  # Add gridlines with decreased width
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    axis.text.y = element_text(color = "black", size = 25)  # Add labels for latitude
  ) +
  labs(title = NULL, x = "Longitude", y = "Latitude") +
  scale_x_continuous(breaks = seq(-120.4, -119.6, by = 0.4)) +
  scale_y_continuous(breaks = seq(33.9, 34.4, by = 0.4))

print(map_stations)

# Export the map as an image file
ggsave("outputs/map_stations.png", plot = map_stations, width = 15, height = 10, dpi = 300, units = "in")

# Create the scatter plot for detections per station with shapefile
map_total_detections <- ggplot() +
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = sanctuary, fill = "gray90", color = "black") +  
  geom_sf(data = islands, fill = "gray70", color = "black") +  # Plot the islands
  geom_point(data = detections_per_station, aes(x = Longitude, y = Latitude, color = Detections), size = 3) +
  scale_color_viridis_c(limits = c(0, max(detections_per_station$Detections))) +  # Color gradient with appropriate limits
  coord_sf(xlim = c(-120.65, -118.9), ylim = c(33.3, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
   panel.background = element_rect(fill = "white", color = NA),  # White background
   plot.background = element_rect(fill = "white", color = NA),   # White plot background
   axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Detections per Station", x = "Longitude", y = "Latitude", color = "Detections")

print(map_total_detections)

# Now make the same map with detections associated with santa barbara filtered out
map_total_detections_no_sb <- ggplot() +
  geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +  # Add the coastline shapefile with thicker line
  geom_sf(data = sanctuary, fill = "#D7E6F9", color = "#B2B2B2") +  
  geom_sf(data = islands, fill = "#B2B2B2", color = "#D7E6F9") +  # Plot the islands
  geom_point(data = detections_per_station %>% filter(Island != "Santa Barbara"), aes(x = Longitude, y = Latitude, color = Detections, size = Detections), shape = 16) +
  scale_color_gradient(low = "blue", high = "red", guide = guide_colorbar(title = "Detections", label = list(size = 40)), name = "Detections") +  # Color gradient with yellow as low and red as high
  scale_size(range = c(3, 8), guide = "none") +  # Size range for points
  theme(legend.position = "bottom") +  # Combine legends at the bottom
  coord_sf(xlim = c(-120.62, -119.27), ylim = c(33.81, 34.42)) +  # Set boundaries
  theme_map() +
  theme(
    legend.title = element_text(size = 25),
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
    panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),  # Add gridlines with decreased width
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    axis.text.y = element_text(color = "black", size = 25),  # Add labels for latitude
    legend.text = element_text(size = 15)  # Increase text size for legend
  ) +
  labs(title = NULL, x = "Longitude", y = "Latitude") +
  scale_x_continuous(breaks = seq(-120.4, -119.6, by = 0.4)) +
  scale_y_continuous(breaks = seq(33.9, 34.4, by = 0.4))

print(map_total_detections_no_sb)

# Export the map as an image file
ggsave("outputs/map_total_detections_no_sb.png", plot = map_total_detections_no_sb, width = 15, height = 10, dpi = 300, units = "in")

# Now make the same map with detections associated with santa barbara filtered out
map_total_detections_sb <- ggplot() +
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = sanctuary, fill = "gray80", color = "black") +  # Plot the islands
  geom_point(data = detections_per_station %>% filter(Island == "Santa Barbara"), aes(x = Longitude, y = Latitude, color = Detections, size = Detections), shape = 16) +
  scale_color_viridis_c(limits = c(0, 200000), guide = guide_colorbar(title = "Detections"), name = "Detections") +  # Color gradient with appropriate limits and colorbar title
  scale_size(range = c(3, 8), name = "Detections") +  # Size range for points
  theme(legend.position = "bottom") +  # Combine legends at the bottom
  coord_sf(xlim = c(-119.015, -119.07), ylim = c(33.46, 33.5)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = NULL, x = "Longitude", y = "Latitude")

print(map_total_detections_sb)

```

# Summary Stats for species

## For each individual transmitter, we want to find the days at liberty so that we can standardize the detections and compare movement and behavior without bias from the length of time the transmitter was active in the study.

### Days at liberty is calculated:
\text{{Days at liberty}} = \text{{last detection date}} - \text{{tagging date}}

```{r days-at-liberty, echo=TRUE, eval=TRUE}

# First, we need to convert First_Detection, Last_Detection, and when_encountered to Date format in the WCO dataframe
WCO <- WCO %>%
  mutate(First_Detection = as.Date(First_Detection, format = "%Y-%m-%d"),
         Last_Detection = as.Date(Last_Detection, format = "%Y-%m-%d"),
         when_encountered = as.Date(when_encountered, format = "%Y-%m-%d"))


# Calculate the days at liberty for each individual transmitter
WCO <- WCO %>%
  mutate(Days_at_liberty = Last_Detection - when_encountered)

  # Let's see which unique SciNames we have a Days_at_liberty for and which are NA
unique_days_at_liberty <- WCO %>%
  group_by(SciName) %>%
  summarise(Days_at_liberty = sum(!is.na(Days_at_liberty)))

print(unique_days_at_liberty)
```

## We are going to look deeper at some general trends including the Residency Index (RI) among groups and the maximum daily displacement for each individual. 

### Calculate the Residency Index (RI) for each species:
#### As described in Gong et al. (2023) the Residency Index (RI) is calculated as: 

```{r setup, echo=FALSE}
reticulate::py_install("matplotlib", envname = "r-reticulate")
```

```{python}

import matplotlib.pyplot as plt

# Corrected LaTeX equation without $ symbols
equation = r"$\text{RI} = \frac{\text{Days detected at a particular station}}{\text{Total detection days}}$"

# Create a figure and axis with no frame
fig = plt.figure(figsize=(10, 2))
ax = fig.add_subplot(111)
ax.axis('off')

# Render the equation
ax.text(0.5, 0.5, equation, fontsize=15, ha='center', va='center')

# Save the figure as an image
fig.savefig("outputs/equation_RI.png", dpi=300, bbox_inches='tight')
```

#### The RI was "calculated regionally for each individual and then averaged by species, where an RI of 0 indicated low residence and an RI of 1 indicated high residence." Here, we will calculate the RI for each individual and then evaluate trends among groups (sex, life stage, species).

```{r RI, echo=TRUE, eval=TRUE}

# Calculate the number of unique days each individual was detected at each station
station_days <- WCO %>%
  group_by(Transmitter, Station.Name) %>%
  summarise(Days_at_Station = n_distinct(Date), .groups = "drop")

# Calculate the total number of unique days each individual was detected
total_days <- WCO %>%
  group_by(Transmitter) %>%
  summarise(Total_Days = n_distinct(Date), .groups = "drop")

# Join the two dataframes together
RI_WCO <- left_join(station_days, total_days, by = "Transmitter")

# Calculate the Residency Index (RI) for each individual at each station
RI_WCO <- RI_WCO %>%
  mutate(RI = Days_at_Station / Total_Days)

  # Check to make sure RI was calculated correctly by looking at summary statistics
summary(RI_WCO$RI)

# Add RI column to the WCO dataframe
WCO <- WCO %>%
  left_join(RI_WCO %>% dplyr::select(Transmitter, Station.Name, RI), by = c("Transmitter", "Station.Name"))

  ```

  ### Calculate the Maximum daily displacement for each individual:

  #### The maximum daily displacement is calculated as the maximum distance detected among transmitters on a given day for each individual. 

  ```{r MDD, echo=TRUE, eval=TRUE}

# Calculate the distance between consecutive detections for each individual on each day
distances <- WCO %>%
  arrange(Transmitter, Date) %>%
  group_by(Transmitter, Date) %>%
  filter(!is.na(Longitude), !is.na(Latitude)) %>%
  mutate(Distance = distVincentySphere(matrix(c(Longitude, Latitude), ncol = 2), 
                                       matrix(c(lead(Longitude), lead(Latitude)), ncol = 2)) / 1000)

# Find the total distance swam for each individual on each day
total_daily_displacement <- distances %>%
  group_by(Transmitter, Date) %>%
  summarise(Total_Distance = sum(Distance, na.rm = TRUE), .groups = "drop")

# Filter out any unrealistic values
total_daily_displacement <- total_daily_displacement %>%
  filter(Total_Distance < 2000)

# Check to make sure total_daily_displacement was calculated correctly by looking at summary statistics. 
summary(total_daily_displacement$Total_Distance)

# Assign each transmitter a maximum daily displacement in the Filtered_Carcharodon dataframe
WCO <- WCO %>%
  left_join(total_daily_displacement, by = c("Transmitter", "Date"))
  
  ```

  ## Summary Stats for species

  ### For the species in the WCO data, we will look at the Residency Index (RI) and Maximum Daily Displacement (MDD). 
  ```{r summary-stats-species, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

install.packages("ggforce")
  library(ggforce)

  # Reorder the factor levels for SciName
WCO$SciName <- factor(WCO$SciName, levels = c("Galeorhinus galeus", 
                         "Carcharodon carcharias", "Triakis semifasciata", "Stereolepis gigas", "Myliobatis californica",
                         "Notorynchus cepedianus",
                         "Seriola lalandi"))

# Histogram of Residency Index (RI) for each species
RI_hist <- ggplot(WCO %>% filter(!is.na(SciName) & Island != "Santa Barbara"), aes(x = RI)) +
  geom_histogram(fill = "#095558", color = "black", bins = 30) +
  facet_wrap(~SciName, scales = "free", ncol = 3, 
             labeller = labeller(SciName = c("Triakis semifasciata" = "Leopard Shark", 
                                             "Stereolepis gigas" = "Giant Sea Bass", 
                                             "Notorynchus cepedianus" = "Sevengill Shark", 
                                             "Galeorhinus galeus" = "Soupfin Shark", 
                                             "Seriola lalandi" = "Yellowtail Amberjack", 
                                             "Myliobatis californica" = "Bat Ray", 
                                             "Carcharodon carcharias" = "White Shark"))) +
  theme_classic() +
  labs(title = NULL, x = "Residency Index", y = "Frequency") +
  theme(text = element_text(family = "Arial", size = 25),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
        axis.title = element_text(size = 30),
        axis.text.y = element_text(size = 25),
        strip.text = element_text(face = "bold")) + # Make species names bold
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(limits = c(0, 1.05), breaks = seq(0, 1, by = 0.25), expand = expansion(mult = c(0, 0.05))) +
  geom_vline(data = WCO %>% filter(!is.na(SciName)) %>% group_by(SciName) %>% summarize(Average_RI = mean(RI)), aes(xintercept = Average_RI), color = "red", linetype = "solid", linewidth = 2) +
  geom_vline(data = WCO %>% filter(!is.na(SciName)) %>% group_by(SciName) %>% summarize(Median_RI = median(RI)), aes(xintercept = Median_RI), color = "blue", linetype = "solid", linewidth = 2)


print(RI_hist)

ggsave("outputs/RI_Hist_pub.png", plot = RI_hist, width = 15, height = 10, dpi = 300, units = "in")

# Let's run an ANOVA test to see if the RI values are signifigantly different among species


# Print the results
print(shapiro_test)

# Histogram of Maximum Daily Displacement (MDD) for each species
MDD_hist <- ggplot(WCO %>% filter(SciName %in% c("Carcharodon carcharias", "Galeorhinus galeus", "Myliobatis californica")), aes(x = Total_Distance)) +
  geom_histogram(fill = "#095558", color = "black", bins = 30) +
  facet_wrap(~SciName, scales = "free") +
  theme_classic() +
  labs(title = NULL, x = "Cumulative Daily Distance (km)", y = "Frequency") +
  theme(text = element_text(family = "Baskerville", size = 25),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
        axis.title = element_text(size = 30),
        axis.text.y = element_text(size = 25)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))

print(MDD_hist)

ggsave("outputs/Daily_Distance.png", plot = MDD_hist, width = 15, height = 10, dpi = 300, units = "in")

```

# Represent the same RI data as a fasceted map for each species
```{r RI-map, echo=TRUE, eval=TRUE}

  # Read the shapefiles
sanctuary <- st_read("data/cinms_py2/cinms_py.shp")
coast <- st_read("data/stanford-vx275xn8886-shapefile/vx275xn8886.shp")
islands <- st_read("data/California_Counties/California_Counties.shp")

# Reorder the factor levels for SciName
WCO$SciName <- factor(WCO$SciName, levels = c("Galeorhinus galeus", 
                         "Carcharodon carcharias", "Triakis semifasciata", "Stereolepis gigas", "Myliobatis californica",
                         "Notorynchus cepedianus",
                         "Seriola lalandi"))

# Create the base map with faceting by SciName
map_RI_faceted <- ggplot() +
  geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
  geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
  geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
  geom_point(data = WCO %>% filter(!is.na(SciName) & SciName != "Pseudobatos productus"), aes(x = Longitude, y = Latitude, color = RI, size = RI), shape = 16) +
  scale_color_gradient(low = "blue", high = "red", guide = guide_colorbar(title = "Residency Index", label = list(size = 40), direction = "horizontal", barwidth = 15, barheight = 3, title.position = "top", title.hjust = 0.5, title.vjust = 0.5), name = "Residency Index") +
  scale_size(range = c(3, 8), guide = "none") +
  facet_wrap(~SciName, ncol = 3, labeller = labeller(SciName = c("Triakis semifasciata" = "Leopard Shark", 
                                                                 "Stereolepis gigas" = "Giant Sea Bass", 
                                                                 "Notorynchus cepedianus" = "Sevengill Shark", 
                                                                 "Galeorhinus galeus" = "Soupfin Shark", 
                                                                 "Seriola lalandi" = "Yellowtail Amberjack", 
                                                                 "Myliobatis californica" = "Bat Ray", 
                                                                 "Carcharodon carcharias" = "White Shark"))) +
  coord_sf(xlim = c(-120.62, -119.27), ylim = c(33.81, 34.42)) +
  theme_map() +
  theme(
    legend.position.inside = c(0.95, 0.05),
    legend.justification = c(1, 0),
    legend.box.just = "right",
    legend.title = element_text(size = 25),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
    panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(color = "black", size = 20),
    legend.text = element_text(size = 20),
    strip.text = element_text(face = "bold", size = 20)
  ) +
  labs(title = NULL, x = "Longitude", y = "Latitude") +
  scale_x_continuous(breaks = seq(-120.4, -119.6, by = 0.4)) +
  scale_y_continuous(breaks = seq(33.9, 34.4, by = 0.4))

print(map_RI_faceted)

# Export the map as an image file
ggsave("outputs/map_RI_faceted.png", plot = map_RI_faceted, width = 15, height = 10, dpi = 300, units = "in")

# Let's find the detections during the highest MDD day for transmitter A69-9002-4561 and oserve the location and time of each detection on a map that day. 

# Filter the data for the highest MDD day for transmitter A69-9002-4561
highest_MDD_day <- WCO %>%
  filter(Transmitter == "A69-9002-4561") %>%
  group_by(Date) %>%
  summarise(Max_MDD = max(Total_Distance)) %>%
  filter(Max_MDD == max(Max_MDD))

  highest_MDD_day

# Highest MDD day for transmitter A69-9002-4561 is 2021-10-30

## Make a map of the detections for the highest MDD day for transmitter A69-9002-4561

# Filter the data for the highest MDD day for transmitter A69-9002-4561
highest_MDD_day_data <- WCO %>%
  filter(Transmitter == "A69-9002-4561", Date == "2021-10-30") %>%
  filter(!is.na(Longitude), !is.na(Latitude)) %>%
  arrange(Local_Time) %>%
  select(Local_Time, Latitude, Longitude)

  highest_MDD_day_data

# Create the map of detections for the highest MDD day for transmitter A69-9002-4561
map_highest_MDD_day <- ggplot() +
  geom_sf(data = coast, color = "#D7E6F9", size = 1.5) +
  geom_sf(data = sanctuary, fill = "#D7E6F9", color = "grey40") +
  geom_sf(data = islands, fill = "#B2B2B2", color = "grey40") +
  geom_point(data = highest_MDD_day_data[1:2, ], aes(x = Longitude, y = Latitude, color = as.factor(1:2)), size = 7) +
  scale_color_manual(values = c("green", "red"), guide = guide_legend(title = "Point Order")) +
  coord_sf(xlim = c(-120.62, -119.27), ylim = c(33.81, 34.42)) +
  theme_map() +
  theme(
    legend.position = "bottomleft",
    legend.title = element_text(size = 25),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
    panel.grid.major = element_line(color = "black", linetype = "solid", size = 0.5),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(color = "black", size = 20),
    legend.text = element_text(size = 20)
  ) +
  labs(title = NULL, x = "Longitude", y = "Latitude") +
  scale_x_continuous(breaks = seq(-120.4, -119.6, by = 0.4)) +
  scale_y_continuous(breaks = seq(33.9, 34.4, by = 0.4))

print(map_highest_MDD_day)

# Export the map as an image file
ggsave("outputs/map_highest_MDD_day.png", plot = map_highest_MDD_day, width = 15, height = 10, dpi = 300, units = "in")

```

### For "Carcharodon carcharias", "Galeorhinus galeus", and "Myliobatis californica", let's take a look at the detections per season
```{r species-seasonal, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}

# Average RI per season with SE
average_ri_season <- WCO %>%
  filter(SciName %in% c("Carcharodon carcharias", "Galeorhinus galeus", "Myliobatis californica")) %>%
  group_by(SciName, Season) %>%
  summarise(Average_RI = mean(RI), SE = sd(RI)/sqrt(n()), .groups = "drop")

  # Plot average RI per season
  species_seasonal_ri_plot <- ggplot(average_ri_season, aes(x = Season, y = Average_RI, fill = SciName)) +
    geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "black") +
    geom_errorbar(aes(ymin = Average_RI - SE, ymax = Average_RI + SE), width = 0.2, position = position_dodge(width = 0.7)) +
    theme_classic() +
    labs(title = NULL, x = "Season", y = "Residency Index") +
    theme(text = element_text(family = "Baskerville", size = 30),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
          axis.title = element_text(size = 30),
          axis.text.y = element_text(size = 25)) +
    scale_x_discrete(limits = c("Spring", "Summer", "Fall", "Winter")) +
    scale_fill_manual(values = c("#93E9BE", "#095558", "#71FBFF"),
                      labels = c("White Shark", "Soupfin Shark", "Bat Ray")) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    labs(fill = "Species")

print(species_seasonal_ri_plot)

# Save plot as image file
ggsave("outputs/Species_Seasonal_RI.png", plot = species_seasonal_ri_plot, width = 15, height = 10, dpi = 300, units = "in")

# Set up Month Column in the WCO dataframe

## Convert Month column to character
WCO$Month <- as.character(WCO$Month)

# Replace the month numbers with the month names
WCO$Month <- recode(WCO$Month, "1" = "January",
                    "2" = "February",
                    "3" = "March",
                    "4" = "April",
                    "5" = "May",
                    "6" = "June",
                    "7" = "July",
                    "8" = "August",
                    "9" = "September",
                    "10" = "October",
                    "11" = "November",
                    "12" = "December")

# Set month order to be January to December
WCO$Month <- factor(WCO$Month, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))

## Check the updated Month column
unique(WCO$Month)

# Average RI per Month with SE
average_ri_month <- WCO %>%
  filter(SciName %in% c("Carcharodon carcharias", "Galeorhinus galeus", "Myliobatis californica")) %>%
  group_by(SciName, Month) %>%
  summarise(Average_RI = mean(RI), SE = sd(RI)/sqrt(n()), .groups = "drop")

# Plot average RI per month
species_monthly_ri_plot <- ggplot(average_ri_month, aes(x = Month, y = Average_RI, fill = SciName)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = Average_RI - SE, ymax = Average_RI + SE), width = 0.2, position = position_dodge(width = 0.7)) +
  theme_classic() +
  labs(title = NULL, x = "Month", y = "Residency Index") +
  theme(text = element_text(family = "Baskerville", size = 30),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
        axis.title = element_text(size = 30),
        axis.text.y = element_text(size = 25)) +
  scale_fill_manual(values = c("#93E9BE", "#095558", "#71FBFF"),
                    labels = c("White Shark", "Soupfin Shark", "Bat Ray")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    labs(fill = "Species")


print(species_monthly_ri_plot)

# Save plot as image file
ggsave("outputs/Species_Monthly_RI.png", plot = species_monthly_ri_plot, width = 15, height = 10, dpi = 300, units = "in")

## Convert Year column to character
WCO$Year <- as.character(WCO$Year)

# Average RI per Year with SE
average_ri_year <- WCO %>%
  filter(SciName %in% c("Carcharodon carcharias", "Galeorhinus galeus", "Myliobatis californica")) %>%
  group_by(SciName, Year) %>%
  summarise(Average_RI = mean(RI), SE = sd(RI)/sqrt(n()), .groups = "drop")

# Plot average RI per year
species_yearly_ri_plot <- ggplot(average_ri_year, aes(x = Year, y = Average_RI, fill = SciName)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = Average_RI - SE, ymax = Average_RI + SE), width = 0.2, position = position_dodge(width = 0.7)) +
  theme_classic() +
  labs(title = NULL, x = "Year", y = "Residency Index") +
  theme(text = element_text(family = "Baskerville", size = 30),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 25),
        axis.title = element_text(size = 30),
        axis.text.y = element_text(size = 25)) +
  scale_fill_manual(values = c("#93E9BE", "#095558", "#71FBFF"),
                    labels = c("White Shark", "Soupfin Shark", "Bat Ray")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    labs(fill = "Species")

print(species_yearly_ri_plot)

# Save plot as image file
ggsave("outputs/Species_Yearly_RI.png", plot = species_yearly_ri_plot, width = 15, height = 10, dpi = 300, units = "in")

```

### Summary Stats for table comparing species
```{r summary-table-species, echo=TRUE, eval=TRUE}

# Summary table for species

Species_table <- WCO %>%
  filter(!is.na(SciName)) %>%
  group_by(SciName) %>%
  summarise(
    Min_Days_at_Liberty = min(Days_at_liberty, na.rm = TRUE),
    Max_Days_at_Liberty = max(Days_at_liberty, na.rm = TRUE),
    Max_Detection_Span = max(as.numeric(Last_Detection - First_Detection), na.rm = TRUE), 
    Min_Detection_Span = min(as.numeric(Last_Detection - First_Detection), na.rm = TRUE),
    Average_RI = mean(RI, na.rm = TRUE),
    SE_RI = sd(RI, na.rm = TRUE) / sqrt(n()),
    Max_MDD = max(Total_Distance, na.rm = TRUE)
  ) %>%
  arrange(desc(Max_MDD))

Species_table

Species_table_filtered <- WCO %>%
  filter(!is.na(SciName) & Days_at_liberty < 11000) %>%
  group_by(SciName) %>%
  summarise(
    Min_Days_at_Liberty = min(Days_at_liberty, na.rm = TRUE),
    Max_Days_at_Liberty = max(Days_at_liberty, na.rm = TRUE),
    Max_Detection_Span = max(as.numeric(Last_Detection - First_Detection), na.rm = TRUE) / 365, # Convert to years
    Min_Detection_Span = min(as.numeric(Last_Detection - First_Detection), na.rm = TRUE) / 365, # Convert to years
    Average_RI = mean(RI, na.rm = TRUE),
    SE_RI = sd(RI, na.rm = TRUE) / sqrt(n()),
    Max_MDD = max(Total_Distance, na.rm = TRUE)
  ) %>%
  arrange(desc(Max_MDD))

# I want to look a a summary of days at liberty for each transmitter among the Carcharodon carcharias species
Carcharodon_days_at_liberty <- WCO %>%
  filter(SciName == "Carcharodon carcharias" & !is.na(Days_at_liberty) & Days_at_liberty < 11000) %>% 
  group_by(Transmitter) %>%
  summarise(Days_at_liberty = max(Days_at_liberty),
            when_encountered = max(when_encountered),
            Last_detection = max(Last_Detection)) %>%
  arrange(desc(Days_at_liberty)) %>%
  ungroup()

Carcharodon_days_at_liberty


```

### Summary Stats for each species

```{r summary-table-species, echo=TRUE, eval=TRUE}

# Summary table for species

# Filter the dataset for the specific species
Carcharodon_RI_Table <- WCO %>%
  filter(SciName == "Carcharodon carcharias") %>%
  group_by(Transmitter) %>%
  summarise(
    Max_Detection_Span = max(as.numeric(Last_Detection - First_Detection, na.rm = TRUE))/365, # Convert to years
    Average_RI = mean(RI, na.rm = TRUE),
    NA,
    SE_RI = sd(RI, na.rm = TRUE) / sqrt(n()),
    Max_MDD = max(Total_Distance, na.rm = TRUE),
    Sex = unique(sex),
    TNL_cm = unique(TNL_cm)
  ) %>%
  filter(Max_Detection_Span >= 0.02) %>%
  arrange(desc(Max_MDD))


Carcharodon_RI_Table

# Export the table to a CSV file
write.csv(Carcharodon_RI_Table, "outputs/Carcharodon_RI_Table.csv")

# Filter the dataset for the specific species
Galeorhinus_RI_Table <- WCO %>%
  filter(SciName == "Galeorhinus galeus") %>%
  group_by(Transmitter) %>%
  summarise(
    Max_Detection_Span = max(as.numeric(Last_Detection - First_Detection, na.rm = TRUE))/365, # Convert to years
    Average_RI = mean(RI, na.rm = TRUE),
    NA,
    SE_RI = sd(RI, na.rm = TRUE) / sqrt(n()),
    Max_MDD = max(Total_Distance, na.rm = TRUE),
    Sex = unique(sex),
    TNL_cm = unique(TNL_cm)
  ) %>%
  filter(Max_Detection_Span >= 0.02) %>%
  arrange(desc(Max_MDD))

Galeorhinus_RI_Table

# Export the table to a CSV file
write.csv(Galeorhinus_RI_Table, "outputs/Galeorhinus_RI_Table.csv")

# Filter the dataset for the specific species
Myliobatis_RI_Table <- WCO %>%
  filter(SciName == "Myliobatis californica") %>%
  group_by(Transmitter) %>%
  summarise(
    Max_Detection_Span = max(as.numeric(Last_Detection - First_Detection, na.rm = TRUE))/365, # Convert to years
    Average_RI = mean(RI, na.rm = TRUE),
    NA,
    SE_RI = sd(RI, na.rm = TRUE) / sqrt(n()),
    Max_MDD = max(Total_Distance, na.rm = TRUE),
    Sex = unique(sex),
    disk_length = unique(disk_length)
  ) %>%
  filter(Max_Detection_Span >= 0.02) %>%
  arrange(desc(Max_MDD))


Myliobatis_RI_Table

# Export the table to a CSV file
write.csv(Myliobatis_RI_Table, "outputs/Myliobatis_RI_Table.csv")

# Let's also make a .csv table for Myliobatis californica ID's to share with the Lowe Lab team
Myliobatis_IDs <- WCO %>%
  filter(SciName == "Myliobatis californica") %>%
  select(Transmitter) %>%
  distinct() %>%
  arrange(Transmitter)

  Myliobatis_IDs

# Export the table to a CSV file
write.csv(Myliobatis_IDs, "outputs/Myliobatis_IDs.csv")
```

# Now, we will investigate our research question for Carcharodon carcharias

## Research Question: Do sexually mature male and female white sharks share habitats within the CINMS?

### We will create an abacus plot to visualize the movement patterns of Carcharodon carcharias that are more than 330 TNL_cm in the CINMS. The plot will be binned by month with the islands on the y axis and months on the x axis. Colors will represent sex of individuals. 

```{r abacus-Carcharodon, echo=TRUE, eval=TRUE}

# Filter for Carcharodon carcharias with non-NA sex, sex is M or F, and TNL_cm greater than 180
Filtered_Carcharodon_adult <- WCO %>%
  filter(SciName == "Carcharodon carcharias", !is.na(sex), sex %in% c("M", "F"), TNL_cm > 180)

# Create a vector of custom colors for sex of individuals
custom_colors <- c("#FF7F50", "#095558")

# Create abacus plot with the Month on the x axis and island on the y axis in order (Santa Barbara, Mainland, Anacapa, Santa Cruz, Santa Rosa, San Miguel) and all the individuals labeled with unique colors.
Carcharodon_abacus <- ggplot(Filtered_Carcharodon_adult, aes(x = Month, y = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Mainland", "Santa Barbara")), color = sex, size = RI)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point() +
  scale_color_manual(values = custom_colors) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
        axis.text.y = element_text(size = 12), # Increase size of y-axis text
        axis.title = element_text(size = 14), # Increase size of axis titles
        plot.title = element_text(hjust = 0.5, size = 16), # Increase size of plot title
        legend.text = element_text(size = 12), # Increase size of legend text
        legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Carcharodon carcharias (Adults)",
       x = "Month",
       y = "Island",
       color = "Sex",
       size = "RI") +
  guides(color = guide_legend(title = "Sex"),
         size = guide_legend(title = "RI")) +
  scale_x_discrete(limits = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")) +
  scale_y_discrete(limits = c("San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Mainland", "Santa Barbara")) +
  theme(legend.position = "bottom") +
  theme(legend.box = "horizontal") +
  scale_size(range = c(0.5, 3))

print(Carcharodon_abacus)

# Save plot as image file
ggsave("outputs/Abacus_Plot_Carcharodon.png", plot = Carcharodon_abacus, width = 15, height = 10, dpi = 300, units = "in")

```


  ```{r summary-stats-species, echo=TRUE, eval=TRUE}

# Abacus Plots for Species

## Myliobatis californica
```{r abacus-Myliobatis, echo=TRUE, eval=TRUE}

# check the number of transmitters associated with Mylobatis californica
Number_of_Myliobatis <- WCO %>%
  filter(SciName == "Myliobatis californica") %>%
  summarise(Transmitter_count = n_distinct(Transmitter))

  print(Number_of_Myliobatis)

# Filter the dataset for the specific species
Filtered_Myliobatis <- WCO %>%
  filter(SciName == "Myliobatis californica")

# Create a vector of custom colors and shapes for each individual
custom_colors <- c("#C58ABB", "#B73944", "#8ACE89", "#C15901", "#0072B2", 
                   "#656565", "#009E73", "#56B4E9", "#E69F00", "#000000")
shapes <- c(15, 16, 17, 18, 19, 20, 21, 22, 23, 24)

# Convert 'Date' from character to Date format in the Filtered_Myliobatis dataframe
Filtered_Myliobatis <- Filtered_Myliobatis %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.

# Filter data for pre 2020
Filtered_Myliobatis_pre_2020 <- Filtered_Myliobatis %>% 
  filter(Date < as.Date("2020-01-01"))

# Plot for pre 2020
mylobatis_abacus_pre_2020 <- ggplot(Filtered_Myliobatis_pre_2020, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 month"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
        axis.text.y = element_text(size = 12), # Increase size of y-axis text
        axis.title = element_text(size = 14), # Increase size of axis titles
        plot.title = element_text(hjust = 0.5, size = 16), # Increase size of plot title
        legend.text = element_text(size = 12), # Increase size of legend text
        legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Myliobatis californica (Pre 2020)",
       x = "Island",
       y = "Date",
       color = "Ray",
       shape = "Ray")

print(mylobatis_abacus_pre_2020)

ggsave("outputs/Abacus_Plot_Myliobatis_Pre_2020.png", plot = mylobatis_abacus_pre_2020, width = 10, height = 15, dpi = 300, units = "in")

# Filter data for post 2020
Filtered_Myliobatis_post_2020 <- Filtered_Myliobatis %>% 
  filter(Date >= as.Date("2020-01-01"))

# Plot for post 2020
mylobatis_abacus_post_2020 <- ggplot(Filtered_Myliobatis_post_2020, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 month"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Myliobatis californica (Post 2020)",
       x = "Island",
       y = "Date",
       color = "Ray",
       shape = "Ray")

print(mylobatis_abacus_post_2020)
ggsave("outputs/Abacus_Plot_Myliobatis_Post_2020.png", plot = mylobatis_abacus_post_2020, width = 10, height = 15, dpi = 300, units = "in")

```

## Triakis semifasciata, Seriola lalandi, Stereolepis gigas, and Pseudobatos productus.
```{r abacus-Triakis-Seriola-Stereolepis-Pseudobatos, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Multi_Species <- WCO %>%
  filter(SciName == "Triakis semifasciata" | SciName == "Seriola lalandi" | SciName == "Stereolepis gigas" | SciName == "Pseudobatos productus")

# Create a vector of custom colors and shapes for each species
custom_colors <- c("#C58ABB", "#B73944", "#8ACE89", "#C15901")
shapes <- c(15, 16, 17, 18)

# Convert 'Date' from character to Date format in the Filtered_Multi_Species dataframe
Filtered_Multi_Species <- Filtered_Multi_Species %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where Island is NA
Filtered_Multi_Species <- Filtered_Multi_Species %>%
  filter(!is.na(Island))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the species labeled with unique colors and shapes.
  
  # Plot
Abacus_Multi_Species <- ggplot(Filtered_Multi_Species, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = SciName, shape = SciName)) +
  geom_jitter(width = 0.5, height = 0.3, alpha = 1) +
  geom_point(size = 5) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Triakis semifasciata, Seriola lalandi, Stereolepis gigas, and Pseudobatos productus",
       x = "Island",
       y = "Date",
       color = "Species",
       shape = "Species")

print(Abacus_Multi_Species)

ggsave("outputs/Abacus_Plot_Multi_Species.png", plot = Abacus_Multi_Species, width = 10, height = 15, dpi = 300, units = "in")

```

## Galeorhinus galeus
```{r abacus-Galeorhinus, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Galeorhinus <- WCO %>%
  filter(SciName == "Galeorhinus galeus")

# Create a vector of custom colors and shapes for each individual (We have 31 individuals)
custom_colors <- c("#000000", "grey", "#D7792C", "#D79C2C", "#D7BB2C", 
                   "#D7D42C", "#BFD72C", "#A3D72C", "#7CD72C", "#3AD72C",
                   "#2CD779", "#2CD79F", "#2CD7BF", "#2CCDD7", "#2CA6D7", 
                   "#2C83D7", "#2C60D7", "#2C41D7", "#672CD7", "#8A2CD7",
                   "#AD2CD7", "#CD2CD7", "#D72CA3", "#D72C80", "#D72C60", 
                   "#D72C3A", "#D7592C", "#D7792C", "#D79C2C", "#D7BB2C", "#D7D42C")
shapes <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10,
            1, 2, 3, 4, 5, 6, 7, 8, 9, 10,
            1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11)

# Convert 'Date' from character to Date format in the Filtered_Galeorhinus dataframe
Filtered_Galeorhinus <- Filtered_Galeorhinus %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where Island is NA
Filtered_Galeorhinus <- Filtered_Galeorhinus %>%
  filter(!is.na(Island))

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.
  
  # Plot
  Abacus_Galeorhinus <- ggplot(Filtered_Galeorhinus, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa")), y = Date, color = Transmitter, shape = Transmitter)) +
  geom_jitter(width = 0.3, height = 0.4, alpha = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  guides(color = guide_legend(ncol = 1), shape = guide_legend(ncol = 1)) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Galeorhinus galeus",
       x = "Island",
       y = "Date",
       color = "Shark",
       shape = "Shark")

print(Abacus_Galeorhinus)

ggsave("outputs/Abacus_Plot_Galeorhinus.png", plot = Abacus_Galeorhinus, width = 10, height = 15, dpi = 300, units = "in")

```

## Carcharodon carcharias
```{r abacus-Carcharodon, echo=TRUE, eval=TRUE}

# Filter the dataset for the specific species
Filtered_Carcharodon <- WCO %>%
  filter(SciName == "Carcharodon carcharias")

# Look at the number of unique transmitters for Carcharodon carcharias
Number_of_Carcharodon <- Filtered_Carcharodon %>%
  summarise(Transmitter_count = n_distinct(Transmitter))

print(Number_of_Carcharodon)

# Create a vector of custom shapes to represent sex for the species. 
shapes <- c(1, 3, 2)

# Convert 'Date' from character to Date format in the Filtered_Carcharodon dataframe
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))

# Filter for detections where sex is NA
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  filter(!is.na(sex))

# Filter for detections where TNL_cm is NA
Filtered_Carcharodon <- Filtered_Carcharodon %>%
  filter(!is.na(TNL_cm))

  # Find the minimum and maximum values of 'TNL_cm' for Carcharodon carcharias. 
min_max_TNL_cm <- Filtered_Carcharodon %>%
  summarise(min_TNL_cm = min(TNL_cm), max_TNL_cm = max(TNL_cm))

print(min_max_TNL_cm)

# Format when_encountered as Date
Filtered_Carcharodon$when_encountered <- as.Date(Filtered_Carcharodon$when_encountered, format = "%Y-%m-%d")

# Create column for adjusted 'TNL_cm' values to account for growth. The species grows 45cm per year until they reach 180cm, then 12cm per year until they reach 330cm. We can keep the values above 330cm as is.

Filtered_Carcharodon$TNL_cm_adjusted <- Filtered_Carcharodon$TNL_cm

# Calculate the number of days since the when_encountered date for each detection
Filtered_Carcharodon$days_since_encountered <- as.numeric(difftime(Filtered_Carcharodon$Date, Filtered_Carcharodon$when_encountered, units = "days"))

# Adjust TNL_cm based on the conditions
Filtered_Carcharodon$TNL_cm_adjusted[Filtered_Carcharodon$TNL_cm < 180] <- Filtered_Carcharodon$TNL_cm[Filtered_Carcharodon$TNL_cm < 180] + 0.12328767 * Filtered_Carcharodon$days_since_encountered[Filtered_Carcharodon$TNL_cm < 180]
Filtered_Carcharodon$TNL_cm_adjusted[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331] <- Filtered_Carcharodon$TNL_cm[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331] + 0.03287671 * Filtered_Carcharodon$days_since_encountered[Filtered_Carcharodon$TNL_cm >= 180 & Filtered_Carcharodon$TNL_cm < 331]


  # Create a new variable 'life_stage' based on 'TNL_cm'
Filtered_Carcharodon$life_stage <- cut(Filtered_Carcharodon$TNL_cm_adjusted, 
                                       breaks = c(-Inf, 180, 330, Inf), 
                                       labels = c("YOY", "Juvenile", "Adult"))

# Create a vector of custom colors
custom_colors <- c("YOY" = "#03C41F", "Juvenile" = "#031FC4", "Adult" = "#C4030F")

  # Create abacus plot with time (binned by date) on the y axis and island on the x axis in order based on longitude (from west to east) and all the individuals labeled with unique colors and shapes.
  
  # Plot
  Abacus_Carcharodon <- ggplot(Filtered_Carcharodon, aes(x = factor(Island, levels = c("San Miguel", "Santa Rosa", "Santa Cruz", "Mainland", "Anacapa", "Santa Barbara")), y = Date, color = life_stage, shape = sex)) +
  geom_jitter(width = 0.3, height = 0.4, alpha = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors) +
  scale_shape_manual(values = shapes) +
  guides(color = guide_legend(ncol = 1), shape = guide_legend(ncol = 1)) +
  scale_y_date(breaks = scales::date_breaks("1 year"), labels = scales::date_format("%Y-%m-%d")) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Increase size of x-axis text
      axis.text.y = element_text(size = 12), # Increase size of y-axis text
      axis.title = element_text(size = 14), # Increase size of axis titles
      plot.title = element_text(hjust = 0.5, size = 16), # Correctly set size of plot title
      legend.text = element_text(size = 12), # Increase size of legend text
      legend.title = element_text(size = 14)) + # Increase size of legend title
  labs(title = "Abacus Plot for Carcharodon carcharias",
       x = "Island",
       y = "Date",
       color = "Life Stage",
       shape = "Sex")

print(Abacus_Carcharodon)

# Save the plot
ggsave("outputs/Abacus_Plot_Carcharodon.png", plot = Abacus_Carcharodon, width = 10, height = 15, dpi = 300, units = "in")

# Make a table with the following parameters for each individual: when_encountered, date first detected, date last detected, and detections per life stage with columns for (YOY, Juvenile, Adult).
 Filtered_Carcharodon_table <- Filtered_Carcharodon %>%
  group_by(Transmitter) %>%
  summarise(
    when_encountered = min(when_encountered),
    date_first_detected = min(Date),
    date_last_detected = max(Date),
    TNL_cm = mean(TNL_cm, na.rm = TRUE), # Add TNL_cm column
    YOY_detections = sum(life_stage == "YOY"),
    Juvenile_detections = sum(life_stage == "Juvenile"),
    Adult_detections = sum(life_stage == "Adult")
  ) %>%
  arrange(TNL_cm) # Sort by TNL_cm in ascending order

  print(Filtered_Carcharodon_table)

  # Write the table to a csv file
write.csv(Filtered_Carcharodon_table, "data/Carcharodon_Life_Stage_table.csv", row.names = FALSE)

```

### Make a movie for shark A69-9001-13501 
```{r shark-movie-13501, echo=TRUE, eval=TRUE}

# Filter the data for the specific shark transmitter
A69_9001_13501 <- Filtered_Carcharodon %>% 
  filter(Transmitter == "A69-9001-13501")

# Read the shapefiles
islands <- st_read("data/cinms_py2/cinms_py.shp")
coast <- st_read("data/stanford-vx275xn8886-shapefile/vx275xn8886.shp")

# Create a Daily column for easier plotting
detections_per_station_daily_13501 <- A69_9001_13501 %>%
  mutate(Daily = as.Date(Date)) %>%
  group_by(Longitude, Latitude, Island, Year = year(Daily), Daily) %>%
  summarise(Detections = n(), TNL_cm_adjusted = mean(TNL_cm_adjusted)) %>%
  ungroup()

base_plot_13501 <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_daily_13501, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted), size = 5) +
  scale_color_gradient(low = "yellow", high = "red") +  # Set the color gradient from red to yellow
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Total Length (cm)")

# Create the animation
animation_plot_13501 <- base_plot_13501 +
  geom_point(data = detections_per_station_daily_13501, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted)) +
  transition_states(states = Daily, transition_length = 20, state_length = 20) +
  labs(title = paste("Detections M 13501:", "{closest_state}"))
```

#### Animate 13510 and save
```{r, eval=FALSE}

animate(animation_plot_13501, nframes = length(unique(as.vector(detections_per_station_daily_13501$Daily))), fps = 3, width = 800, height = 600, renderer = gifski_renderer("outputs/13501_animation.gif"))

# Set levels for stations based on longitude (West to East)
station_order <- A69_9001_13501 %>%
  arrange(Longitude) %>%
  pull(Station.Name)

# Remove duplicated levels
station_order <- unique(station_order)

# Convert Island and Station Name to factors with specified order
A69_9001_13501 <- A69_9001_13501 %>%
  mutate(Station.Name = factor(Station.Name, levels = station_order))

# Create abacus for 13510
Abacus_13501 <- ggplot(A69_9001_13501, aes(x = factor(`Station.Name`, levels = station_order), y = Date, color = TNL_cm_adjusted)) +
  geom_point(shape = 19, size = 5) +
  scale_color_viridis_c(option = "plasma", direction = -1) +
  labs(title = "Abacus Plot for 13510", x = "Station", y = "Date", color = "TNL") +
  theme_minimal() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_rect(fill = "darkgrey"),
        plot.background = element_rect(fill = "darkgrey"))

print(Abacus_13501)
ggsave("outputs/Abacus_Plot_13501.png", plot = Abacus_13501, width = 10, height = 15, dpi = 300, units = "in")

```

## Ok, now that we know how to do this, let's look at some other Charcarodon carcharias transmitters that may be of interest.

### Identify the unique transmitters for Carcharodon carcharias with the greatest detection span and the greatest number of detections.
```{r shark-summary, echo=TRUE, eval=TRUE}

# Find the unique transmitters for Carcharodon carcharias with the greatest detection span
Carcharodon_span <- Filtered_Carcharodon %>%
  group_by(Transmitter) %>%
  summarize(First_Detection = min(Date),
            Last_Detection = max(Date),
            Detection_Span = as.numeric(max(Date) - min(Date)),
            Number_of_Detections = n()) %>%
  arrange(desc(Detection_Span))

View(Carcharodon_span)

```

### Make a movie for shark A69-1601-30029 
```{r shark-movie-30029, echo=TRUE, eval=TRUE}

# Filter the data for the specific shark transmitter
A69_1601_30029 <- Filtered_Carcharodon %>% 
  filter(Transmitter == "A69-1601-30029")

# Create a Daily column for easier plotting
detections_per_station_daily_30029 <- A69_1601_30029 %>%
  mutate(Daily = as.Date(Date)) %>%
  group_by(Longitude, Latitude, Island, Year = year(Daily), Daily) %>%
  summarise(Detections = n(), TNL_cm_adjusted = mean(TNL_cm_adjusted)) %>%
  ungroup()

base_plot_30029 <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_daily_30029, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted), size = 5) +
  scale_color_gradient(low = "yellow", high = "red") +  # Set the color gradient from red to yellow
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Total Length (cm)")

# Create the animation
animation_plot_30029 <- base_plot_30029 +
  geom_point(data = detections_per_station_daily_30029, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted)) +
  transition_states(states = Daily, transition_length = 20, state_length = 20) +
  labs(title = paste("Detections M 30029:", "{closest_state}"))
  ```

#### Animate 30029 and save
```{r, eval=FALSE}
animate(animation_plot_30029, nframes = length(unique(as.vector(detections_per_station_daily_30029$Daily))), fps = 3, width = 800, height = 600, renderer = gifski_renderer("outputs/30029_animation.gif"))

# Set levels for stations based on longitude (West to East)
station_order <- A69_1601_30029 %>%
  arrange(Longitude) %>%
  pull(Station.Name)

# Remove duplicated levels
station_order <- unique(station_order)

# Convert Island and Station Name to factors with specified order
A69_1601_30029 <- A69_1601_30029 %>%
  mutate(Station.Name = factor(Station.Name, levels = station_order))

# Create abacus for 30029
Abacus_30029 <- ggplot(A69_1601_30029, aes(x = factor(`Station.Name`, levels = station_order), y = Date, color = TNL_cm_adjusted)) +
  geom_point(shape = 19, size = 5) +
  scale_color_viridis_c(option = "plasma", direction = -1) +
  labs(title = "Abacus Plot for 30029", x = "Station", y = "Date", color = "TNL") +
  theme_minimal() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_rect(fill = "darkgrey"),
        plot.background = element_rect(fill = "darkgrey"))

print(Abacus_30029)
ggsave("outputs/Abacus_Plot_30029.png", plot = Abacus_30029, width = 10, height = 15, dpi = 300, units = "in")
```

### Make a movie for shark A69-9001-24384 
```{r shark-movie-24384, echo=TRUE, eval=TRUE}

# Filter the data for the specific shark transmitter
A69_9001_24384 <- Filtered_Carcharodon %>% 
  filter(Transmitter == "A69-9001-24384")

# Create a Daily column for easier plotting
detections_per_station_daily_24384 <- A69_9001_24384 %>%
  mutate(Daily = as.Date(Date)) %>%
  group_by(Longitude, Latitude, Island, Year = year(Daily), Daily) %>%
  summarise(Detections = n(), TNL_cm_adjusted = mean(TNL_cm_adjusted)) %>%
  ungroup()

base_plot_24384 <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_daily_24384, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted), size = 5) +
  scale_color_gradient(low = "yellow", high = "red") +  # Set the color gradient from red to yellow
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Total Length (cm)")

# Create the animation
animation_plot_24384 <- base_plot_24384 +
  geom_point(data = detections_per_station_daily_24384, aes(x = Longitude, y = Latitude, color = TNL_cm_adjusted)) +
  transition_states(states = Daily, transition_length = 20, state_length = 20) +
  labs(title = paste("Detections M 24384:", "{closest_state}"))
  ```

#### Animate 24384 and save
```{r, eval=FALSE}
animate(animation_plot_24384, nframes = length(unique(as.vector(detections_per_station_daily_24384$Daily))), fps = 3, width = 800, height = 600, renderer = gifski_renderer("outputs/24384_animation.gif"))

# Set levels for stations based on longitude (West to East)
station_order <- A69_9001_24384 %>%
  arrange(Longitude) %>%
  pull(Station.Name)

# Remove duplicated levels
station_order <- unique(station_order)

# Convert Island and Station Name to factors with specified order
A69_9001_24384 <- A69_9001_24384 %>%
  mutate(Station.Name = factor(Station.Name, levels = station_order))

# Create abacus for 24384
Abacus_24384 <- ggplot(A69_9001_24384, aes(x = factor(`Station.Name`, levels = station_order), y = Date, color = TNL_cm_adjusted)) +
  geom_point(shape = 19, size = 5) +
  scale_color_viridis_c(option = "plasma", direction = -1) +
  labs(title = "Abacus Plot for 24384", x = "Station", y = "Date", color = "TNL") +
  theme_minimal() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_rect(fill = "darkgrey"),
        plot.background = element_rect(fill = "darkgrey"))

print(Abacus_24384)
ggsave("outputs/Abacus_Plot_24384.png", plot = Abacus_24384, width = 10, height = 15, dpi = 300, units = "in")

```

### Make a movie for all three sharks 
```{r shark-movie-all, echo=TRUE, eval=TRUE}

# Filter the data for the specific shark transmitters
male_cohort <- Filtered_Carcharodon %>% 
  filter(Transmitter %in% c("A69-1601-30029", "A69-9001-24384", "A69-9001-13501"))

# Create a Daily column for easier plotting
detections_per_station_daily_mcoh <- male_cohort %>%
  mutate(Daily = as.Date(Date)) %>%
  group_by(Longitude, Latitude, Island, Year = year(Daily), Daily, Transmitter) %>%
  summarise(Detections = n(), TNL_cm_adjusted = mean(TNL_cm_adjusted)) %>%
  ungroup()

base_plot_mcoh <- ggplot() + 
  geom_sf(data = coast, color = "black") +  # Add the coastline shapefile
  geom_sf(data = islands, fill = "gray80", color = "black") +  # Plot the shapefile
  geom_point(data = detections_per_station_daily_mcoh, aes(x = Longitude, y = Latitude, color = Transmitter), size = 5) +
  coord_sf(xlim = c(-120.65, -119.25), ylim = c(33.8, 34.42)) +  # Set boundaries
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),  # White background
    plot.background = element_rect(fill = "white", color = NA),   # White plot background
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Longitude", y = "Latitude", color = "Shark")

# Create the animation
animation_plot_mcoh <- base_plot_mcoh +
  geom_point(data = detections_per_station_daily_mcoh, aes(x = Longitude, y = Latitude, color = Transmitter)) +
  transition_states(states = Daily, transition_length = 20, state_length = 20) +
  labs(title = paste("Detections Male Cohort:", "{closest_state}"))
```

#### Animate all three sharks and save
```{r, eval=FALSE}
animate(animation_plot_mcoh, nframes = length(unique(as.vector(detections_per_station_daily_mcoh$Daily))), fps = 3, width = 800, height = 600, renderer = gifski_renderer("outputs/male_cohort_animation.gif"))

# Set levels for stations based on longitude (West to East)
station_order <- male_cohort %>%
  arrange(Longitude) %>%
  pull(Station.Name)

# Remove duplicated levels
station_order <- unique(station_order)

# Convert Island and Station Name to factors with specified order
male_cohort <- male_cohort %>%
  mutate(Station.Name = factor(Station.Name, levels = station_order))

# Create abacus for all
Abacus_mcoh <- ggplot(male_cohort, aes(x = factor(`Station.Name`, levels = station_order), y = Date, color = Transmitter)) +
  geom_point(shape = 19, size = 5) +
  labs(title = "Abacus Plot Male Cohort", x = "Station", y = "Date", color = "Shark") +
  theme_minimal() +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_rect(fill = "darkgrey"),
        plot.background = element_rect(fill = "darkgrey"))

print(Abacus_mcoh)
ggsave("outputs/Abacus_Male_Cohort.png", plot = Abacus_mcoh, width = 10, height = 15, dpi = 300, units = "in")
```

